### 2016-06-01 DraftOne.Rmd
First commit (test)

---
title: "DraftOne"
author: "Sarah Koehler"
date: "June 1, 2016"
output: html_document
---

###First draft of eesim.
```{r}
sum(1:4)
```

### 2016-06-01 DraftOne.Rmd
Simulating data with seasonal trends

@@
 
-###First draft of eesim.
+```{r echo = FALSE, message=FALSE, warning=FALSE}
+library(ggplot2)
+library(dplyr)
+library(splines)
+```
+
+# Methods
+
+## Generating the simulated data
+
+### Exposure data
+
+There are several different possible ways to generate the simulated exposure data (or to pull in observed data to use for this).
+
+For all of the simulated exposure series, this starts by specifying the number of days to simulate, `n`:
+
 ```{r}
-sum(1:4)
+n <- 1000
 ```
+
+#### Binary exposure
+
+For a binary exposure, then it's necessary to specify the probability of exposure, `p_exp`. Then it's pretty straightforward to generate the series of exposures:
+
+```{r}
+# Generate a binary exposure
+p_exp <- 0.1
+x <- sample(c(0, 1), size = n, replace = TRUE, prob = c(1-p_exp, p_exp))
+head(x)
+```
+
+Here's a function that takes arguments of `n` and `p_exp` and returns a generated series (I'm also adding in a part to put in dates-- unless I specify otherwise, this will default to start at Jan.
+1, 2000):
+
+```{r}
+# Function to generate a binary exposure
+binary_exposure <- function(n, p_exp, start.date = "2000-01-01", ...){
+        start.date <- as.Date(start.date)
+        date <- seq(from = start.date, by = 1, length.out = n)
+        x <- sample(c(0, 1), size = n, replace = TRUE,
+                    prob = c(1-p_exp, p_exp))
+        df <- data.frame(date, x)
+        return(df)
+}
+binary_exposure(n = 5, p_exp = 0.25)
+```
+
+#### Continuous exposure
+
+For a continuous exposure (which I'll assume it's okay to assume is normally distributed), it's necessary to specify the mean, `mu`, and standard deviation, `sd`, of the distribution of the exposure.
+Then it's possible to simulate the data:
+
+```{r}
+# Generate a continuous exposure
+mu <- 10 # These values are both from the Chicago NMMAPS values for 
+sd <- 10 # temperature in degrees Celcius
+x <- rnorm(n, mean = mu, sd = sd)
+head(x)
+```
+
+```{r}
+# Function to generate a continuous exposure
+continuous_exposure <- function(n, mu, sd, start.date = "2000-01-01", ...){
+        start.date <- as.Date(start.date)
+        date <- seq(from = start.date, by = 1, length.out = n)
+        x <- rnorm(n, mean = mu, sd = sd)
+        df <- data.frame(date, x)
+        return(df)
+}
+continuous_exposure(n = 5, mu = 10, sd = 10)
+```
+
+#### Generate exposure data with seasonal pattern
+
+```{r}
+# Get Chicago NMMAPS data and use CVD
+library(dlnm)
+data(chicagoNMMAPS)
+chic <- chicagoNMMAPS
+names(chic)
+cvd <- chic$cvd
+
+```
+
+#### Pull real data
+
+```{r}
+# Pull exposure with a seasonal pattern
+library(dlnm)
+data(chicagoNMMAPS)
+chic <- chicagoNMMAPS
+
+if(n < nrow(chic)){
+        start <- sample(1:(nrow(chic) - n), 1) # randomly choose a starting point
+        x <- chic$temp[start:(start + n -1)]
+}
+head(x)
+```
+
+```{r}
+# Function to pull exposure with a seasonal pattern
+# Note: Your `data` must have a column called `date` as well as one with the 
+# measure you specify, with that column name.
+pull_exposure <- function(n, data = "chicagoNMMAPS", measure = "temp", ...){
+        if(data == "chicagoNMMAPS"){
+                require(dlnm)
+                data(chicagoNMMAPS)
+                data <- chicagoNMMAPS
+        }
+
+        data$measure <- data[ , measure]
+
+        if(n < nrow(data)){
+                # randomly choose a starting point
+                start <- sample(1:(nrow(data) - n), 1) 
+                df <- data[start:(start + n - 1), c("date", measure)]
+        } else{
+                stop("You asked for too many simulated observations given the dataset you're pulling the data from. Try a smaller value of n.")
+        }
+
+        return(df)
+}
+
+pull_exposure(n = 3)
+pull_exposure(n = 3)
+pull_exposure(n = 3, measure = "pm10")
+```
+
+### Expected mean baseline for outcome data
+
+#### Constant baseline hazard rate
+
+To generate the expected mean baseline for an outcome with a constant baseline hazard rate over time, you just need to specify the $\lambda$ value for the distribution of the outcome (if you're using a Poisson distribution):
+
+```{r}
+# Contant baseline of outcome
+lambda <- 100 # Approximate mean value of daily deaths for chicagoNMMAPS
+exp_base_y <- rep(lambda, n)
+```
+
+```{r}
+constant_baseline <- function(n, lambda, start.date = "2000-01-01", ...){
+        start.date <- as.Date(start.date)
+        date <- seq(from = start.date, by = 1, length.out = n)
+        exp_base_y <- rep(lambda, n)
+        df <- data.frame(date = date,
+                         exp_base_y = exp_base_y)
+        return(df)
+}
+constant_baseline(n = 3, lambda = 100)
+```
+
+#### Baseline hazard rate varies by season
+
+First, we need to generate the explanatory exposure variable.
+Here is a function for binary exposure in which the probability varies by season.
+Later we will add a function for continuous exposure which varies by season (~Normal(mu, sigma)).
+
+```{r}
+# A function for binary exposure which varies by season:
+season_binexp <- function(n, p){
+  p <- p #Change this later to reflect probability varying by season using trends
+  binexp <- sample(c(0,1), size = n, replace = T, prob = c(1-p, p))
+  return(binexp)
+}
+```
+
+Now a function to calculate the trend variable. Choices for trend are cvd, cos1, cos2, cos3, linear, curvilinear, cos1linear.
+
+```{r}
+calc_t <- function(n, trend){
+  day <- c(1:n)
+  if (trend == "cos1"){
+    seasont <- 1+.6*cos(2*pi*(day/365))
+  }
+  if (trend == "cos2"){
+    seasont <- 1+.6*cos(2*pi*(day/365)) + ifelse(day<639 & day>274, .4*cos(2*(pi*(day/365))), 0)
+  }
+  if (trend == "cos3"){
+    seasont <- 1+.75^(day/365)*.6*cos(2*pi*(day/365))
+  }
+  if (trend == "linear"){
+    seasont <- 1+ (day/n)
+  }
+  if (trend == "curvilinear"){
+    seasont <- 1+day*(2/n)+day^2*(-1/n^2)
+  }
+  if (trend == "cos1linear"){
+    seasont <- (1+(day/n)) *(1+.6*cos(2*pi*(day/365)))
+  }
+  if (trend == "cvd"){ #Figure this out later.
+  } #else{
+    #stop("Specify a trend variable from cos1, cos2, cos3, linear, curvilinear, cos1linear, or cvd."
+    #)}
+  return(seasont)
+}
+```
+
+```{r}
+#Make a sweet plot of these with n=1500 (just for fun, not working yet):
+trendname <- c(rep("cos1", 1500), rep("cos2",1500), rep("cos3",1500), rep("linear",1500),  rep("curvilinear",1500), rep("cos1linear",1500))
+trendtests <- c(calc_t(1500, trend = "cos1"), calc_t(1500, trend = "cos2"), calc_t(1500, trend = "cos3"), calc_t(1500, trend = "linear"), calc_t(1500, trend = "curvilinear"), calc_t(1500, trend = "cos1linear"))
+trendframe <- data.frame(trendname, trendtests)
+```
+
+Next, a function to simulate the outcomes.
+
+```{r}
+#Here, exposure would be the output from the season_binexp function and t would be the output from the calc_t function. 
+sim_outcome <- function(n, mean_out, t, exposure, rr){
+  day <- c(1:n)
+  rr <- ifelse(exposure == 1, rr, 1)
+  mean_outcome <- mean_out*t*rr
+  sim_outcome <- rpois(n, mean_outcome)
+  return(sim_outcome)
+}
+```
+
+Putting it all together, here is a function for data with seasonal trends. This function will return a vector of simulated outcomes.
+
+```{r}
+#Make this function output a dataframe with date, x, etc.
+sim_seasdata <- function(n, rr, trend, x_type="binary", p, mean_out){
+  if(x_type == "binary"){
+    x <- season_binexp(n, p)
+  } 
+  else if(x_type == "continuous"){
+
+  }
+  t <- calc_t(n, trend)
+  outcomes <- sim_outcome(n, mean_out, t, x, rr)
+  return(outcomes)
+}
+```
+
+Here are some examples of this function:
+
+```{r}
+sim_seasdata(n = 3, rr = 1.2, trend="cos2", x_type="binary", p=.2, mean_out = 100)
+```
+
+```{r}
+# Outcome with mortality displacement
+
+# Outcome with delayed, slightly elevated risk
+```
+
+### Simulated outcome data
+
+Moving from the expected value of the baseline outcome to simulated outcome data requires a few things. First, it's necessary to specify the true effect associated with the exposure:
+
+```{r}
+# Specify true effect
+rr <- 1.2
+```
+
+Then, it's necessary to combine the exposure, `x`, and expected baseline outcome values, `exp_base_y`, in a dataframe:
+
+```{r}
+n <- 1000
+p_exp <- 0.25
+lambda <- 100
+
+x <- binary_exposure(n = n, p_exp = p_exp)
+exp_base_y <- constant_baseline(n = n, lambda = lambda)
+df <- full_join(x, exp_base_y, by = "date")
+```
+
+Then you can generate the `exp_y` using the equation:
+
+$$
+E(log(\mu_t)) = \beta_0 + \beta_1X_1
+$$
+
+(For this, it will be important that the relative risk is specified for a **one-unit** increase in the exposure value.)
+
+```{r}
+df <- mutate(df, exp_y = exp(log(exp_base_y) + log(rr) * x))
+```
+
+Now, you need to decide on how the outcome data is distributed around the expected value. This is simulating outcome values from a Poisson distribution using `rpois`.
+
+```{r}
+# Simulate outcome data
+df$y <- sapply(df$exp_y, FUN = function(x) rpois(1, x))
+head(df, 3)
+```
+
+```{r fig.width = 2.5, fig.height = 2}
+to_plot <- mutate(df, x = as.factor(x))
+ggplot(to_plot, aes(x = x, y = y)) + 
+        geom_point(position = position_jitter(), alpha = 0.1) + 
+        geom_boxplot( fill = NA) + 
+        theme_bw()
+```
+
+```{r}
+group_by(df, exp_y) %>% summarize(mean_y = mean(y))
+```
+
+Here is a function to do all this:
+
+```{r}
+sim_data <- function(n, rr, x_type = "binary", baseline = "constant",
+                     start.date = "2000-01-01", ...){
+
+        require(dplyr)
+
+        if(x_type == "binary"){
+                x <- binary_exposure(n = n, ...)
+        } else if(x_type == "continuous"){
+                x <- continuous_exposure(n = n, ...)
+        }
+
+        if(baseline == "constant"){
+                exp_base_y <- constant_baseline(n = n, ...)
+        }
+
+        df <- full_join(x, exp_base_y, by = "date") %>%
+                mutate(exp_y = exp(log(exp_base_y) + log(rr) * x))
+        df$y <- sapply(df$exp_y, FUN = function(x) rpois(1, x))
+
+        return(df)
+
+}
+```
+
+Here are some examples of this function:
+
+```{r}
+sim_data(n = 3, rr = 1.2, p_exp = .2, lambda = 100)
+sim_data(n = 3, rr = 1.02, x_type = "continuous", mu = 10, sd = 10, lambda = 100)
+```
+
+## Fitting the models
+
+### Generalized linear model
+
+The formula for the generalized linear model (GLM) is:
+
+$$
+E(log(\mu_t)) = \beta_0 + \beta_1X_t + \mbox{ns(time, d.f.)}
+$$
+
+where:
+
+- $\mu_t$: Count of outcomes on day $t$, dispersed as a quasi-Poisson distribution
+- $\beta_0$: Intercept (expected log count of health outcome when $X_t$ equals 0 and for middle of the timeframe [so `time` = 0])
+- $\beta_1$: log relative risk for a 1-unit increase in exposure $X_t$
+- $X_t$: Exposure on day $t$
+- $\mbox{ns(time, d.f.)}$: Spline of time for $\mbox{d.f.}$ degrees of freedom per year
+
+```{r}
+# Fit a generalized linear model
+spline_mod <- function(df, df_year = 7){
+        require(splines)
+
+        dgrs_free <- df_year * as.numeric(diff(df[c(1, nrow(df)), "date"])) / 365.4
+        df$time <- scale(df$date, center = TRUE, scale = FALSE)
+        mod <- glm(y ~ x + ns(time, round(dgrs_free)),
+                   data = df,
+                   family = quasipoisson(link = "log"))
+
+        out_1 <- summary(mod)$coef[2, ]
+        out_2 <- confint.default(mod)[2, ]
+        out <- c(out_1, out_2)
+        return(out)
+}
+```
+
+Here are some examples of fitting this model:
+
+```{r}
+df <- sim_data(n = 5 * 365, rr = 1.2, p_exp = .2, lambda = 100)
+spline_mod(df)
+exp(spline_mod(df)[1])
+
+df <- sim_data(n = 5 * 365, rr = 1.02, x_type = "continuous",
+               mu = 10, sd = 10, lambda = 100)
+spline_mod(df)
+exp(spline_mod(df)[1])
+```
+
+### Case-crossover model
+
+Here is the formula for the case-crossover model. It's actually using a quasi-Poisson GLM with a factor included identifying stratum (month and year):
+
+$$
+E(log(\mu_t)) = \beta_0 + \beta_1X_t + \beta_2S_t
+$$
+
+where:
+
+- $\mu_t$: Count of outcomes on day $t$, assuming these outcomes follow a quasi-Poisson distribution
+- $\beta_0$: Intercept (expected log count of health outcome when $X_t$ equals 0 and for the baseline level of the time strata
+- $\beta_1$: log relative risk for a 1-unit increase in exposure $X_t$
+- $X_t$: Exposure on day $t$
+- $\beta_2$: offset for log baseline counts for each stratum 
+- $S_t$: Time stratum (month and year)
+
+```{r}
+# Fit a case-crossover model
+casecross_mod <- function(df){
+        df$stratum <- factor(format(df$date, "%Y.%m"))
+
+        if (sum(df$x == 0 | df$x == 1) == length(df$x)){
+                event.check <- as.matrix(table(df$stratum, df$x))
+                informative.strata <- rownames(event.check)[apply(event.check,
+                                                          1, prod) > 0]
+                df <- subset(df, stratum %in% informative.strata)
+
+                if(length(informative.strata) > 1){
+                        mod <- glm(y ~ x + stratum,
+                                   data = df,
+                                   family = quasipoisson(link = "log"))
+                } else {
+                        mod <- glm(y ~ x,
+                                   data = df,
+                                   family = quasipoisson(link = "log"))
+                        }
+        } else {
+                mod <- glm(y ~ x + stratum,
+                                   data = df,
+                                   family = quasipoisson(link = "log"))
+        }
+
+        out_1 <- summary(mod)$coef[2, ]
+        out_2 <- confint.default(mod)[2, ]
+        out <- c(out_1, out_2)
+        return(out)
+}
+```
+
+Here are some examples of fitting this model:
+
+```{r}
+df <- sim_data(n = 5 * 365, rr = 1.2, p_exp = .2, lambda = 100)
+casecross_mod(df)
+exp(casecross_mod(df)[1])
+
+df <- sim_data(n = 5 * 365, rr = 1.02, x_type = "continuous",
+               mu = 10, sd = 10, lambda = 100)
+casecross_mod(df)
+exp(casecross_mod(df)[1])
+```
+
+### Cross-year model
+
+$$
+E(log(\mu_t)) = \beta_0 + \beta_1X_t + \beta_2D_t + \beta_3Y_t
+$$
+
+```{r}
+# Fit a cross-year model
+# Fit a case-crossover model
+crossyear_mod <- function(df){
+        df$stratum <- factor(format(df$date, "%j"))
+        df$year <- factor(format(df$date, "%Y"))
+
+        if (sum(df$x == 0 | df$x == 1) == length(df$x)){
+                event.check <- as.matrix(table(df$stratum, df$x))
+                informative.strata <- rownames(event.check)[apply(event.check,
+                                                          1, prod) > 0]
+                df <- subset(df, stratum %in% informative.strata)
+
+                if(length(informative.strata) > 1){
+                        mod <- glm(y ~ x + stratum,
+                                   data = df,
+                                   family = quasipoisson(link = "log"))
+                } else {
+                        mod <- glm(y ~ x,
+                                   data = df,
+                                   family = quasipoisson(link = "log"))
+                        }
+        } else {
+                mod <- glm(y ~ x + stratum,
+                                   data = df,
+                                   family = quasipoisson(link = "log"))
+        }
+
+        out_1 <- summary(mod)$coef[2, ]
+        out_2 <- confint.default(mod)[2, ]
+        out <- c(out_1, out_2)
+        return(out)
+}
+```
+
+Here are some examples of applying this model:
+
+```{r}
+df <- sim_data(n = 5 * 365, rr = 1.2, p_exp = .2, lambda = 100)
+crossyear_mod(df)
+exp(crossyear_mod(df)[1])
+
+df <- sim_data(n = 5 * 365, rr = 1.02, x_type = "continuous",
+               mu = 10, sd = 10, lambda = 100)
+crossyear_mod(df)
+exp(crossyear_mod(df)[1])
+```
+
+## Run lots of simulations
+
+Now it's pretty easy to run lots of simulations using `replicate`:
+
+```{r cache = TRUE, fig.width = 10, fig.height = 4}
+n_sim <- 50
+ex_spline <- replicate(n_sim, exp(spline_mod(sim_data(n = 5 * 365,
+                                          rr = 1.02, x_type = "continuous",
+                                          mu = 10, sd = 10, lambda = 100))[1]))
+
+ex_casecross <- replicate(n_sim, exp(casecross_mod(sim_data(n = 5 * 365,
+                                          rr = 1.02, x_type = "continuous",
+                                          mu = 10, sd = 10, lambda = 100))[1]))
+
+ex_crossyear <- replicate(n_sim, exp(crossyear_mod(sim_data(n = 5 * 365,
+                                          rr = 1.02, x_type = "continuous",
+                                          mu = 10, sd = 10, lambda = 100))[1]))
+
+ex <- data.frame(beta_hat = c(ex_spline, ex_casecross, ex_crossyear),
+                 model = rep(c("spline", "casecross", "crossyear"),
+                             each = n_sim)) %>%
+        mutate(model = factor(model))
+
+ggplot(ex, aes(x = beta_hat)) + 
+        geom_histogram(fill = "lightgray", color = "white") + 
+        facet_wrap(~ model, ncol = 3) + 
+        theme_bw()
+
+group_by(ex, model) %>% 
+        summarize(mean_beta = mean(beta_hat), sd_beta = sd(beta_hat))
+```
+
+Here is a function that will run a lot of simulations for a given dataset and model and will output a dataframe with the $\hat{\beta}$, $sd(\hat\beta)$, and p-values for the $\beta$ for each replication:
+
+```{r}
+rep_sims <- function(n_sims, model, n = 5 * 365, rr = 1.01, 
+                     x_type = "continuous", mu = 10, sd = 2, lambda = 100){
+        library(dplyr)
+        out <- replicate(n_sims, eval(call(model, 
+                                           sim_data(n = n, rr = rr, 
+                                                    x_type = x_type,
+                                                    mu = mu, sd = sd,
+                                                    lambda = lambda))))
+        out <- as.data.frame(t(out))
+        names(out) <- c("est", "se", "t", "p", "lower_ci", "upper_ci")
+        return(out)
+}
+```
+
+Here are some examples of using this function:
+
+```{r}
+rep_sims(n_sims = 3, model = "spline_mod")
+rep_sims(n_sims = 3, model = "casecross_mod")
+rep_sims(n_sims = 3, model = "crossyear_mod")
+```
+
+## Assess performance based on simulations
+
+All of these functions will take, as input, a dataframe of replicated simulations:
+
+```{r}
+rep_df <- rep_sims(n_sims = 50, model = "spline_mod", rr = 1.02)
+head(rep_df, 4)
+```
+
+### Average estimate coefficient
+
+This is the mean value of all the $\hat{\beta}$s over $n$ simulations [@figueiras2005analysis].
+I'm also having this function calculate the mean of the $\hat{RR}$ values calculated by all the simulations.
+
+```{r}
+# Measure mean of estimated coefficient
+mean_beta <- function(df){
+        beta_hat <- mean(df$est)
+        rr_hat <- mean(exp(df$est))
+        out <- data.frame(beta_hat, rr_hat)
+        return(out)
+}
+mean_beta(rep_df)
+```
+
+### Standard deviation of estimated coefficients
+
+This is the standard deviation of all the $\hat{\beta}$s over $n$ simulations [@figueiras2005analysis].
+(In other words, I think they took all the $\hat\beta$s and took the standard deviation of all those point estimates.)
+
+Bateson and Schwartz measure the *variance* of the estimated coefficients, instead, and they took the mean value of the variance of each of the estimated coefficients (rather than the variance across all the point estimates) [-@bateson1999control].
+In other words, I think they measured:
+
+$$
+\mbox{Variance} = E(Var(\hat\beta))
+$$
+
+I'll write a function that will estimate both, `var_across_betas` to correspond with Figueiras and `mean_beta_var` to correspond with Bateson and Schwartz:
+
+```{r}
+# Measure standard deviation across estimated coefficient
+beta_var <- function(df){
+        var_across_betas <- var(df$est)
+        mean_beta_var <- mean(df$se^2)
+        out <- data.frame(var_across_betas, mean_beta_var)
+        return(out)
+}
+beta_var(rep_df)
+```
+
+### % bias in estimated coefficient
+
+Based on Figueiras and coauthors, this is measured as [-@figueiras2005analysis]:
+
+$$
+\mbox{% bias} = 100 \frac{(\hat{\beta} - \beta)}{\beta}
+$$
+
+where $\hat{\beta}$ is the mean estimated coefficient (e.g., log relative risk) over $n$ simulations.
+
+Wang and co-authors measure *relative bias* instead [-@wang2011potential]:
+
+$$
+\mbox{% bias} = 100 \frac{(\beta - \hat{\beta})}{\beta}
+$$
+
+I'll write a function that measures both (`fig_bias` and `wang_bias`):
+
+```{r}
+# Measure percent bias
+beta_bias <- function(df, true_rr){
+        fig_bias <- 100 * (mean(df$est) - true_rr) / true_rr
+        wang_bias <- 100 * (true_rr - mean(df$est)) / true_rr
+        out <- data.frame(fig_bias, wang_bias)
+        return(out)
+}
+beta_bias(rep_df, true_rr = 1.02)
+```
+
+As a note, one of these will just be the negative of the other... Does it make more sense to just take the absolute value of the difference between the two estimates and divide it by $\beta$?
+
+### % coverage of estimated coefficients
+
+$$
+\mbox{% coverage} = \mbox{% of simulations with 95% CIs that cover }\beta
+$$
+
+where $\beta$ is the true value of the coefficient (e.g., log relative risk) [@figueiras2005analysis].
+
+From Butland and coauthors [@butland2013measurement]:
+
+> "An estimate of coverage probability records the percentage of simulations where the 95% confidence interval contains the 'true' value of $\beta$."
+
+From Bateson and Schwartz [@bateson1999control]:
+
+> "Coverage probabilities are the percentages of the 1,000 confidence intervals that included the true relative risk of exposure."
+
+These all seem to be saying the same thing.
+
+```{r}
+# Measure coverage
+coverage_beta <- function(df, true_rr){
+        true_beta <- log(true_rr)
+        coverage <- df$lower_ci <= true_beta & df$upper_ci >= true_beta
+        out <- data.frame(coverage = sum(coverage) / nrow(df))
+        return(out)
+}
+
+coverage_beta(rep_df, true_rr = 1.02)
+```
+
+### Power
+
+This is the percent of all estimated coefficients for the $n$ simulations where the lower 95% confidence interval is above 0 [@figueiras2005analysis]. This is a metric of statistical efficiency.
+
+From Butland and coauthors [@butland2013measurement]:
+
+> "An estimate of power records the percentage of simulations that would have detected the health effect estimate as statistically significant at the 5% significance level."
+
+These also seem to be saying the same thing.
+
+```{r}
+# Measure power
+power_beta <- function(df){
+        no_zero <- df$lower_ci >= 0 | df$upper_ci <= 0
+        out <- data.frame(power = sum(no_zero) / nrow(df))
+        return(out)
+}
+
+power_beta(rep_df)
+```
+
+### Relative efficiency
+
+From Bateson and Schwartz [@bateson1999control]:
+
+> "Relative efficiency is the ratio of the $V_{Poi}$ to the $V_{SBI}$."
+
+Where $V_{Poi}$ is the means of 1,000 parameter variances from the regressions of simulated values and $V_{SBI}$ is the means of 1,000 parameter variances from a case-crossover model.
+
+This looks like it is a comparative metric. I have already measured the mean variance of the $\hat{\beta}$s with the `beta_var` function I defined above.
+
+### Combined function to assess performance
+
+I created a function that will perform all these checks:
+
+```{r}
+check_sims <- function(df, true_rr){
+        a <- mean_beta(df)
+        b <- beta_var(df)
+        c <- beta_bias(df, true_rr = true_rr)
+        d <- coverage_beta(df, true_rr = true_rr)
+        e <- power_beta(df)
+
+        out <- cbind(a, b, c, d, e)
+        return(out)
+}
+check_sims(rep_df, true_rr = 1.02)
+```
+
+## Power calculations
+
+Power can vary with the following elements of the analysis:
+
+- Size of the dataset
+- Effect size
+- Baseline mortality rate (here, $\lambda$)
+- Patterns in exposure (variance for continuous exposure, percent of exposed days for binary exposure)
+
+```{r}
+power_calc <- function(varying, values, plot = FALSE, ...){
+        out <- data.frame(x = values, power = NA)
+        if(varying == "n"){
+                for(i in 1:nrow(out)){
+                        rep_df <- rep_sims(n = out$x[i], ...)
+                        out$power[i] <- power_beta(rep_df)[1,1]
+                }
+        } else if(varying == "rr"){
+                for(i in 1:nrow(out)){
+                        rep_df <- rep_sims(rr = out$x[i], ...)
+                        out$power[i] <- power_beta(rep_df)[1,1]
+                }
+        } else if(varying == "lambda"){
+                for(i in 1:nrow(out)){
+                        rep_df <- rep_sims(lambda = out$x[i], ...)
+                        out$power[i] <- power_beta(rep_df)[1,1]
+                }
+        }
+
+        if(plot == TRUE){
+                        library(ggplot2)
+                        my_plot <- ggplot(out, aes(x = x, y = power)) + 
+                                geom_line() + theme_minimal() + 
+                                xlab(varying)
+                        print(my_plot)
+        }
+
+        colnames(out)[1] <- varying
+        return(out)
+}
+```
+
+Here are some examples of varying the sample size, `n`:
+
+```{r fig.height = 3, fig.width = 3, cache = TRUE}
+power_calc(varying = "n", values = c(50 * (1:5)),
+           n_sims = 50, model = "spline_mod", rr = 1.02)
+power_calc(varying = "n", values = c(365 * (1:5)),
+           n_sims = 100, model = "spline_mod", rr = 1.003, plot = TRUE)
+```
+
+Here are some examples of varying the relative risk, `r`:
+
+```{r fig.height = 3, fig.width = 3, cache = TRUE}
+power_calc(varying = "rr", values = c(1.002, 1.005, 1.01, 1.02, 1.03, 1.05, 1.1),
+           n_sims = 50, model = "spline_mod", n = 365 * 5)
+power_calc(varying = "rr", values = c(1.002, 1.005, 1.01, 1.02, 1.03, 1.05, 1.1),
+           n_sims = 100, model = "spline_mod", n = 365, plot = TRUE)
+```
+
+Here are some examples of varying the baseline mortality rate, `lambda`:
+
+```{r fig.height = 3, fig.width = 3, cache = TRUE}
+power_calc(varying = "lambda", values = c(1, 5, 10, 25, 50, 75, 100),
+           n_sims = 50, model = "spline_mod", n = 365 * 5, rr = 1.02)
+power_calc(varying = "lambda", values = c(1, 5, 10, 25, 50, 75, 100),
+           n_sims = 100, model = "spline_mod", n = 365 * 5, rr = 1.005, plot = TRUE)
+```
+
+I can compare power from different models using the same data parameters (i.e., sample size, relative risk, etc.) for these:
+
+```{r, fig.width = 5, fig.height = 3, cache = TRUE}
+for(model in c("spline_mod", "casecross_mod", "crossyear_mod")){
+        df_mod <- power_calc(varying = "n", values = c(365 * (1:5)),
+                             n_sims = 100, model = "spline_mod", rr = 1.005)
+        df_mod$model <- model
+        if(model == "spline_mod"){
+                df <- df_mod
+        } else {
+                df <- rbind(df, df_mod)
+        }
+}
+df$model <- factor(df$model)
+
+ggplot(df, aes(x = n, y = power, color = model)) +
+        geom_line() + 
+        theme_minimal() +
+        geom_hline(aes(yintercept = 0.8), linetype = 2)
+```
+
+## References

### 2016-06-02 DraftOne.Rmd
Simulating continuous exposure with seasonal trends

@@
 }
+
+# A function for continuous exposure which varies by season:
+season_contexp <- function(n, mu, sd, trend){
+  day <- c(1:n)
+  if (trend=="cos1"){
+    mu <- mu*(1+.6*cos(2*pi*(day/365)))
+  }
+  if (trend == "cos2"){
+    mu <- mu*(1+.6*cos(2*pi*(day/365)) + ifelse(day<639 & day>274, .4*cos(2*(pi*(day/365))), 0))
+  }
+  if (trend == "cos3"){
+    mu <- mu*(1+.75^(day/365)*.6*cos(2*pi*(day/365)))
+  }
+  if (trend == "linear"){
+    mu <- mu* (1+ (day/n))
+  }
+  if (trend == "curvilinear"){
+    mu <- mu*(1+day*(2/n)+day^2*(-1/n^2))
+  }
+  if (trend == "cos1linear"){
+    mu <- mu*((1+(day/n)) *(1+.6*cos(2*pi*(day/365))))
+  }
+  if (trend == "cvd"){ #Figure this out later.
+  } #else{
+    #stop("Specify a trend variable from cos1, cos2, cos3, linear, curvilinear, cos1linear, or cvd."
+    #)}
+  contexp <- rnorm(n, mean=mu, sd=sd)
+  return(contexp)
+}
 ```
@@
 #Here, exposure would be the output from the season_binexp function and t would be the output from the calc_t function. 
-sim_outcome <- function(n, mean_out, t, exposure, rr){
+sim_binout <- function(n, mean_out, t, exposure, rr){
   day <- c(1:n)
   rr <- ifelse(exposure == 1, rr, 1)
-  mean_outcome <- mean_out*t*rr
-  sim_outcome <- rpois(n, mean_outcome)
-  return(sim_outcome)
+  exp_base_y <- mean_out*t
+  exp_y <- mean_out*t*rr
+  y <- rpois(n, exp_y)
+  df <- data.frame(exp_base_y=exp_base_y, exp_y=exp_y, y=y)
+  return(df)
+}
+
+#Here, exposure would be the output from the season_contexp function and t is from calc_t.
+sim_contout<- function(n, mu, t, exposure, rr){
+  day <- c(1:n)
+  exp_y <- mu*t*rr*(exposure/sd(exposure))
+  y <- rpois(n, exp_y)
+  df <- data.frame( exp_y=exp_y, y=y)
+  return(df)
 }
@@
 ```{r}
-#Make this function output a dataframe with date, x, etc.
-sim_seasdata <- function(n, rr, trend, x_type="binary", p, mean_out){
+#Put "date" in the data frame like the others? 
+sim_seasdata <- function(n, rr, trend, x_type="binary", p, mu, sd, mean_out){
   if(x_type == "binary"){
     x <- season_binexp(n, p)
+    t <- calc_t(n, trend)
+    simdata <- data.frame(x, sim_binout(n, mean_out, t, x, rr))
   } 
   else if(x_type == "continuous"){
-
+    x <- season_contexp(n, mu, sd, trend)
+    t <- calc_t(n, trend)
+    simdata <- sim_contout(n, mu, t, x, rr)
   }
-  t <- calc_t(n, trend)
-  outcomes <- sim_outcome(n, mean_out, t, x, rr)
-  return(outcomes)
+ return(simdata)
 }

### 2016-06-03 R/sim_exposure.R
Add function for simulating binary exposure

@@ -1,0 +1,19 @@
+#' Simulate binary exposure data
+#'
+#' This function simulated binary exposure data.
+#'
+#' @param n A numeric value giving the number of days to simulate.
+#' @param p A numeric value giving the baseline probability of exposure for
+#'    a binary exposure.
+#'
+#' @return A numeric vector with simulated exposure of length n.
+#'
+#' @examples
+#' season_binexp(5, 0.25)
+#'
+#' @export
+season_binexp <- function(n, p){
+  p <- p #Change this later to reflect probability varying by season using trends
+  binexp <- sample(c(0,1), size = n, replace = T, prob = c(1-p, p))
+  return(binexp)
+}

### 2016-06-03 DraftOne.Rmd
Add function to simulate continuous exposure

@@
 library(splines)
+library(eesim)
 ```
@@
 # A function for binary exposure which varies by season:
-season_binexp <- function(n, p){
-  p <- p #Change this later to reflect probability varying by season using trends
-  binexp <- sample(c(0,1), size = n, replace = T, prob = c(1-p, p))
-  return(binexp)
-}
+season_binexp
 

### 2016-06-03 R/sim_exposure.R

@@ -17,3 +17,48 @@ season_binexp <- function(n, p){
   binexp <- sample(c(0,1), size = n, replace = T, prob = c(1-p, p))
   return(binexp)
 }
+
+#' Simulate continuous exposure data
+#'
+#' This function simulates a time series of continuous exposure data.
+#'
+#' @param mu A numeric vector giving the average of the exposure distribution.
+#' @param sd A numeric vector giving the standard deviation of the exposure
+#'    distribution.
+#' @param trend A character string that gives the trend function to use. Options
+#'    include ...
+#' @inheritParams season_binexp
+#'
+#' @return A numeric vector with simulated exposure of length n.
+#'
+#' @examples
+#' season_contexp(n = 5, mu = 100, sd = 10, trend = "cos1")
+#'
+#' @export
+season_contexp <- function(n, mu, sd, trend){
+  day <- c(1:n)
+  if (trend=="cos1"){
+    mu <- mu*(1+.6*cos(2*pi*(day/365)))
+  }
+  if (trend == "cos2"){
+    mu <- mu*(1+.6*cos(2*pi*(day/365)) + ifelse(day<639 & day>274, .4*cos(2*(pi*(day/365))), 0))
+  }
+  if (trend == "cos3"){
+    mu <- mu*(1+.75^(day/365)*.6*cos(2*pi*(day/365)))
+  }
+  if (trend == "linear"){
+    mu <- mu* (1+ (day/n))
+  }
+  if (trend == "curvilinear"){
+    mu <- mu*(1+day*(2/n)+day^2*(-1/n^2))
+  }
+  if (trend == "cos1linear"){
+    mu <- mu*((1+(day/n)) *(1+.6*cos(2*pi*(day/365))))
+  }
+  if (trend == "cvd"){ #Figure this out later.
+  } #else{
+  #stop("Specify a trend variable from cos1, cos2, cos3, linear, curvilinear, cos1linear, or cvd."
+  #)}
+  contexp <- rnorm(n, mean=mu, sd=sd)
+  return(contexp)
+}

### 2016-06-03 DraftOne.Rmd
Update DraftOne

@@
 # A function for continuous exposure which varies by season:
-season_contexp <- function(n, mu, sd, trend){
-  day <- c(1:n)
-  if (trend=="cos1"){
-    mu <- mu*(1+.6*cos(2*pi*(day/365)))
-  }
-  if (trend == "cos2"){
-    mu <- mu*(1+.6*cos(2*pi*(day/365)) + ifelse(day<639 & day>274, .4*cos(2*(pi*(day/365))), 0))
-  }
-  if (trend == "cos3"){
-    mu <- mu*(1+.75^(day/365)*.6*cos(2*pi*(day/365)))
-  }
-  if (trend == "linear"){
-    mu <- mu* (1+ (day/n))
-  }
-  if (trend == "curvilinear"){
-    mu <- mu*(1+day*(2/n)+day^2*(-1/n^2))
-  }
-  if (trend == "cos1linear"){
-    mu <- mu*((1+(day/n)) *(1+.6*cos(2*pi*(day/365))))
-  }
-  if (trend == "cvd"){ #Figure this out later.
-  } #else{
-    #stop("Specify a trend variable from cos1, cos2, cos3, linear, curvilinear, cos1linear, or cvd."
-    #)}
-  contexp <- rnorm(n, mean=mu, sd=sd)
-  return(contexp)
-}
+season_contexp
 ```

### 2016-06-06 R/sim_exposure.R
Add calc_t function to generate trend variable

@@ -1,6 +1,6 @@
 #' Simulate binary exposure data
 #'
-#' This function simulated binary exposure data.
+#' This function simulates binary exposure data.
 #'
 #' @param n A numeric value giving the number of days to simulate.
 #' @param p A numeric value giving the baseline probability of exposure for
@@ -37,28 +37,8 @@ season_binexp <- function(n, p){
 #' @export
 season_contexp <- function(n, mu, sd, trend){
   day <- c(1:n)
-  if (trend=="cos1"){
-    mu <- mu*(1+.6*cos(2*pi*(day/365)))
-  }
-  if (trend == "cos2"){
-    mu <- mu*(1+.6*cos(2*pi*(day/365)) + ifelse(day<639 & day>274, .4*cos(2*(pi*(day/365))), 0))
-  }
-  if (trend == "cos3"){
-    mu <- mu*(1+.75^(day/365)*.6*cos(2*pi*(day/365)))
-  }
-  if (trend == "linear"){
-    mu <- mu* (1+ (day/n))
-  }
-  if (trend == "curvilinear"){
-    mu <- mu*(1+day*(2/n)+day^2*(-1/n^2))
-  }
-  if (trend == "cos1linear"){
-    mu <- mu*((1+(day/n)) *(1+.6*cos(2*pi*(day/365))))
-  }
-  if (trend == "cvd"){ #Figure this out later.
-  } #else{
-  #stop("Specify a trend variable from cos1, cos2, cos3, linear, curvilinear, cos1linear, or cvd."
-  #)}
+  t <- calc_t(n, trend)
+  mu <- mu*t
   contexp <- rnorm(n, mean=mu, sd=sd)
   return(contexp)
 }

### 2016-06-06 DraftOne.Rmd
Adding calc_t to package

@@
 
-First, we need to generate the explanatory exposure variable.
-Here is a function for binary exposure in which the probability varies by season.
-Later we will add a function for continuous exposure which varies by season (~Normal(mu, sigma)).
+A function to calculate the trend variable. Choices for trend are cvd, cos1, cos2, cos3, linear, curvilinear, cos1linear.
+
+```{r}
+calc_t 
+```
+
+We need to generate the explanatory exposure variable.  Here are functions for binary and continuous exposure which varies by season.
 
@@
 season_contexp
-```
-
-Now a function to calculate the trend variable. Choices for trend are cvd, cos1, cos2, cos3, linear, curvilinear, cos1linear.
-
-```{r}
-calc_t <- function(n, trend){
-  day <- c(1:n)
-  if (trend == "cos1"){
-    seasont <- 1+.6*cos(2*pi*(day/365))
-  }
-  if (trend == "cos2"){
-    seasont <- 1+.6*cos(2*pi*(day/365)) + ifelse(day<639 & day>274, .4*cos(2*(pi*(day/365))), 0)
-  }
-  if (trend == "cos3"){
-    seasont <- 1+.75^(day/365)*.6*cos(2*pi*(day/365))
-  }
-  if (trend == "linear"){
-    seasont <- 1+ (day/n)
-  }
-  if (trend == "curvilinear"){
-    seasont <- 1+day*(2/n)+day^2*(-1/n^2)
-  }
-  if (trend == "cos1linear"){
-    seasont <- (1+(day/n)) *(1+.6*cos(2*pi*(day/365)))
-  }
-  if (trend == "cvd"){ #Figure this out later.
-  } #else{
-    #stop("Specify a trend variable from cos1, cos2, cos3, linear, curvilinear, cos1linear, or cvd."
-    #)}
-  return(seasont)
-}
 ```

### 2016-06-06 R/sim_exposure.R

@@ -1,5 +1,45 @@
 #' Simulate binary exposure data
 #'
+#' This function creates a trend variable.
+#'
+#' @param n A numeric value giving the number of days to simulate.
+#' @param trend A character string that gives the trend function to use.  Options include
+#'    cos1, cos2, cos3, linear, curvilinear, cos1linear, and no trend.
+#'
+#' @return A numeric vector used to generate data with seasonal trends.
+#'
+#' @examples
+#' calc_t(5, "cos3")
+#'
+calc_t <- function(n, trend = "no trend"){
+day <- c(1:n)
+if (trend == "cos1"){
+  seasont <- 1+.6*cos(2*pi*(day/365))
+}
+if (trend == "cos2"){
+  seasont <- 1+.6*cos(2*pi*(day/365)) + ifelse(day<639 & day>274, .4*cos(2*(pi*(day/365))), 0)
+}
+if (trend == "cos3"){
+  seasont <- 1+.75^(day/365)*.6*cos(2*pi*(day/365))
+}
+if (trend == "linear"){
+  seasont <- 1+ (day/n)
+}
+if (trend == "curvilinear"){
+  seasont <- 1+day*(2/n)+day^2*(-1/n^2)
+}
+if (trend == "cos1linear"){
+  seasont <- (1+(day/n)) *(1+.6*cos(2*pi*(day/365)))
+}
+if (trend == "no trend"){
+  seasont <- 1
+}
+else{
+}
+return(seasont)
+}
+
+#'
 #' This function simulates binary exposure data.
 #'
 #' @param n A numeric value giving the number of days to simulate.

### 2016-06-07 DraftOne.Rmd
Adding functions for exposure data without seasonal trends to package

@@
 # Function to generate a binary exposure
-binary_exposure <- function(n, p_exp, start.date = "2000-01-01", ...){
-        start.date <- as.Date(start.date)
-        date <- seq(from = start.date, by = 1, length.out = n)
-        x <- sample(c(0, 1), size = n, replace = TRUE,
-                    prob = c(1-p_exp, p_exp))
-        df <- data.frame(date, x)
-        return(df)
-}
+binary_exposure
+
 binary_exposure(n = 5, p_exp = 0.25)
@@
 # Function to generate a continuous exposure
-continuous_exposure <- function(n, mu, sd, start.date = "2000-01-01", ...){
-        start.date <- as.Date(start.date)
-        date <- seq(from = start.date, by = 1, length.out = n)
-        x <- rnorm(n, mean = mu, sd = sd)
-        df <- data.frame(date, x)
-        return(df)
-}
+continuous_exposure
+
 continuous_exposure(n = 5, mu = 10, sd = 10)

### 2016-06-07 R/sim_exposure.R

@@ -1,4 +1,50 @@
-#' Simulate binary exposure data
+#' Simulate binary exposure data without a seasonal trend
+#'
+#' This function generates binary exposure data without a seasonal trend.
+#'
+#' @inheritParams season_binexp
+#' @param p A numeric value between 0 and 1 specifying the probability of exposure.
+#' @param start.date A date in the format "yyyy-mm-dd" specifying the first day of simulated measurements.
+#'
+#' @return A data frame with the date and a 0 or 1 exposure value for each day.
+#'
+#' @examples
+#' binary_exposure(n = 5, p_exp = 0.25)
+#'
+#' @export
+binary_exposure <- function(n, p, start.date = "2000-01-01", ...){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by = 1, length.out = n)
+  x <- sample(c(0, 1), size = n, replace = TRUE,
+              prob = c(1-p, p))
+  df <- data.frame(date, x)
+  return(df)
+}
+#'
+#' Simulate continuous exposure data without a seasonal trend
+#'
+#' This function simulates continuous exposure data without a seasonal trend.
+#'
+#' @inheritParams season_binexp
+#' @inheritParams binary_exposure
+#' @inheritParams season_contexp
+#'
+#' @return A data frame with date and exposure value for each day.
+#'
+#' @examples
+#' continuous_exposure(n = 5, mu = 10, sd = 10)
+#'
+#' @export
+#'
+continuous_exposure <- function(n, mu, sd, start.date = "2000-01-01", ...){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by = 1, length.out = n)
+  x <- rnorm(n, mean = mu, sd = sd)
+  df <- data.frame(date, x)
+  return(df)
+}
+#'
+#' Create a trend variable
 #'
 #' This function creates a trend variable.
 #'
@@ -40,7 +86,9 @@ return(seasont)
 }
 
 #'
-#' This function simulates binary exposure data.
+#' Simulate binary exposure data with a seasonal trend
+#'
+#' This function simulates binary exposure data with a seasonal trend.
 #'
 #' @param n A numeric value giving the number of days to simulate.
 #' @param p A numeric value giving the baseline probability of exposure for
@@ -58,9 +106,9 @@ season_binexp <- function(n, p){
   return(binexp)
 }
 
-#' Simulate continuous exposure data
+#' Simulate continuous exposure data with a seasonal trend
 #'
[…]

### 2016-06-08 DraftOne.Rmd
Adding outcome data simulation

@@
 binary_exposure 
-
 binary_exposure(n = 5, p_exp = 0.25)
@@
 continuous_exposure 
-
 continuous_exposure(n = 5, mu = 10, sd = 10)
@@
 cvd <- chic$cvd
-
 ```
@@
 
-### Expected mean baseline for outcome data
+### Expected values for outcome data
 
@@
 ```{r}
-sim_data <- function(n, rr, x_type = "binary", baseline = "constant",
+sim_data <- function(n, rr, x_type = "binary", trend = "constant",
                      start.date = "2000-01-01", ...){
@@
         require(dplyr)
-
-        if(x_type == "binary"){
+        t <- calc_t(n, trend)
+        if(x_type == "binary" & trend = "constant"){
                 x <- binary_exposure(n = n, ...)
-        } else if(x_type == "continuous"){
+                simdata <- data.frame(x, sim_binout(n=n, ...))
+        } else if(x_type == "continuous" & trend=="constant"){
                 x <- continuous_exposure(n = n, ...)
+                simdata <- data.frame(x, sim_contout(n=n, ...))
+        } else if(x_type == "binary" & trend != "constant"){
+                x <- season_binexp(n = n, ...)
+                simdata <- data.frame(x, sim_binout(n=n, ...))
+        } else if(x_type == "continuous" & trend != "constant"){
+                x <- season_contexp(n = n, ...)
+                simdata <- data.frame(x, sim_contout(n=n, ...))
         }
 
-        if(baseline == "constant"){
+        if(trend == "constant"){
                 exp_base_y <- constant_baseline(n = n, ...)

### 2016-06-08 R/sim_outcome.R

@@ -1,0 +1,93 @@
+#' Expected baseline outcomes for constant hazard rate
+#'
+#' This is a function to compute the expected baseline outcomes when the hazard rate for exposure is constant.
+#'
+#' @param n A numeric vector specifying the number of days for which to simulate data
+#' @param lambda A numeric value specifying the mean for the expected outcomes
+#' @param start.date A date in the format "yyyy-mm-dd" specifying the first day for which to simulate data
+#'
+#' @return A data frame with the date and expected baseline value of outcomes for each day of simulated data
+#'
+#' @example
+#' constant_baseline(n=3, lambda=100)
+#'
+#' @export
+#'
+constant_baseline <- function(n, lambda, start.date = "2000-01-01", ...){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by = 1, length.out = n)
+  exp_base_y <- rep(lambda, n)
+  df <- data.frame(date = date,
+                   exp_base_y = exp_base_y)
+  return(df)
+}
+constant_baseline(n = 3, lambda = 100)
+#'
+#' Simulate outcome data for binary exposure with a seasonal trend
+#'
+#' This function simulates outcome counts for binary exposure with a seasonal trend.
+#'
+#' @inheritParams constant_baseline
+#' @param mean_out A numeric value specifying the mean outcome count per day
+#' @param t A numeric vector for the trend variable resulting from the calc_t function
+#' @param exposure A numeric vector of exposure values
+#' @param rr A numeric value specifying the relative risk
+#'
+#' @return A data frame with the expected outcomes and simulated outcomes for each day
+#'
+#' @example
+#' sim_binout(n=5, mean_out=22, t=t, exposure = exposure, rr=1.1)
+#'
+sim_binout <- function(n, mean_out, t, exposure, rr){
+  day <- c(1:n)
+  rr <- ifelse(exposure == 1, rr, 1)
+  exp_y <- log(mean_out)*log(t)*log(rr)
+  y <- rpois(n, exp_y)
+  df <- data.frame(exp_y=exp_y, y=y)
+  return(df)
+}
+#'
+#' Simulate outcomes for continuous exposure with a seasonal trend
+#'
+#' This function simulates outcome counts for continuous exposure with a seasonal trend.
+#'
[…]

### 2016-06-10 DraftOne.Rmd
Fixing sim_data function

@@
         require(dplyr)
-        t <- calc_t(n, trend)
-        if(x_type == "binary" & trend = "constant"){
+
+        if(trend=="constant"){
+          if(x_type == "binary"){
+            x <- binary_exposure(n=n, ...)
+          }
+          else if(x_type == "continuous"){
+            x <- continuous_exposure(n=n, ...)
+          }
+          exp_base_y <- constant_baseline(n=n, ...)
+          df <- full_join(x, exp_base_y, by="date") %>% mutate(exp_y = 
+                 exp(log(exp_base_y)+log(rr)*x))
+          df$y <- sapply(df$exp_y, FUN = function(x) rpois(1,x))
+        }
+        else if(trend != "constant"){
+          if(x_type == "binary"){
+            x <- season_binexp(n = n,...)
+          }
+          else if(x_type == "continuous"){
+            x <- season_contexp(n = n, ...)
+          }
+
+        }
+  return(df)
+}
+
+#????
+
+        if(x_type == "binary" & trend == "constant"){
                 x <- binary_exposure(n = n, ...)
-                simdata <- data.frame(x, sim_binout(n=n, ...))
+                exp_base_y <- constant_baseline(n=n, ...)
+                df <- full_join(x, exp_base_y, by="date")
+                df <- mutate(df, exp_y=exp(log(exp_base_y)+log(rr)*x))
+                df$y <- sapply(df$exp_y, FUN = function(x) rpois(1,x))
         } else if(x_type == "continuous" & trend=="constant"){
                 x <- continuous_exposure(n = n, ...)
-                simdata <- data.frame(x, sim_contout(n=n, ...))
+                exp_base_y <- constant_baseline(n=n, ...)
+                df <- full_join(x, exp_base_y, by="date")
+                df <- mutate(df, exp_y=exp(log(exp_base_y)+log(rr)*x))
+                df$y <- sapply(df$exp_y, FUN=function(x) rpois(1,x))
+
         } else if(x_type == "binary" & trend != "constant"){
@@
         }
-
-        if(trend == "constant"){
-                exp_base_y <- constant_baseline(n = n, ...)
-        }
-
-        df <- full_join(x, exp_base_y, by = "date") %>%
-                mutate(exp_y = exp(log(exp_base_y) + log(rr) * x))
-        df$y <- sapply(df$exp_y, FUN = function(x) rpois(1, x))
-
         return(df)

### 2016-06-10 R/sim_outcome.R

@@ -28,7 +28,7 @@ constant_baseline(n = 3, lambda = 100)
 #' This function simulates outcome counts for binary exposure with a seasonal trend.
 #'
 #' @inheritParams constant_baseline
-#' @param mean_out A numeric value specifying the mean outcome count per day
+#' @param lambda A numeric value specifying the mean outcome count per day
 #' @param t A numeric vector for the trend variable resulting from the calc_t function
 #' @param exposure A numeric vector of exposure values
 #' @param rr A numeric value specifying the relative risk
@@ -38,10 +38,10 @@ constant_baseline(n = 3, lambda = 100)
 #' @example
 #' sim_binout(n=5, mean_out=22, t=t, exposure = exposure, rr=1.1)
 #'
-sim_binout <- function(n, mean_out, t, exposure, rr){
+sim_binout <- function(n, lambda, t, exposure, rr){
   day <- c(1:n)
   rr <- ifelse(exposure == 1, rr, 1)
-  exp_y <- log(mean_out)*log(t)*log(rr)
+  exp_y <- log(lambda)*log(t)*log(rr)
   y <- rpois(n, exp_y)
   df <- data.frame(exp_y=exp_y, y=y)
   return(df)

### 2016-06-10 DraftOne.Rmd
Creating seasonal_outcome function

@@
 ```{r}
+#Seasonal baseline
+seasonal_baseline <- function(n, lambda, trend, start.date="2000-01-01", ...){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by=1, length.out=n)
+  t <- calc_t(n=n, ...)
+  exp_base_y <- lambda*t
+  df <- data.frame(date=date, exp_base_y=exp_base_y)
+  return(df)
+}
+```
+
+```{r}
 #Make a sweet plot of these with n=1500 (just for fun, not working yet):
@@
         else if(trend != "constant"){
+          t <- calc_t(n=n, ...)
           if(x_type == "binary"){
@@
           }
-
+          exp_base_y <- 
         }
@@
 #????
-
-        if(x_type == "binary" & trend == "constant"){
-                x <- binary_exposure(n = n, ...)
-                exp_base_y <- constant_baseline(n=n, ...)
-                df <- full_join(x, exp_base_y, by="date")
-                df <- mutate(df, exp_y=exp(log(exp_base_y)+log(rr)*x))
-                df$y <- sapply(df$exp_y, FUN = function(x) rpois(1,x))
-        } else if(x_type == "continuous" & trend=="constant"){
-                x <- continuous_exposure(n = n, ...)
-                exp_base_y <- constant_baseline(n=n, ...)
-                df <- full_join(x, exp_base_y, by="date")
-                df <- mutate(df, exp_y=exp(log(exp_base_y)+log(rr)*x))
-                df$y <- sapply(df$exp_y, FUN=function(x) rpois(1,x))
 
@@
 
-}
 ```

### 2016-06-10 R/sim_outcome.R

@@ -21,7 +21,27 @@ constant_baseline <- function(n, lambda, start.date = "2000-01-01", ...){
                    exp_base_y = exp_base_y)
   return(df)
 }
-constant_baseline(n = 3, lambda = 100)
+#'
+#' Expected baseline outcomes for seasonal hazard rate
+#'
+#' This function generates expected baseline outcomes when the hazard rate for exposure is seasonal.
+#'
+#' @inheritParams constant_baseline
+#' @param trend A character string specifying the desired seasonal trend
+#'
+#' @return A data frame with the date and expected baseline outcomes for each day of simulated data
+#'
+#' @example
+#' seasonal_baseline(n=3, lambda=100, trend="cos1")
+#'
+seasonal_baseline <- function(n, lambda, trend, start.date="2000-01-01", ...){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by=1, length.out=n)
+  t <- calc_t(n=n, ...)
+  exp_base_y <- lambda*t
+  df <- data.frame(date=date, exp_base_y=exp_base_y)
+  return(df)
+}
 #'
 #' Simulate outcome data for binary exposure with a seasonal trend
 #'
@@ -41,7 +61,8 @@ constant_baseline(n = 3, lambda = 100)
 sim_binout <- function(n, lambda, t, exposure, rr){
   day <- c(1:n)
   rr <- ifelse(exposure == 1, rr, 1)
-  exp_y <- log(lambda)*log(t)*log(rr)
+  exp_base_y <-
+  exp_y <- log(lambda*t)+log(rr)*exposure
   y <- rpois(n, exp_y)
   df <- data.frame(exp_y=exp_y, y=y)
   return(df)

### 2016-06-13 R/sim_outcome.R
Update documentation

@@ -32,7 +32,7 @@ constant_baseline <- function(n, lambda, start.date = "2000-01-01", ...){
 #' exposure is seasonal.
 #'
 #' @inheritParams constant_baseline
-#' @param trend A character string specifying the desired seasonal trend
+#' @inheritParams calc_t
 #'
 #' @return A data frame with the date and expected baseline outcomes for each
 #'    day of simulated data
@@ -40,10 +40,11 @@ constant_baseline <- function(n, lambda, start.date = "2000-01-01", ...){
 #' @examples
 #' seasonal_baseline(n = 3, lambda = 100, trend = "cos1")
 #'
-seasonal_baseline <- function(n, lambda, trend, start.date = "2000-01-01", ...){
+seasonal_baseline <- function(n, lambda, start.date = "2000-01-01",
+                              trend = "no trend"){
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
-  t <- calc_t(n = n, ...)
+  t <- calc_t(n = n, trend = trend)
   exp_base_y <- lambda * t
   df <- data.frame(date = date, exp_base_y = exp_base_y)
   return(df)

### 2016-06-13 DraftOne.Rmd
Get everything in DraftOne.Rmd up to `sim_data` running

@@
 binary_exposure 
-binary_exposure(n = 5, p_exp = 0.25)
+binary_exposure(n = 5, p = 0.25)
 ```
@@
 n <- 1000
-p_exp <- 0.25
+p <- 0.25
 lambda <- 100
 
-x <- binary_exposure(n = n, p_exp = p_exp)
+x <- binary_exposure(n = n, p = p)
 exp_base_y <- constant_baseline(n = n, lambda = lambda)

### 2016-06-13 R/sim_exposure.R

@@ -68,6 +68,7 @@ continuous_exposure <- function(n, mu, sd, start.date = "2000-01-01", ...){
 #' @examples
 #' calc_t(5, "cos3")
 #'
+#' @export
 calc_t <- function(n, trend = "no trend"){
   day <- c(1:n)
   if (trend == "cos1"){

### 2016-06-13 DraftOne.Rmd
Finishing sim_data function

@@
 ```{r}
-sim_data <- function(n, rr, x_type = "binary", trend = "constant",
+sim_data <- function(n, rr, lambda, x_type = "binary", trend = "constant",
                      start.date = "2000-01-01", ...){
@@
           }
-          exp_base_y <- constant_baseline(n=n, ...)
+          exp_base_y <- constant_baseline(n=n, lambda = lambda, start.date = start.date)
           df <- full_join(x, exp_base_y, by="date") %>% mutate(exp_y = 
@@
         }
+
         else if(trend != "constant"){
-          t <- calc_t(n=n, ...)
           if(x_type == "binary"){
-            x <- season_binexp(n = n,...)
+            x <- season_binexp(n = n, ...)
           }
           else if(x_type == "continuous"){
-            x <- season_contexp(n = n, ...)
+            x <- season_contexp(n = n, trend = trend, ...)
           }
-          exp_base_y <- 
+          exp_base_y <- seasonal_baseline(n = n, lambda = lambda, start.date = start.date,
+                                          trend = trend)
+          df <- full_join(x, exp_base_y, by="date") %>% mutate(exp_y = 
+                exp(log(exp_base_y)+log(rr)*x))
+          df$y <- sapply(df$exp_y, FUN = function(x) rpois(1,x))
         }
@@
 }
-
-#????
-
-        } else if(x_type == "binary" & trend != "constant"){
-                x <- season_binexp(n = n, ...)
-                simdata <- data.frame(x, sim_binout(n=n, ...))
-        } else if(x_type == "continuous" & trend != "constant"){
-                x <- season_contexp(n = n, ...)
-                simdata <- data.frame(x, sim_contout(n=n, ...))
-        }
-        return(df)
 

### 2016-06-13 R/sim_exposure.R

@@ -107,10 +107,13 @@ calc_t <- function(n, trend = "no trend"){
 #' season_binexp(n = 5, p = 0.25)
 #'
 #' @export
-season_binexp <- function(n, p){
+season_binexp <- function(n, p, start.date = "2000-01-01", ...){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by = 1, length.out = n)
   p <- p #Change this later to reflect probability varying by season using trends
-  binexp <- sample(c(0, 1), size = n, replace = T, prob = c(1 - p, p))
-  return(binexp)
+  x <- sample(c(0, 1), size = n, replace = T, prob = c(1 - p, p))
+  df <- data.frame(date, x)
+  return(df)
 }
 
 #' Simulate continuous exposure data with a seasonal trend
@@ -130,10 +133,13 @@ season_binexp <- function(n, p){
 #' season_contexp(n = 5, mu = 100, sd = 10, trend = "cos1")
 #'
 #' @export
-season_contexp <- function(n, mu, sd, trend){
+season_contexp <- function(n, mu, sd, trend, start.date = "2000-01-01", ...){
   day <- c(1:n)
   t <- calc_t(n, trend)
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by = 1, length.out = n)
   mu <- mu * t
-  contexp <- rnorm(n, mean = mu, sd = sd)
-  return(contexp)
+  x <- rnorm(n, mean = mu, sd = sd)
+  df <- data.frame(date, x)
+  return(df)
 }

### 2016-06-13 R/sim_outcome.R

@@ -49,51 +49,65 @@ seasonal_baseline <- function(n, lambda, start.date = "2000-01-01",
   df <- data.frame(date = date, exp_base_y = exp_base_y)
   return(df)
 }
-
-#' Simulate outcome data for binary exposure with a seasonal trend
 #'
-#' This function simulates outcome counts for binary exposure with a seasonal
-#' trend.
+#' Simulated Exposure and Outcome Data
+#'
+#' This function simulates data for exposure levels and health outcomes by date.
 #'
 #' @inheritParams constant_baseline
-#' @param lambda A numeric value specifying the mean outcome count per day
-#' @param t A numeric vector for the trend variable resulting from the
-#'    \code{calc_t} function
-#' @param exposure A numeric vector of exposure values
-#' @param rr A numeric value specifying the relative risk
+#' @inheritParams calc_t
+#' @inheritParams season_binexp
+#' @inheritParams season_contexp
+#' @param rr A numeric value specifying the relative risk of the outcome for each 1-unit increase in exposure
+#' @param x_type A character string specifying the type of exposure data.  Options are "binary" and "continuous".
 #'
-#' @return A data frame with the expected outcomes and simulated outcomes for
-#'    each day
+#' @return A data frame with the date, exposure value, expected baseline outcome,
+#' expected outcome, and simulated outcome for each day of simulated data.
 #'
-#' @examples
-#' exposure <- season_contexp(n = 5, mu = 100, sd = 10, trend = "cos1")
-#' t <- calc_t(n = 5, trend = "no trend")
-#' sim_binout(n = 5, lambda = 22, t = t, exposure = exposure, rr = 1.1)
+#' @example
+#' sim_data(n=10, rr = 1.1, lambda = 100, p = .5)
 #'
-sim_binout <- function(n, lambda, t, exposure, rr){
-  day <- c(1:n)
-  rr <- ifelse(exposure == 1, rr, 1)
-  exp_y <- log(lambda * t) + log(rr) * exposure
-  y <- rpois(n, exp_y)
-  df <- data.frame(exp_y = exp_y, y = y)
+#' @export
+#'
+sim_data <- function(n, rr, lambda, x_type = "binary", trend = "constant",
+                     start.date = "2000-01-01", ...){
+
+  require(dplyr)
+
+  if(trend=="constant"){
+    if(x_type == "binary"){
+      x <- binary_exposure(n=n, ...)
+    }
[…]

### 2016-06-13 DraftOne.Rmd
Adding functions to assess model performance

@@
 ```{r}
-sim_data <- function(n, rr, lambda, x_type = "binary", trend = "constant",
-                     start.date = "2000-01-01", ...){
-
-        require(dplyr)
-
-        if(trend=="constant"){
-          if(x_type == "binary"){
-            x <- binary_exposure(n=n, ...)
-          }
-          else if(x_type == "continuous"){
-            x <- continuous_exposure(n=n, ...)
-          }
-          exp_base_y <- constant_baseline(n=n, lambda = lambda, start.date = start.date)
-          df <- full_join(x, exp_base_y, by="date") %>% mutate(exp_y = 
-                 exp(log(exp_base_y)+log(rr)*x))
-          df$y <- sapply(df$exp_y, FUN = function(x) rpois(1,x))
-        }
-
-        else if(trend != "constant"){
-          if(x_type == "binary"){
-            x <- season_binexp(n = n, ...)
-          }
-          else if(x_type == "continuous"){
-            x <- season_contexp(n = n, trend = trend, ...)
-          }
-          exp_base_y <- seasonal_baseline(n = n, lambda = lambda, start.date = start.date,
-                                          trend = trend)
-          df <- full_join(x, exp_base_y, by="date") %>% mutate(exp_y = 
-                exp(log(exp_base_y)+log(rr)*x))
-          df$y <- sapply(df$exp_y, FUN = function(x) rpois(1,x))
-        }
-  return(df)
-}
-
+sim_data 
 ```
@@
 # Fit a generalized linear model
-spline_mod <- function(df, df_year = 7){
-        require(splines)
-
-        dgrs_free <- df_year * as.numeric(diff(df[c(1, nrow(df)), "date"])) / 365.4
-        df$time <- scale(df$date, center = TRUE, scale = FALSE)
-        mod <- glm(y ~ x + ns(time, round(dgrs_free)),
-                   data = df,
-                   family = quasipoisson(link = "log"))
-
-        out_1 <- summary(mod)$coef[2, ]
-        out_2 <- confint.default(mod)[2, ]
-        out <- c(out_1, out_2)
-        return(out)
-}
+spline_mod 
 ```
@@
 # Fit a case-crossover model
-casecross_mod <- function(df){
-        df$stratum <- factor(format(df$date, "%Y.%m"))
-
-        if (sum(df$x == 0 | df$x == 1) == length(df$x)){
-                event.check <- as.matrix(table(df$stratum, df$x))
-                informative.strata <- rownames(event.check)[apply(event.check,
-                                                          1, prod) > 0]
-                df <- subset(df, stratum %in% informative.strata)
-
-                if(length(informative.strata) > 1){
-                        mod <- glm(y ~ x + stratum,
-                                   data = df,
-                                   family = quasipoisson(link = "log"))
-                } else {
-                        mod <- glm(y ~ x,
-                                   data = df,
-                                   family = quasipoisson(link = "log"))
-                        }
-        } else {
-                mod <- glm(y ~ x + stratum,
-                                   data = df,
-                                   family = quasipoisson(link = "log"))
-        }
-
-        out_1 <- summary(mod)$coef[2, ]
-        out_2 <- confint.default(mod)[2, ]
-        out <- c(out_1, out_2)
-        return(out)
-}
+casecross_mod 
 ```
@@
 # Fit a cross-year model
-# Fit a case-crossover model
-crossyear_mod <- function(df){
-        df$stratum <- factor(format(df$date, "%j"))
-        df$year <- factor(format(df$date, "%Y"))
-
-        if (sum(df$x == 0 | df$x == 1) == length(df$x)){
-                event.check <- as.matrix(table(df$stratum, df$x))
-                informative.strata <- rownames(event.check)[apply(event.check,
-                                                          1, prod) > 0]
-                df <- subset(df, stratum %in% informative.strata)
-
-                if(length(informative.strata) > 1){
-                        mod <- glm(y ~ x + stratum,
-                                   data = df,
-                                   family = quasipoisson(link = "log"))
-                } else {
-                        mod <- glm(y ~ x,
-                                   data = df,
-                                   family = quasipoisson(link = "log"))
-                        }
-        } else {
-                mod <- glm(y ~ x + stratum,
-                                   data = df,
-                                   family = quasipoisson(link = "log"))
-        }
-
-        out_1 <- summary(mod)$coef[2, ]
-        out_2 <- confint.default(mod)[2, ]
-        out <- c(out_1, out_2)
-        return(out)
-}
+crossyear_mod 
 ```
@@
 ```{r}
-rep_sims <- function(n_sims, model, n = 5 * 365, rr = 1.01, 
-                     x_type = "continuous", mu = 10, sd = 2, lambda = 100){
-        library(dplyr)
-        out <- replicate(n_sims, eval(call(model, 
-                                           sim_data(n = n, rr = rr, 
-                                                    x_type = x_type,
-                                                    mu = mu, sd = sd,
-                                                    lambda = lambda))))
-        out <- as.data.frame(t(out))
-        names(out) <- c("est", "se", "t", "p", "lower_ci", "upper_ci")
-        return(out)
-}
+rep_sims 
 ```
@@
 # Measure mean of estimated coefficient
-mean_beta <- function(df){
-        beta_hat <- mean(df$est)
-        rr_hat <- mean(exp(df$est))
-        out <- data.frame(beta_hat, rr_hat)
-        return(out)
-}
+mean_beta 
 mean_beta(rep_df)
@@
 # Measure standard deviation across estimated coefficient
-beta_var <- function(df){
-        var_across_betas <- var(df$est)
-        mean_beta_var <- mean(df$se^2)
-        out <- data.frame(var_across_betas, mean_beta_var)
-        return(out)
-}
+beta_var 
 beta_var(rep_df)

### 2016-06-13 R/assess_performance.R

@@ -1,0 +1,60 @@
+#' Average Estimated Coefficient
+#'
+#' This function gives the mean value of the $\hat{\beta}$s and the mean estimated
+#' relative risk over the n simulations.
+#'
+#' @param df A data frame of replicated simulations which must include a column titled "est"
+#'
+#' @return A data frame with the mean estimated coefficient and mean estimated relative risk
+#'
+#' @example
+#' mean_beta(df)
+#'
+#' @export
+#'
+mean_beta <- function(df){
+  beta_hat <- mean(df$est)
+  rr_hat <- mean(exp(df$est))
+  out <- data.frame(beta_hat, rr_hat)
+  return(out)
+}
+#'
+#' Standard Deviation of Estimated Coefficients
+#'
+#' This function gives the variance of the point estimates of $\hat{\beta}$ over the n simulations
+#' and the mean of the variances of each $\hat{\beta}$.
+#'
+#' @param df A data frame of replicated simulations which must include columns titles "est" and "se".
+#'
+#' @return A data frame of the variance across $\hat{\beta}$s and the mean variance of the $\hat{\beta}$s
+#'
+#' @example
+#' beta_var(df)
+#'
+#' @export
+#'
+beta_var <- function(df){
+  var_across_betas <- var(df$est)
+  mean_beta_var <- mean(df$se^2)
+  out <- data.frame(var_across_betas, mean_beta_var)
+  return(out)
+}
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+#'

### 2016-06-17 DraftOne.Rmd
Get Rmds running on my computer

@@
 ```{r}
-sim_data(n = 3, rr = 1.2, p_exp = .2, lambda = 100)
-sim_data(n = 3, rr = 1.02, x_type = "continuous", mu = 10, sd = 10, lambda = 100)
+sim_data(n = 3, rr = 1.2, p = .2, lambda = 100)
+sim_data(n = 3, rr = 1.02, x_type = "continuous", mu = 10, sd = 10,
+         lambda = 100)
 ```
@@
 ```{r}
-df <- sim_data(n = 5 * 365, rr = 1.2, p_exp = .2, lambda = 100)
+df <- sim_data(n = 5 * 365, rr = 1.2, p = .2, lambda = 100)
 spline_mod(df)
@@
 ```{r}
-df <- sim_data(n = 5 * 365, rr = 1.2, p_exp = .2, lambda = 100)
+df <- sim_data(n = 5 * 365, rr = 1.2, p = .2, lambda = 100)
 casecross_mod(df)
@@
 ```{r}
-df <- sim_data(n = 5 * 365, rr = 1.2, p_exp = .2, lambda = 100)
+df <- sim_data(n = 5 * 365, rr = 1.2, p = .2, lambda = 100)
 crossyear_mod(df)

### 2016-06-17 R/assess_performance.R

@@ -7,7 +7,7 @@
 #'
 #' @return A data frame with the mean estimated coefficient and mean estimated relative risk
 #'
-#' @example
+#' @examples
 #' mean_beta(df)
 #'
 #' @export
@@ -28,7 +28,7 @@ mean_beta <- function(df){
 #'
 #' @return A data frame of the variance across $\hat{\beta}$s and the mean variance of the $\hat{\beta}$s
 #'
-#' @example
+#' @examples
 #' beta_var(df)
 #'
 #' @export

### 2016-06-17 DraftOne.Rmd
Add figure with time trends

@@
 
-```{r}
-#Make a sweet plot of these with n=1500 (just for fun, not working yet):
-trendname <- c(rep("cos1", 1500), rep("cos2",1500), rep("cos3",1500), rep("linear",1500),  rep("curvilinear",1500), rep("cos1linear",1500))
-trendtests <- c(calc_t(1500, trend = "cos1"), calc_t(1500, trend = "cos2"), calc_t(1500, trend = "cos3"), calc_t(1500, trend = "linear"), calc_t(1500, trend = "curvilinear"), calc_t(1500, trend = "cos1linear"))
-trendframe <- data.frame(trendname, trendtests)
+Here's a plot of the trends:
+
+```{r fig.width = 7, fig.height = 4}
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+
+data.frame(day = 1:1000) %>%
+  mutate(none = calc_t(n = 1000, trend = "no trend"),
+         cos1 = calc_t(n = 1000, trend = "cos1"),
+         cos2 = calc_t(n = 1000, trend = "cos2"),
+         cos3 = calc_t(n = 1000, trend = "cos3"),
+         linear = calc_t(n = 1000, trend = "linear"),
+         curvilinear = calc_t(n = 1000, trend = "curvilinear"),
+         cos1linear = calc_t(n = 1000, trend = "cos1linear")) %>%
+  gather(trend_method, trend_value, -day) %>%
+  ggplot(aes(x = day, y = trend_value)) + 
+  geom_line() + facet_wrap(~ trend_method, ncol = 4) + 
+  theme_bw()
 ```

### 2016-06-20 DraftOne.Rmd
Fixing beta_bias, adding coverage_beta

@@
 date: "June 1, 2016"
-output: html_document
+output: word_document
 ---
@@
 beta_bias <- function(df, true_rr){
-        fig_bias <- 100 * (mean(df$est) - true_rr) / true_rr
-        wang_bias <- 100 * (true_rr - mean(df$est)) / true_rr
-        out <- data.frame(fig_bias, wang_bias)
+        percent_bias <- 100 * (true_rr - mean(exp(df$est))) / true_rr
+        out <- data.frame(percent_bias)
         return(out)

### 2016-06-20 R/assess_performance.R

@@ -40,13 +40,42 @@ beta_var <- function(df){
   return(out)
 }
 #'
+#' Percent Bias of Estimated Coefficient
+#'
+#' This function returns the relative bias of the mean of the estimated coefficients.
+#'
+#' @inheritParams mean_beta
+#' @param true_rr The true relative risk used to simulate your data
+#'
+#' @return The percent bias of the mean of the estimated coefficients over n simulations
+#'
+#' @examples
+#' beta_bias(df, true_rr = 1.02)
+#'
+beta_bias <- function(df, true_rr){
+  percent_bias <- 100 * (true_rr - mean(exp(df$est))) / true_rr
+  out <- data.frame(percent_bias)
+  return(out)
+}
+#'
 #' Percent Coverage of Estimated Coefficients
 #'
-#' This function gives the percent coverage for the true coefficient.
+#' This function gives the percent coverage of the true coefficient.
 #'
-#' @param df A data frame of replicated simulations which must include columns titled
+#' @param df A data frame of replicated simulations which must include columns titled "lower_ci" and "upper_ci"
+#' @param true_rr The true relative risk used to simulate your data
 #'
+#' @return The percent of confidence intervals for the estimated relative risk over n simulations which include the true relative risk
 #'
+#' @examples
+#' coverage_beta(df, true_rr = 1.02)
+#'
+coverage_beta <- function(df, true_rr){
+  true_beta <- log(true_rr)
+  coverage <- df$lower_ci <= true_beta & df$upper_ci >= true_beta
+  out <- data.frame(coverage = sum(coverage) / nrow(df))
+  return(out)
+}
 #'
 #'
 #'

### 2016-06-20 DraftOne.Rmd
Cleaning up document

@@
 # Measure percent bias
-beta_bias <- function(df, true_rr){
-        percent_bias <- 100 * (true_rr - mean(exp(df$est))) / true_rr
-        out <- data.frame(percent_bias)
-        return(out)
-}
+beta_bias 
 beta_bias(rep_df, true_rr = 1.02)
@@
 # Measure coverage
-coverage_beta <- function(df, true_rr){
-        true_beta <- log(true_rr)
-        coverage <- df$lower_ci <= true_beta & df$upper_ci >= true_beta
-        out <- data.frame(coverage = sum(coverage) / nrow(df))
-        return(out)
-}
-
+coverage_beta 
 coverage_beta(rep_df, true_rr = 1.02)

### 2016-06-22 DraftOne.Rmd
Finishing script files

@@
 
-This is the percent of all estimated coefficients for the $n$ simulations where the lower 95% confidence interval is above 0 [@figueiras2005analysis]. This is a metric of statistical efficiency.
+This is the percent of all estimated coefficients for the $n$ simulations where the lower 95% confidence interval bound is above 0 [@figueiras2005analysis]. This is a metric of statistical efficiency.
 
@@
 # Measure power
-power_beta <- function(df){
-        no_zero <- df$lower_ci >= 0 | df$upper_ci <= 0
-        out <- data.frame(power = sum(no_zero) / nrow(df))
-        return(out)
-}
+power_beta
 
@@
 ```{r}
-check_sims <- function(df, true_rr){
-        a <- mean_beta(df)
-        b <- beta_var(df)
-        c <- beta_bias(df, true_rr = true_rr)
-        d <- coverage_beta(df, true_rr = true_rr)
-        e <- power_beta(df)
-
-        out <- cbind(a, b, c, d, e)
-        return(out)
-}
+check_sims 
 check_sims(rep_df, true_rr = 1.02)
@@
 ```{r}
-power_calc <- function(varying, values, plot = FALSE, ...){
-        out <- data.frame(x = values, power = NA)
-        if(varying == "n"){
-                for(i in 1:nrow(out)){
-                        rep_df <- rep_sims(n = out$x[i], ...)
-                        out$power[i] <- power_beta(rep_df)[1,1]
-                }
-        } else if(varying == "rr"){
-                for(i in 1:nrow(out)){
-                        rep_df <- rep_sims(rr = out$x[i], ...)
-                        out$power[i] <- power_beta(rep_df)[1,1]
-                }
-        } else if(varying == "lambda"){
-                for(i in 1:nrow(out)){
-                        rep_df <- rep_sims(lambda = out$x[i], ...)
-                        out$power[i] <- power_beta(rep_df)[1,1]
-                }
-        }
-
-        if(plot == TRUE){
-                        library(ggplot2)
-                        my_plot <- ggplot(out, aes(x = x, y = power)) + 
-                                geom_line() + theme_minimal() + 
-                                xlab(varying)
-                        print(my_plot)
-        }
-
-        colnames(out)[1] <- varying
-        return(out)
-}
+power_calc 
 ```

### 2016-06-22 R/assess_performance.R

@@ -52,6 +52,8 @@ beta_var <- function(df){
 #' @examples
 #' beta_bias(df, true_rr = 1.02)
 #'
+#' @export
+#'
 beta_bias <- function(df, true_rr){
   percent_bias <- 100 * (true_rr - mean(exp(df$est))) / true_rr
   out <- data.frame(percent_bias)
@@ -70,6 +72,8 @@ beta_bias <- function(df, true_rr){
 #' @examples
 #' coverage_beta(df, true_rr = 1.02)
 #'
+#' @export
+#'
 coverage_beta <- function(df, true_rr){
   true_beta <- log(true_rr)
   coverage <- df$lower_ci <= true_beta & df$upper_ci >= true_beta
@@ -77,11 +81,99 @@ coverage_beta <- function(df, true_rr){
   return(out)
 }
 #'
+#' Power
 #'
+#' This function gives the power of the test at a 5% siginificance level.
 #'
+#' @inheritParams coverage_beta
 #'
+#' @return Power at 5% significance level
 #'
+#' @examples
+#' power_beta(df)
 #'
+#' @export
+#'
+power_beta <- function(df){
+  no_zero <- df$lower_ci >= 0 | df$upper_ci <= 0
+  out <- data.frame(power = sum(no_zero) / nrow(df))
+  return(out)
+}
+#'
+#' Model Performance Assessment
+#'
+#' This function gives several measures of model performance.
+#'
+#' @inheritParams beta_bias
+#'
+#' @return Mean beta estimate, mean relative risk estimate, variance across betas, mean variance of
+#' the estimates, percent bias, coverage, and power.
+#'
+#' @examples
+#' check_sims(df, true_rr = 1.02)
+#'
+#' @export
+#'
+check_sims <- function(df, true_rr){
+  a <- mean_beta(df)
+  b <- beta_var(df)
+  c <- beta_bias(df, true_rr = true_rr)
+  d <- coverage_beta(df, true_rr = true_rr)
+  e <- power_beta(df)
+
+  out <- cbind(a, b, c, d, e)
+  return(out)
+}
+check_sims(rep_df, true_rr = 1.02)
+#'
+#' Power Calculations
+#'
+#' This function gives the power for a model with varying parameters.
+#'
+#' @param varying A character string of the parameter to be varied.  Choices are "n", "rr", or "lambda".
[…]

### 2016-06-23 DraftOne.Rmd

@@
 ```{r}
-constant_baseline <- function(n, lambda, start.date = "2000-01-01", ...){
-        start.date <- as.Date(start.date)
-        date <- seq(from = start.date, by = 1, length.out = n)
-        exp_base_y <- rep(lambda, n)
-        df <- data.frame(date = date,
-                         exp_base_y = exp_base_y)
-        return(df)
-}
+constant_baseline 
 constant_baseline(n = 3, lambda = 100)
@@
 calc_t
+calc_t(n=10, trend="curvilinear")
 ```
@@
 season_binexp
+season_binexp(n=100, p=.25)
 
@@
 season_contexp
+season_contexp(10, mu = 100, sd = 10, trend = "curvilinear")
 ```

### 2016-06-23 writing/JSSDraft/JSSDraft.Rmd
Add template for JSS draft

---
author:
  - name: Sarah Koehler
    affiliation: Colorado State University
    address: >
      Lake Street
      Fort Collins, CO
    email: sakoehler7@gmail.com
  - name: G. Brooke Anderson
    affiliation: Colorado State University
    address: >
      Lake Street
      Fort Collins, CO
    email: brooke.anderson@colostate.edu
title:
  formatted: "\\pkg{eesim}: A Package to Simulate Environmental Epidemiology Time Series for Simulation Studies and Power Analyses"
  # If you use tex in the formatted title, also supply version without
  plain:     "eesim: A Package to Simulate Environmental Epidemiology Time Series for Simulation Studies and Power Analyses"
  # For running headers, if needed
  short:     "\\pkg{eesim}: Simulations for Environmental Epidemiology"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [environmental epidemiology, simulation studies, power analysis, air pollution]
  plain:     [environmental epidemiology, simulation studies, power analysis, air pollution]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

# Purpose of package

## Environmental time series studies and why they're important

## Using simulated time series for power analysis

## Using simulated time series to test modeling questions

## Challenges in creating simulated time series

* Not consistent across different studies
* Time consuming to create, often only need to customize certain parts

# Package functionality

## Simulating time series

## Using simulated time series for power analysis

## Using simulated time series to test modeling questions

## Evaluating simulations based on these time series

## Customizing simulations

### 2016-06-23 writing/SimulationPaper/SimulationPaper.Rmd
Add template for simulation paper draft

---
title: Performance of time series, case-crossover, and cross-year models in estimating the effects of natural disasters on human health in the presence of delayed or displaced effects 
author:
  - name: Sarah Koehler
    email: sakoehler7@gmail.com
    affiliation: Colorado State University Statistics
  - name: G. Brooke Anderson
    email: brooke.anderson@colostate.edu
    affiliation: Colorado State University ERHS
    footnote: Corresponding Author
address:
  - code: Colorado State University Statistics
    address: Colorado State University, Department of Statistics, Street, Fort Collins, CO, Zip
  - code: Colorado State University ERHS
    address: Colorado State University, Department of Environmental & Radiological Health Sciences, Lake Street, Fort Collins, CO, Zip
abstract: |
  Time series and case-crossover studies have been extensively used to determine the health effects of both continuous environmental exposures (e.g., air pollution concentration, temperature) and binary exposures (e.g., heat waves). However, exposures to severe natural disasters, like heat waves and tropical storms, could affect mid-term patterns in health risks through either delayed effects or through mortality displacement.

  Here, we conduct a simulation study to determine the performance of time series, case-crossover, and cross-year models to assess the acute health effects of natural disasters under different scenarios of delayed health effects or mortality displacement.

bibliography: mybibfile.bib
output: rticles::elsevier_article
---

Introduction
==========================

Data and Methods
==========================

#### Data

#### Methods

Results
==========================

Discussion
==========================

References {#references .unnumbered}
==========

### 2016-06-23 writing/JSSDraft/JSSDraft.Rmd
Figure out bibliography for JSS draft

@@
 title:
-  formatted: "\\pkg{eesim}: A Package to Simulate Environmental Epidemiology Time Series for Simulation Studies and Power Analyses"
+  formatted: "\\pkg{eesim}: An R Package to Simulate Environmental Epidemiology Time Series for Simulation Studies and Power Analyses"
   # If you use tex in the formatted title, also supply version without
-  plain:     "eesim: A Package to Simulate Environmental Epidemiology Time Series for Simulation Studies and Power Analyses"
+  plain:     "eesim: An R Package to Simulate Environmental Epidemiology Time Series for Simulation Studies and Power Analyses"
   # For running headers, if needed
@@
   # at least one keyword must be supplied
-  formatted: [environmental epidemiology, simulation studies, power analysis, air pollution]
-  plain:     [environmental epidemiology, simulation studies, power analysis, air pollution]
+  formatted: [R, environmental epidemiology, simulation studies, power analysis, air pollution]
+  plain:     [R, environmental epidemiology, simulation studies, power analysis, air pollution]
 preamble: >
   \usepackage{amsmath}
-output: rticles::jss_article
+output:
+  rticles::jss_article:
+    citation_package: natbib
+bibliography: mybibfile.bib
 ---
 
-# Purpose of package
+# Introduction
 
@@
 
-# Package functionality
+# Basic overview of package
+
+# Implementing the package
 
@@
 ## Customizing simulations
+
+Example citation [@Dirac1953888].
+
+# Discussion
+
+# Acknowledgments
+
+# References

### 2016-06-23 R/assess_performance.R
Move an example from regular code for `check_sims`
Otherwise, it's causing build errors with the package.

@@ -111,6 +111,9 @@ power_beta <- function(df){
 #'
 #' @examples
 #' check_sims(df, true_rr = 1.02)
+#' \dontrun{
+#' check_sims(rep_df, true_rr = 1.02)
+#' }
 #'
 #' @export
 #'
@@ -124,7 +127,7 @@ check_sims <- function(df, true_rr){
   out <- cbind(a, b, c, d, e)
   return(out)
 }
-check_sims(rep_df, true_rr = 1.02)
+
 #'
 #' Power Calculations
 #'

### 2016-06-24 writing/JSSDraft/JSSDraft.Rmd
Add some references to Bibtex file
Work on paper drafts

@@
 abstract: >
-  The abstract of the article.
+  Simulation studies are important in environmental epidemiology research on air pollution, temperature, and other exposures.
+For example, studies have simulated time series of environmental exposure and health outcome data to compare time series and case-crossover methods; explore models to test for mortality-displacement; investigate the impacts of measurement error and spatial misalignment of exposure and outcome measurements; and search for confounding by other time-varying covariates.
+
+  Developing the code to simulate environmental time series for these types of studies can be time consuming, and simulation methods have varied across previous simulation studies, complicating the comparison of results across different studies.
+Here, we present an R package that automates simulation of environmental time series for such simulations, offering both sensible defaults for each stage of the simulation process and also the option to customize any of the simulation steps while using defaults for any steps that do not require customization.
+This packge can be used to aid simulation studies to explore environmental epidemiology models relating acute health outcomes and daily environmental exposures; it can also be used to estimate the power of studies when planning or proposing future research.
 keywords:
@@
 
-* Not consistent across different studies
-* Time consuming to create, often only need to customize certain parts
+For many environmental time series studies, including those investigating air pollution and temperature, both exposures and outcomes can have complex seasonal and longer-term patterns over the study period.
+For example, air pollution concentrations can change seasonally, both because of changes in emission patterns and meteorological patterns (for example, winter inversion layers can cause a build-up of particulate matter in the winter, while high temperatures in the summer can accelate ozone formation from precursors); air pollution concentrations can also change over years, especially with evolving Clean Air policies.
+
+Exposures, outcomes, and their relationship with each other can all have complex patterns that could cause problems in epidemiological studies.
+For example, \citet{Armstrong2014} explain how autocorrelation in temperature data (e.g., days right after a very severe heat wave are likely to be hotter than days right after a milder heat wave) can introduce spurious results in a model testing for evidence of short-term mortality displacement following heat waves.
+
+There are a few important advantages to having an open-source software package that can simulate this type of environmental time series for simulation studies.
+First, current methods of simulating environmental time series, to use to assess modeling choices, vary across different studies.
+While, ideally, results from a simulation study would be robust to different choices in how to simulate the data used in the study, this may not always be the case; if so, results across studies using different methods of simulating data would not be comparable.
+This package offers a tool that both allows different research to simulate environmental time series for simulation studies in a consistent way, it also offers several methods of simulating time series, and so allows researchers to check if simulation results are sensitive to the method of simulating study data.
+Tools for consistency across studies have helped in the progress of environmental epidemiology research; for example, the public availability of the NMMAPs dataset--which included air pollution, weather, and mortality counts in over 100 US communities for 14 years-- allowed researchers around the world to explore hypotheses and investigate health effects of air pollution and temperature using the same data, and so differences in results could be traced to methodological differences rather than random differences in environmental datasets.
+[We could add some on the usefulness of (1) reproducible studies and (2) open source software for scientific studies here.]
+
+Further, developing the code to simulate environmental time series in an appropriate way can be time consuming.
+While researchers often need to customize certain parts of a simulation, to test a particular research question, often most of the simulation can be performed in a way that is generalizable across most other environmental time series studies.
+Therefore, by providing simulation tools with reasonable defaults, as well as room to customize at each step of the simulation, this package can make environmental epidemiolgoy simulation studies much more efficient and also make simulation a reasonable alternative for determining expected power when planning and proposing future studies.
+
+# Examples of studies that simulated environmental time series
+
+## Mortality displacement
+
+\citet{Armstrong2014} used a simulation study to investigate whether evidence of short-term mortality displacement following heat waves was a result of model artifacts.
+To do this, they aimed to simulate a time series of heat wave (binary) effects with heat wave effects, none of which were caused by short-term mortality displacement.
+They then showed that, when they fit the model from the previous study to this simulated dataset, the model generated results suggestive of short-term mortality displacement, and less displacement with more severe heat waves than less severe heat waves, which were similar to the findings of the study they were assessing.
+Overall, this analysis suggested that the earlier study's results might have been an artifact of the modeling approach they used, and so their results could be consistent with a hypothesis of no mortality displacement.
+
+To do this simulation, which required simulating a time series with a heat wave effect on mortality, but no short-term displacement to explain some of that excess mortality, \citet{Armstrong2014} started with a real dataset of daily temperature and mortality counts.
+For this, they used the NMMAPS dataset (1987--2000) from Atlanta, GA, which was one of the datasets used in the original study.
+They fit a model of expected log mortality count for this real data regressed on daily (lag 0) temperature and a smooth function of time (to capture seasonal and long-term variation).
+They then predicted from this model for the days in their simulation, to generate a smoothed estimate of expected daily mortality on each day (note: this daily expected count is a function of observed temperature over the study period; it sounds like they used observed data directly for the exposure time series).
+They then drew simulated daily mortality counts from a Poisson distribution with $\lambda$ equal to this daily expected mean mortality for each day in the simulation period.
+[Code they used for this simulation analysis is available as an online supplement.]
+
+## Comparing case-crossover and time series
+
+## Investigating confounding
+
+## Exploring spatial misalignment between exposure and outcome measurements
 
@@
 
-Example citation [@Dirac1953888].
-
 # Discussion

### 2016-06-24 writing/SimulationPaper/SimulationPaper.Rmd
Work on paper drafts

@@
 
+#### Mortality displacement
+
+Some immediate fatalaties during natural disasters might result from very frail people dying a few days earlier than under the counterfactual of no disaster.
+For example, [evidence on mortality displacement following heat waves, other disasters].
+While these deaths still represent excess premature mortality related to the disaster, they represent a much shorter loss of life-time than premature deaths that were advanced by months or years.
+Further, in model designs that use times near to the disaster as controls, including time series and case-crossover designs, displaced mortality could cause the estimated baseline mortality during control periods to be artifically low, so biasing the acute disaster effect estimate to be too high [we might want to do a figure that sketches this out].
+
+#### Delayed effects
+
+Disasters could elevate mortality risk in the week or weeks after the acute exposure.
+For example, [Florida study of mortality in the 2004 hurricane season].
+If a study uses controls nearby in time to the disaster, here there would be a bias to underestimate acute disaster effects (the opposite direction of bias than in the case of mortality displacement effects).
+
+A higher-than-average mortality count in the days around a disaster could also be related to autocorrelation in the exposure variable, particularly when a "disaster" is identified as a string of days when a continous exposure outcome is above a certain threshold.
+For example, @Armstrong2014 note that mortality is expected to be elevated on days around a severe heat wave because temperatures tend to be higher than normal on days just before and after a severe heat wave, and mortality risk is positively associated with temperature across the range of summer temperatures, whether a day meets the threshold for a heat wave or not.
+A similar pattern could exist for "smokewave" days, used in some studies to explore health risks related to forest fires [Coco's Epidemiology paper].
+
 Data and Methods
@@
 
+We generated simulated time series of disaster exposure and mortality outcomes using the `eesim` R package.
+
+We simulated several environmental time series, based on different scenarios of (1) the size of the acute effect of the disaster on mortality risk; (2) the size of the delayed or mortality displacement effect; and (3) the length of either the delayed or mortality displacement effect.
+We based the range of sizes used for each of these values on reasonable values from studies of the health impacts of heat waves and other natural disasters.
+
 #### Methods
+
+For each simulated time series dataset, we fit three models to estimate acute disaster effects: (1) a time series model; (2) a case-crossover model; and (3) a cross-year model.
+This last model compares mortality counts during the disaster to a set of control periods within the same city.
+However, unlike the case-crossover model, it uses control periods from the same month and days of the week as the case in neighboring years, rather than from the same days of the week and month as the case in the same year as the case.
+This model therefore controls for seasonality by design, but requires addition control to account for year-to-year trends in average mortality.
+
+As a sensitivity analysis, we re-ran our main analysis using environmental time series datasets simulated using some other common simulation methods.
+We tested results using [other simulation methods].
 
@@
 
+#### Performance of three models
+
+#### Sensitivity analysis
+
 Discussion
@@
 
+#### Implications for epidemiological studies of natural disasters
+
+#### More complex models to model delayed or mortality displacement effects
+
+Methods exist to model effect sizes over a period that includes the disaster and the following weeks. [More about dlnm.]
+
+Why researchers might want to model acute effects without using these more complex models:
+
+- Low power (especially with rarer health outcomes, may run into very low power when modeling the more complex models required to assess lags)
+- Interest might be in immediate impacts (e.g., what should city planners be prepared for during a disaster, in terms of health care needs, medical examiners?) rather than in estimating the total impacts (e.g., what is the total attributable impact?)
+- May not have a good idea of the time range of delayed or mortality displacement effects.
+If it's longer than the model includes, what are the impacts?
+Very long time ranges (e.g., month or more) could start to interfere with seasonal patterns, right?
+
+#### Other advantages / disadvantages of each model type
+
+For single disasters, there is the chance for some bias in comparing to controls in the same year if the event occurs during the yearly dip or peak in mortality trends.
+For example, [Jennifer's ozone and Olympics study].
+A model that uses control periods from the same time of the year (i.e., same week) of other years might prevent this bias.
+
+#### Delayed / mortality displacement effects that extend past a year
+
+For very severe disasters, there is some evidence that delayed or mortality displacement effects might span into the following year.
+[French and Russian heat wave examples, hurricane examples].
+In this case, even a study design that uses controls from other years could be biased by how these effects influence control periods.
+
 References {#references .unnumbered}

### 2016-06-25 writing/JSSDraft/JSSDraft.Rmd
Work on JSS draft

@@
 ## Using simulated time series for power analysis
+
+Simulation studies can be used to determine effective sample size or power when planning or proposing future research studies.
+This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable \citep{Bellan2015, Johnson2015, Burton2009}.
+Simulation studies can often be used to generate power analyses under a more realistic set of assumptions \cite{Burton2009}.
+
+For example, \citet{Bellan2015} used a simulation study to determine the power and statistical validity of two potential Ebola vaccine trials, a randomized controlled trial and a stepped-wedge cluster trial.
+Simulation was useful in this case because the trial would be conducting in a population with clustering by geographical district and with a temporally changing hazard of Ebola infection over the trial period \citep{Bellan2015}.
+
+\citet{Johnson2015} give an example of using simulation for a power analysis of a study of numbers of ticks on grouse chicks in a study design with multi-level clustering of the study population, both by geographic area and, within that, by brood.
+
+\citet{Burton2009} used simulation to investigate the power of case-control studies of genetic associations for chronic diseases.
 
@@
 
+A time series of simulated count outcome data with Poisson random noise can be simulated using the following model \citep{Johnson2015}:
+
+\begin{equation}
+\eta_t = \beta_0 + \sum_{m=1}^{p}\beta_{m}X_{m,t} 
+\end{equation}
+
+\begin{equation}
+\eta_t = log(\lambda_t)
+\end{equation}
+
+\begin{equation}
+y_i \sim Pois(\lambda_t)
+\end{equation}
+
+\noindent which is a Generalized Linear Model where $\eta_t$ is the linear predictor and a log link is used.
+
 # Implementing the package
@@
 
+Running a power analysis requires specifications of effect sizes, baseline counts of health outcomes, and shape of seasonal trends in exposures and outcomes.
+These values can be specified at realistic values based on previous research \citep{Johnson2015}; if no relevant previous research exists, in some cases a pilot study can be used to determine reasonable values to use for these inputs \citep{Johnson2015}.
+
 ## Using simulated time series to test modeling questions
@@
 ## Evaluating simulations based on these time series
+
+A number of different metrics are used to evaluate the results of simulation studies. These include:
+
+* False positive rate / type I error rate (percent of time significant effect is detected-- i.e., null hypothesis is rejected-- when there is not a true effect) \citep{Bellan2015, Johnson2015}
+* Statistical power to detect a true effect of a certain size (percent of simulations in which a true effect of a certain size is identified with a p-value of 0.05 or lower; i.e., probability of rejecting the null hypothesis when the null hypothesis is false) \citep{Bellan2015, Johnson2015}
+* Precision of effect estimate (average confidence interval width over all simulations) \citep{Johnson2015}
+* 95% confidence inverval coverage (percent of 95% confidence intervals from simulations that include the true effect size) \citep{Johnson2015}
+* Bias in effect estimate \citep{Johnson2015}
+* Margin of error [?] \citep{Johnson2015}
+* Sample size required to detect a certain effect size \citep{Burton2009}
+* Time to achieve required number of cases \citep{Burton2009}
 
@@
 
+## Limitations
+
+We currently do not include overdispersion when simulating health outcome counts.
+\citet{Johnson2015} discusses why this might be problematic and suggests some methods for incorporating overdispersion in a simulation in cases where count outcome data is expected to include overdispersion.
+They suggest that the following model can be used to generate the expected outcome count with overdispersion-- model the linear predictor as:
+
+\begin{equation}
+\eta_t = \beta_0 + \sum_{m=1}^{p}\beta_{m}X_{m,t} + \epsilon_t
+\end{equation}
+
+\begin{equation}
+\eta_t = log(\lambda_t)
+\end{equation}
+
+\begin{equation}
+y_i \sim Pois(\lambda_t)
+\end{equation}
+
+\noindent where $\epsilon_t$ represents the overdispersion and follows a normal distribution:
+
+\begin{equation}
+\epsilon_t \sim N(0, \sigma_{\epsilon}^2)
+\end{equation}
+
+In the case of no overdispersion, $\sigma_{\epsilon}^2$ equals 0. In the case of overdispersion, $\sigma_{\epsilon}^2$ will be larger than 0.
+
 # Acknowledgments

### 2016-06-25 writing/JSSDraft/JSSDraft.Rmd
Work on JSS draft

@@
 Simulation studies can be used to determine effective sample size or power when planning or proposing future research studies.
-This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable \citep{Bellan2015, Johnson2015, Burton2009}.
+This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable \citep{Bellan2015, Johnson2015, Burton2009, Alfons2010}.
 Simulation studies can often be used to generate power analyses under a more realistic set of assumptions \cite{Burton2009}.
@@
 
+A number of R packages have been created to do power analyses through simulations for a variety of purposes and contexts, including for genetic association studies \citep{Gaye2015}, microbiome studies \citep{Kelly2015}, network data \citep{Handcock2008}, survey statistics \citep{Alfons2010}, ...
+.
+More general-purpose R packages also exist for simulations for power analyses, including packages for power analyses for generalized linear mixed models \cite{Green2016}, ...
+.
+
 ## Using simulated time series to test modeling questions
+
+Simulation studies can also be used to assess different models and modeling choices.
+This way of assessing models can be particularly useful for complex statistical methods, which can often be very hard to assess performance metrics of using analytical methods \citep{Alfons2010}.
+
+Several R packages exist that simulate data to allow model evaluation in a variety of domains.
+For example, \code{ss3sim} allows the simulation of fisheries stock assessment for use in evaluating methods in fishery stock assessments \citep{Anderson2014}.
 
@@
 First, current methods of simulating environmental time series, to use to assess modeling choices, vary across different studies.
-While, ideally, results from a simulation study would be robust to different choices in how to simulate the data used in the study, this may not always be the case; if so, results across studies using different methods of simulating data would not be comparable.
+While, ideally, results from a simulation study would be robust to different choices in how to simulate the data used in the study, this may not always be the case; if so, results across studies using different methods of simulating data would not be comparable \citep{Alfons2010}.
 This package offers a tool that both allows different research to simulate environmental time series for simulation studies in a consistent way, it also offers several methods of simulating time series, and so allows researchers to check if simulation results are sensitive to the method of simulating study data.
@@
 # Basic overview of package
+
+[Table / Figure of functions included and simulation process]
 
@@
 * False positive rate / type I error rate (percent of time significant effect is detected-- i.e., null hypothesis is rejected-- when there is not a true effect) \citep{Bellan2015, Johnson2015}
-* Statistical power to detect a true effect of a certain size (percent of simulations in which a true effect of a certain size is identified with a p-value of 0.05 or lower; i.e., probability of rejecting the null hypothesis when the null hypothesis is false) \citep{Bellan2015, Johnson2015}
+* Statistical power to detect a true effect of a certain size / empirical power (percent of simulations in which a true effect of a certain size is identified with a p-value of 0.05 or lower; i.e., probability of rejecting the null hypothesis when the null hypothesis is false) \citep{Bellan2015, Johnson2015, Gaye2015, Green2016}
 * Precision of effect estimate (average confidence interval width over all simulations) \citep{Johnson2015}
@@
 * Margin of error [?] \citep{Johnson2015}
-* Sample size required to detect a certain effect size \citep{Burton2009}
+* Sample size required to detect a certain effect size \citep{Burton2009, Gaye2015, Green2016}
 * Time to achieve required number of cases \citep{Burton2009}
@@
 ## Customizing simulations
+
+While R packages always offer the opportunity for a user to alter the original code to create customized solutions, this process can require extensive skill in R programming for anything but fairly straightforward functions.
+Here, we provide the opportunity at every step of the simulation process for a user to input custom R functions, from simple to complex, to be run through our functions, without requiring changes to the function code provided by the package.
+
+For each of the customization options, we have provided the input and output rules required for a custom function to use for that step.
+A custom function can include any code, and implement any functionality, as long as it adheres to these rules for the input it takes and output it produces.
+Further, through using R's "dots" functionality, custom functions can include inputs other than those specified by our interface rules, as long as the custom function inputs, at a minimum, the inputs required by the interface rules.
 

### 2016-06-25 writing/JSSDraft/JSSDraft.Rmd
Work on JSS draft

@@
 Simulation studies can be used to determine effective sample size or power when planning or proposing future research studies.
-This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable \citep{Bellan2015, Johnson2015, Burton2009, Alfons2010}.
-Simulation studies can often be used to generate power analyses under a more realistic set of assumptions \cite{Burton2009}.
+This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable \citep{Bellan2015, Johnson2015, Burton2009, Alfons2010, Beaujean2014}.
+Simulation studies can often be used to generate power analyses under a more realistic set of assumptions \cite{Burton2009, Beaujean2014}.
 
@@
 * False positive rate / type I error rate (percent of time significant effect is detected-- i.e., null hypothesis is rejected-- when there is not a true effect) \citep{Bellan2015, Johnson2015}
-* Statistical power to detect a true effect of a certain size / empirical power (percent of simulations in which a true effect of a certain size is identified with a p-value of 0.05 or lower; i.e., probability of rejecting the null hypothesis when the null hypothesis is false) \citep{Bellan2015, Johnson2015, Gaye2015, Green2016}
-* Precision of effect estimate (average confidence interval width over all simulations) \citep{Johnson2015}
-* 95% confidence inverval coverage (percent of 95% confidence intervals from simulations that include the true effect size) \citep{Johnson2015}
+* Power / Statistical power to detect a true effect of a certain size / empirical power (percent of simulations in which a true effect of a certain size is identified with a p-value of 0.05 or lower; i.e., probability of rejecting the null hypothesis when the null hypothesis is false) \citep{Bellan2015, Johnson2015, Gaye2015, Green2016, Beaujean2014}
+* Precision of effect estimate (average confidence interval width over all simulations; some studies give average half-width) \citep{Johnson2015, Beaujean2014}
+* Coverage / 95% confidence inverval coverage (percent of 95% confidence intervals from simulations that include the true effect size) \citep{Johnson2015, Beaujean2014}
 * Bias in effect estimate \citep{Johnson2015}
 * Margin of error [?] \citep{Johnson2015}
-* Sample size required to detect a certain effect size \citep{Burton2009, Gaye2015, Green2016}
+* Sample size required to detect a certain effect size \citep{Burton2009, Gaye2015, Green2016, Beaujean2014}
 * Time to achieve required number of cases \citep{Burton2009}
+* Power curve \citep{Beaujean2014}
+* Accuracy curve \citep{Beaujean2014}
+
+\citet{Beaujean2014} gives the following equation for calculating relative parameter estimate bias, given $\theta_{H}$ (true value of the parameter, as set when creating the simulated data) and $\hat{\theta}$ (value of the parameter estimated from applying a model to the simulated data):
+
+\begin{equation}
+\theta_{bias} = \frac{\hat{\theta} - \theta_{H}}{\theta_{H}}
+\end{equation}
+
+\citet{Beaujean2014} gives the following equation for calculating relative standard error bias, given $\sigma_{\hat{\theta}}$ (the standard deviation of the parameter estimations across the simulations) and $\hat{\sigma_{\theta}}$ (the average of the standard errors estimated for the simultations):
+
+\begin{equation}
+\sigma_{bias} = \frac{\hat{\sigma_{\theta}} - \sigma_{\hat{\theta}}}{\sigma_{\hat{\theta}}}
+\end{equation}
 
@@
 \citet{Johnson2015} discusses why this might be problematic and suggests some methods for incorporating overdispersion in a simulation in cases where count outcome data is expected to include overdispersion.
-They suggest that the following model can be used to generate the expected outcome count with overdispersion-- model the linear predictor as:
+They suggest that the following model can be used to generate the expected outcome count with overdispersion-- modeling the linear predictor as:
 
@@
 \eta_t = \beta_0 + \sum_{m=1}^{p}\beta_{m}X_{m,t} + \epsilon_t
-\end{equation}
-
-\begin{equation}
-\eta_t = log(\lambda_t)
-\end{equation}
-
-\begin{equation}
-y_i \sim Pois(\lambda_t)
 \end{equation}

### 2016-06-25 writing/JSSDraft/JSSDraft.Rmd
Work on JSS draft

@@
 \citet{Burton2009} used simulation to investigate the power of case-control studies of genetic associations for chronic diseases.
+
+\citet{Schoemann2014} demonstrate how to use simulation studies to assess power for study designs with planned missing data.
 
@@
 * Coverage / 95% confidence inverval coverage (percent of 95% confidence intervals from simulations that include the true effect size) \citep{Johnson2015, Beaujean2014}
-* Bias in effect estimate \citep{Johnson2015}
+* Bias in effect estimate / relative parameter estimate bias / relative bias in point estimate \citep{Johnson2015, Beaujean2014, Schoemann2014}
+* Relative standard error bias / standard error bias \citep{Beaujean2014, Schoemann2014}
 * Margin of error [?] \citep{Johnson2015}
-* Sample size required to detect a certain effect size \citep{Burton2009, Gaye2015, Green2016, Beaujean2014}
+* Sample size required to detect a certain effect size \citep{Burton2009, Gaye2015, Green2016, Beaujean2014, Schoemann2014}
 * Time to achieve required number of cases \citep{Burton2009}

### 2016-06-28 R/sim_exposure.R
Trying to figure out season_binexp

@@ -114,8 +114,8 @@ calc_t <- function(n, trend = "no trend", amp = .6, custom_func = NULL, ...){
 season_binexp <- function(n, p, trend, start.date = "2000-01-01"){
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
-  t <- calc_t(n = n, trend = trend)
-  p <- p*t
+  t <- calc_t(n = n, trend = trend, amp=amp)
+
   p <- #Magic
   for (i in 1:n){
     x <- sample(c(0, 1), size = 1, replace = T, prob = c(1 - p[i], p[i]))

### 2016-07-01 DraftOne.Rmd
Trying to figure out season_binexp
Adding bin_t

@@
 date: "June 1, 2016"
-output: word_document
+output: pdf_document
 ---
@@
 calc_t(n=10, trend="curvilinear")
+```
+
+Now a function for calc_t to use for seasonal binary exposure:
+
+```{r}
+bin_t <- function(n, p, trend = "no trend", amp = .3, custom_func = NULL, ...){
+  day <- c(1:n)
+  if (trend == "cos1"){
+    seasont <- p + amp * cos(2 * pi * (day / 365))
+    } else if (trend == "cos2"){
+    seasont <- p + amp/2 * cos(2 * pi * (day / 365)) +
+      ifelse(day < 639 & day > 274, amp/2 * cos(2 * (pi * (day / 365))), 0)
+    } else if (trend == "cos3"){
+      seasont <- p + .75 ^ (day / 365) * amp * cos(2 * pi * (day / 365))
+    } else if (trend == "linear"){
+      seasont <- p+1/(day/n)
+    } else if (trend == "curvilinear"){
+      seasont <- p + day * (2 / n) + day^2 * (-1 / n^2)
+    } else if (trend == "cos1linear"){
+      seasont <- (p + (day / n)) * (p + amp * cos(2 * pi * (day / 365)))
+    } else if (trend == "no trend"){
+      seasont <- p
+    } else if (trend == "custom" & !is.null(custom_func)) {
+      arguments <- list(...)
+      arguments$n <- n
+      seasont <- do.call(custom_func, arguments)
+    } else {
+      stop(paste0("`trend` value is not a valid choice. Please check the",
+                  " function documentation to select a valid option."))
+    }
+}
 ```

### 2016-07-01 R/sim_exposure.R

@@ -96,7 +96,19 @@ calc_t <- function(n, trend = "no trend", amp = .6, custom_func = NULL, ...){
   seasont <- seasont / mean(seasont)
   return(seasont)
 }
-
+#'
+#' Create a trend variable for binary exposure data
+#'
+#' This function creates a trend variable for binary exposure data which is centered at p and
+#' restricts the probability of exposure between 0 and 1.
+#'
+#' @inheritParams calc_t
+#' @param p A numeric value giving the mean probability of exposure
+#'
+#' @return
+#'
+#'
+#'
 #' Simulate binary exposure data with a seasonal trend
 #'
 #' This function simulates binary exposure data with a seasonal trend.

### 2016-07-05 writing/JSSDraft/JSSDraft.Rmd
Copyediting

--- 2016-06-25 writing/JSSDraft/JSSDraft.Rmd
@@
 For many environmental time series studies, including those investigating air pollution and temperature, both exposures and outcomes can have complex seasonal and longer-term patterns over the study period.
-For example, air pollution concentrations can change seasonally, both because of changes in emission patterns and meteorological patterns (for example, winter inversion layers can cause a build-up of particulate matter in the winter, while high temperatures in the summer can accelate ozone formation from precursors); air pollution concentrations can also change over years, especially with evolving Clean Air policies.
+For example, air pollution concentrations can change seasonally, both because of changes in emission patterns and meteorological patterns (for example, winter inversion layers can cause a build-up of particulate matter in the winter, while high temperatures in the summer can accelerate ozone formation from precursors); air pollution concentrations can also change over years, especially with evolving Clean Air policies.
 
@@
 There are a few important advantages to having an open-source software package that can simulate this type of environmental time series for simulation studies.
-First, current methods of simulating environmental time series, to use to assess modeling choices, vary across different studies.
+First, current methods of simulating environmental time series to use to assess modeling choices vary across different studies.
 While, ideally, results from a simulation study would be robust to different choices in how to simulate the data used in the study, this may not always be the case; if so, results across studies using different methods of simulating data would not be comparable \citep{Alfons2010}.
@@
 While researchers often need to customize certain parts of a simulation, to test a particular research question, often most of the simulation can be performed in a way that is generalizable across most other environmental time series studies.
-Therefore, by providing simulation tools with reasonable defaults, as well as room to customize at each step of the simulation, this package can make environmental epidemiolgoy simulation studies much more efficient and also make simulation a reasonable alternative for determining expected power when planning and proposing future studies.
+Therefore, by providing simulation tools with reasonable defaults, as well as room to customize at each step of the simulation, this package can make environmental epidemiology simulation studies much more efficient and also make simulation a reasonable alternative for determining expected power when planning and proposing future studies.
 
@@
 To do this, they aimed to simulate a time series of heat wave (binary) effects with heat wave effects, none of which were caused by short-term mortality displacement.
-They then showed that, when they fit the model from the previous study to this simulated dataset, the model generated results suggestive of short-term mortality displacement, and less displacement with more severe heat waves than less severe heat waves, which were similar to the findings of the study they were assessing.
+They then showed that when they fit the model from the previous study to this simulated dataset, the model generated results suggestive of short-term mortality displacement, and less displacement with more severe heat waves than less severe heat waves, which were similar to the findings of the study they were assessing.
 Overall, this analysis suggested that the earlier study's results might have been an artifact of the modeling approach they used, and so their results could be consistent with a hypothesis of no mortality displacement.

### 2016-07-06 writing/JSSDraft/JSSDraft.Rmd
Adding Strickland summary

@@
 
+\citet{Strickland2013} used a simulation study to investigate the effects of measurement error resulting from instrument imprecision and spacial misalignment on the estimates of health effects related to ambient air pollutant concentrations.
+They first used air-quality measurements obtained from several networks of monitoring stations in Atlanta to estimate various distributional properties associated with pollutant concentrations, including spatial and temporal autocorrelation.
+They used these estimates to generate daily two-dimensional pollutant fields for Atlanta at 5 km grids, producing "true" population-weighted average pollutant concentrations (TPWA).
+They then simulated grid-specific counts of daily ED visits using TPWA as a covariate, among others.
+
+The authors then created three different pollutant metrics based on their simulated data which contained additional measurement error and compared estimates of relative risk using these metrics to the RR = 1.05 they specified while simulating the TPWA.
+As expected they found that the population-weighted average metric resulted in less bias than measurements from a single monitoring station or measurements averaged across stations.
+
 # Basic overview of package

### 2016-07-08 DraftOne.Rmd

--- 2016-06-23 DraftOne.Rmd
@@
 calc_t(n=10, trend="curvilinear")
+```
+
+Now a function for calc_t to use for seasonal binary exposure:
+
+```{r}
+bin_t <- function(n, p, trend = "no trend", amp = .3, custom_func = NULL, ...){
+  day <- c(1:n)
+  if (p > .5 & amp >1-p){
+    stop(paste0("For p>.5, amp must be between 0 and 1-p."))
+  }
+  if (p < .5 & amp>p){
+    stop(paste0("For p<.5, amp must be between 0 and p."))
+  }
+  if (trend == "cos1"){
+    seasont <- p + amp * cos(2 * pi * (day / 365))
+    } else if (trend == "cos2"){
+    seasont <- p + amp/2 * cos(2 * pi * (day / 365)) +
+      ifelse(day < 639 & day > 274, amp/2 * cos(2 * (pi * (day / 365))), 0)
+    } else if (trend == "cos3"){
+      seasont <- p + .75 ^ (day / 365) * amp * cos(2 * pi * (day / 365))
+    } else if (trend == "linear"){
+      seasont <- p+1/(day/n)
+    } else if (trend == "curvilinear"){
+      seasont <- p + day * (2 / n) + day^2 * (-1 / n^2)
+    } else if (trend == "cos1linear"){
+      seasont <- (p + (day / n)) * (p + amp * cos(2 * pi * (day / 365)))
+    } else if (trend == "no trend"){
+      seasont <- p
+    } else if (trend == "custom" & !is.null(custom_func)) {
+      arguments <- list(...)
+      arguments$n <- n
+      seasont <- do.call(custom_func, arguments)
+    } else {
+      stop(paste0("`trend` value is not a valid choice. Please check the",
+                  " function documentation to select a valid option."))
+    }
+}
 ```

### 2016-07-12 writing/JSSDraft/JSSDraft.Rmd
Updating bib file to include Strickland

@@
 
-\citet{Strickland2013} used a simulation study to investigate the effects of measurement error resulting from instrument imprecision and spacial misalignment on the estimates of health effects related to ambient air pollutant concentrations.
+\citet{Strickland2015} used a simulation study to investigate the effects of measurement error resulting from instrument imprecision and spacial misalignment on the estimates of health effects related to ambient air pollutant concentrations.
 They first used air-quality measurements obtained from several networks of monitoring stations in Atlanta to estimate various distributional properties associated with pollutant concentrations, including spatial and temporal autocorrelation.

### 2016-07-22 writing/JSSDraft/JSSDraft.Rmd
Summarizing Roberts 2006

@@
 
+##PM and Mortality
+
+\citet{Roberts2006} used a simulation study to test the reliability of using AIC to determine the shape of the relationship between particulate matter and mortality.
+The authors used Chicago NMMAPS data, which is the default data set in eesim.
+They estimated the effects of confounders such as temperature on the mortality time series, all from this Chicago NMMAPS, and incorporated the estimates in generating mortality time series with author-specified relationships with PM dose.
+These explicitly specified relationships were piecewise linear relationships with and without a lower threshhold for effects.
+They could use our customization functions to specify these relationships.
+Then they estimated the shape of the dose-response relationship using a Poisson log-linear model and four different candidate models for the dose-response relation.
+They found the AIC for each of these models and determined whether the AIC was lowest for the models which reflected the true relationship between PM and response in the simulated data.
+They found that AIC does not always detect nonlinear relationships in dose-response data.
+
 # Basic overview of package

### 2016-07-22 writing/JSSDraft/JSSDraft.Rmd
Adding Gryparis

@@
 
+\citet{Gryparis2009} compared approaches to fitting health models using predicted exposures featuring spatial misalignment.
+The authors considered four levels of heterogeneity in exposure values over a geographic area and simulated exposure data for each.
+They simulated both continuous and binary health outcomes, which could be done using eesim, and evaluated several approaches to relating exposure to outcome, including Bayesian and RC-OOS approaches.
+Their simulation results suggested that several approaches may be appropriate to discover relationships in spatially misaligned data.
+
 ##PM and Mortality

### 2016-07-24 writing/JSSDraft/JSSDraft.Rmd
Summarizing Butland2013

@@
 
+\citet{Butland2013} performed a simulation study to compare models using geographically resolved, simulated air pollution measures and air pollution concentrations averaged over geographic areas.
+Recently, regional air pollution chemistry-transport models (CTMs) have been developed which can simulate air pollution concentrations over highly-resolved geographic areas.
+These are useful where pollutants cannot be measured or can only be measured in a few locations.
+The authors used simulation to investigate if and when it is better to use air pollution measurements from CTMs when estimating health outcomes rather than measures obtained from averaging over geographic grids based on monitor data.
+
+The authors first simulated "true" values for pollutant concentrations over 100 grid squares, each with a specific mean concentration.
+They generated daily values for each grid square over three years, including between-grid covariances.
+Then they sampled the number of daily deaths from a poisson with mean dependent on the "true" concentrations.
+Their generation of daily mortality counts could be done using eesim.
+The authors then simulate  monitor data with classical measurement error and CTM data with both classical and Berkson error.
+They separately fit the CTM data and regional averages from the monitor data to the simulated mortality counts and display several measures of model performance, including standard error of the estimates, coverage, and power, all of which can be produced using eesim.
+Their results suggested that there may be a significant penalty in the form of increased bias in health effect estimates for the use of CTMs over monitors due to the additional measurement error.
+
 ##PM and Mortality

### 2016-07-24 writing/JSSDraft/JSSDraft.Rmd
Summarizing Bobb2015

@@
 
+##High-dimensional problems
+
+\citet{Bobb2015} used simulation to assess the performance of a proposed method for estimating the health effects of multi-pollutant exposures called Bayesian kernel machine regression (BKMR).
+The authors generated exposure data by resampling from two representative data sets.
+They then specified several different relationships between response and the components of each pollutant mixture to generate response data, for which they would need to use the customization features in eesim.
+They fit kernel machine regression using both frequentist and Bayesian approaches, which would again require customization.
+Their model performance measures included standard error of the estimates, power, and coverage, all of which eesim offers.
+They found that there are advantages of using BKMR over KMR to detect relationships between mixture pollutants and health.
+
 # Basic overview of package

### 2016-07-24 writing/JSSDraft/JSSDraft.Rmd
Summarizing Bateson, Bateson, Peng

@@
 
+\citet{Bateson1999} compare several different case-crossover designs in the analysis of data with confounding seasonal trends.
+They generated both exposure and response data exhibiting each of eight different temporal trends, which can be done with eesim.
+They analyzed the simulated data using five different case-crossover designs, which could be done using the customization features of eesim.
+They assessed these models with measures such as mean variance of the estimates, relative efficiency, and coverage probability, all included in eesim.
+Based on the model performances, the authors concluded that symmetric bi-directional design with a lag period of one week can control for a large variety of confounding time trends.
+
 ## Investigating confounding
+
+\citet{Peng2006} investigated the effects of confounding factors and model choice on time series studies of air pollution and mortality.
+In their simulation study, the authors specified four scenarios differing in smoothness and concurvity between pollutant and time trend variable.
+They simulated exposure and mortality data based on a data from Minneapolis-St.
+Paul, for which they could use eesim's customization feature.
+They then fit Poisson regression models to each simulated data set, which could be done using eesim.
+They assessed the models with measures including AIC and BIC and found that different approaches to smoothing functions of time have different effects on the bias and variance of the effect estimates depending on factors such as the concurvity of the data.
+
+\citet{Bateson2001} investigated confounding in case-crossover studies.
+They simulated continuous exposure data with a cosinusoidal trend and generated Poisson-distributed outcome counts with the same trend, all of which can be done with eesim's functions.
 

### 2016-08-01 custom_examples.Rmd
Adding sim_exposure

---
title: "Examples of Customization"
output: pdf_document
---

#Exposure

##calc_t

```{r}
sintrend <- function(n, y){
  day <- c(1:n)
  1 + y * sin(2 * pi * (day / 365))
}
calc_t(n = 5, y = .3, trend = "custom", custom_func = sintrend)
plot(calc_t(1000, y = .3, trend = "custom", custom_func = sintrend))
```

bin_t is similar.

##binary and continuous exposure

```{r}
binary_exposure(n=5, p = .2, y = .1, trend = "custom", custom_func = sintrend)
plot(binary_exposure(n=400, p = .2, y = .1, trend = "custom", custom_func = sintrend))

continuous_exposure(n = 5, y = .6, mu = 100, sd = 10, trend = "custom", custom_func = sintrend)
plot(continuous_exposure(n=1000, y = .6, mu = 5, sd = .5, trend = "custom", custom_func = sintrend))
```

### 2016-08-01 R/sim_exposure.R

@@ -131,10 +131,10 @@ bin_t <- function(n, p, trend = "no trend", amp = .01, start.date = "2000-01-01"
 #'
 #' @export
 #'
-binary_exposure <- function(n, p, trend = "no trend", amp, start.date = "2000-01-01"){
+binary_exposure <- function(n, p, trend = "no trend", amp = .01, start.date = "2000-01-01", custom_func = NULL, ...){
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
-  t <- bin_t(n, p, trend, amp)
+  t <- bin_t(n, p, ...)
   x <- rbinom(length(t), size = 1, prob = t)
   df <- data.frame(date, x)
   return(df)
@@ -142,7 +142,7 @@ binary_exposure <- function(n, p, trend = "no trend", amp, start.date = "2000-01
 
 #' Simulate continuous exposure data
 #'
-#' This function simulates a time series of continuous exposure values with or without a
+#' Simulates a time series of continuous exposure values with or without a
 #' seasonal trend.
 #'
 #' @param mu A numeric value giving the average of the exposure distribution.
@@ -158,11 +158,42 @@ binary_exposure <- function(n, p, trend = "no trend", amp, start.date = "2000-01
 #' @export
 continuous_exposure <- function(n, mu, sd, trend = "no trend", amp, start.date = "2000-01-01", ...){
   day <- c(1:n)
-  t <- calc_t(n, trend, amp)
+  t <- calc_t(n, trend, amp, start.date, ...)
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
-  mu <- mu * t
-  x <- rnorm(n, mean = mu, sd = sd)
+  newmu <- mu * t
+  x <- rnorm(n, newmu, sd)
   df <- data.frame(date, x)
   return(df)
 }
+#'
+#' Simulate exposure data
+#'
+#' Simulates binary or continuous exposure data with or without seasonal trends.
+#'
+#' @param central A numeric value specifying the mean probability of exposure (for binary data) or
+#' the mean exposure value (for continuous data)
+#' @param exposure_type A character string specifying the type of exposure.  Choices are "binary" or "continuous".
+#'
+#' @inheritParams continuous_exposure
+#'
+#' @return A data frame with n rows, a column for the date of exposure, and a column giving the exposure value.
+#'
+#' @examples
[…]

### 2016-08-01 R/sim_exposure.R
Deleting parameter "sd"

@@ -156,13 +156,13 @@ binary_exposure <- function(n, p, trend = "no trend", amp = .01, start.date = "2
 #' continuous_exposure(n = 5, mu = 100, sd = 10, trend = "cos1")
 #'
 #' @export
-continuous_exposure <- function(n, mu, sd, trend = "no trend", amp, start.date = "2000-01-01", ...){
+continuous_exposure <- function(n, mu, trend = "no trend", amp, start.date = "2000-01-01", ...){
   day <- c(1:n)
   t <- calc_t(n, trend, amp, start.date, ...)
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
   newmu <- mu * t
-  x <- rnorm(n, newmu, sd)
+  x <- rnorm(n, newmu, ...)
   df <- data.frame(date, x)
   return(df)
 }
@@ -185,14 +185,14 @@ continuous_exposure <- function(n, mu, sd, trend = "no trend", amp, start.date =
 #'
 #' @export
 #'
-sim_exposure <- function(n, central, sd, trend = "no trend", exposure_type = "binary", amp, start.date = "2000-01-01", ...){
+sim_exposure <- function(n, central, trend = "no trend", exposure_type = "binary", amp, start.date = "2000-01-01", ...){
   if(exposure_type=="binary"){
     p <- central
     df <- binary_exposure(n, p, trend, amp, start.date, ...)
   }
   else if(exposure_type == "continuous"){
     mu <- central
-    df <- continuous_exposure(n, mu, sd, trend, amp, start.date, ...)
+    df <- continuous_exposure(n, mu, trend, amp, start.date, ...)
   }
   return(df)
 }

### 2016-08-09 custom_examples.Rmd
Fixing sim_exposure to include default parameters

@@
   day <- c(1:n)
-  1 + y * sin(2 * pi * (day / 365))
+  base <- 1 + y * sin(2 * pi * (day / 365))
+  rnorm(n, mean = base, sd = 1)
 }

### 2016-08-09 R/customizable_functions.R

@@ -18,8 +18,8 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, central = NA,
 #'                     metric = "temp")
 #'
 #' @export
-sim_exposure <- function(n, central, trend = "no trend", amp = .6,
-                         exposure_type = NA,
+sim_exposure <- function(n, central = NULL, trend = "no trend", amp = .6,
+                         exposure_type = NULL,
                          start.date = "2001-01-01", custom_func = NULL, ...){
   arguments <- list(...)
   arguments$n <- n
@@ -32,11 +32,12 @@ sim_exposure <- function(n, central, trend = "no trend", amp = .6,
   } else if (!(is.null(custom_func))){
     start.date <- as.Date(start.date)
     date <- seq(from = start.date, by = 1, length.out = n)
-    exposure <- do.call(custom_func, arguments)
+    exposure1 <- do.call(custom_func, arguments)
+    exposure <- data.frame(date, exposure1)
   } else {
-    stop(paste0("If a custom function is not used to simulate randomness in the",
-                "exposure variable, then `outcome_type` must be specified as",
-                "either `binary` or `continuous`."))
+    stop(paste0("If a custom function is not used to simulate randomness in the
+                exposure variable, then the parameters central and
+                exposure_type must be specified."))
   }
   return(exposure)
 }

### 2016-08-11 writing/JSSDraft/JSSDraft.Rmd
Add package overview (section 3)

@@
 
-[Table / Figure of functions included and simulation process]
+[Chart / Figure of functions included and simulation process]
+
+eesim consists of four major parts: generating exposure time series, generating outcome time series, fitting models, and assessing models.
+Exposure time series can be generated using the sim\_exposure function with a variety of built-in seasonal trends or using a custom trend specified by the user.
+This function also allows the user to choose either binary or continuous exposure values.
+Using the resulting exposure values as an input, outcomes can then be generated using the sim\_outcome function.
+For outcomes as well as exposure, the user may specify a built-in seasonal trend or a custom trend.
+
+The package includes functions for fitting two different models of the exposure and outcome data: a generalized linear model with splines and a case-crossover model.
+The function rep\_sims runs many simulations for a given data set and model.
+Finally, the function check\_sims returns several measures of model performance including the mean estimate of relative risk, mean variance of the relative risk estimates, percent bias, coverage, and power.
 

### 2016-08-18 R/customizable_functions.R
Attempt to fix sim_outcome, problem with unused arguments

@@ -116,7 +116,7 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
 sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                         amp = .6, rr = 1.01, start.date="2000-01-01",
                         cust_base_func = NULL, cust_lambda_func = NULL,
-                        cust_args = NULL,...){
+                        cust_args = NULL){
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = nrow(exposure))
   if (is.null(cust_args)){
@@ -124,7 +124,6 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
   } else {
     arguments <- cust_args
   }
-  arguments$n <- n
   if(is.null(cust_base_func) & is.null(cust_lambda_func)){
     if(is.null(average_outcome)){
       stop(paste0("If custom functions are not used to generate outcomes,
@@ -140,6 +139,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
     outcome <- rpois(n = nrow(exposure), lambda = lambda)
   }
   else if (is.null(cust_lambda_func) & !is.null(cust_base_func)){
+    arguments$n <- nrow(exposure)
     arguments$average_outcome <- average_outcome
     arguments$exposure <- exposure
     arguments$rr <- rr
@@ -150,11 +150,21 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
     outcome <- rpois(n = nrow(exposure), lambda = lambda)
   }
   else if (is.null(cust_base_func) & !is.null(cust_lambda_func)){
-    arguments$average_outcome <- average_outcome
-    arguments$exposure <- exposure
-    arguments$rr <- rr
-    arguments$trend <- trend
-    arguments$amp <- amp
+    if(!is.null(average_outcome)){
+      arguments$average_outcome <- average_outcome
+    }
+    if(!is.null(exposure)){
+      arguments$exposure <- exposure
+    }
+    if(!is.null(rr)){
+      arguments$rr <- rr
+    }
+    if(!is.null(trend)){
+      arguments$trend <- trend
+    }
+    if(!is.null(amp)){
+      arguments$amp <- amp
+    }
     baseline <- create_baseline(n = nrow(exposure),
                                 average_outcome = average_outcome,
                                 trend = trend,
[…]

### 2016-08-23 custom_examples.Rmd
Attempt to fix sim_outcome, problem with unused arguments
Add example custom functions, all options working for sim_outcome

--- 2016-08-01 custom_examples.Rmd
@@
   day <- c(1:n)
-  1 + y * sin(2 * pi * (day / 365))
+  base <- 1 + y * sin(2 * pi * (day / 365))
+  rnorm(n, mean = base, sd = 1)
 }
 calc_t(n = 5, y = .3, trend = "custom", custom_func = sintrend)
-plot(calc_t(1000, y = .3, trend = "custom", custom_func = sintrend))
+plot(calc_t(1000, y = 2, trend = "custom", custom_func = sintrend))
 ```
@@
 ```
+
+##Custom functions for outcome
+
+```{r}
+#For baseline:
+custombase <- function(n, slope, intercept){
+  day <- c(1:n)
+  baseline <- day*slope + intercept
+  return(baseline)
+}
+
+#Example
+custombase(n=5, slope = .3, intercept = 55)
+
+#For lambda:
+customlambda <- function(exposure, rr, constant, baseline){
+  log_lambda <- log(baseline) + log(rr) * exposure + constant
+  lambda <- exp(log_lambda)
+  return(lambda)
+}
+
+#Example:
+testlambda <- customlambda(testexp$x, testbase$baseline, rr=1.01, constant = 4)
+```

### 2016-08-23 R/customizable_functions.R

@@ -111,6 +111,16 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
   return(lambda)
 }
 #' @param cust_args A list of arguments and their values used in the user-specified custom functions
+#'
+#' @example
+#' sim_outcome(testexp, average_outcome = 22)
+#'
+#' sim_outcome(exposure=testexp, cust_base_func = custombase,
+#' cust_base_args = list(n=nrow(testexp), slope = .2, intercept = 55))
+#'
+#' sim_outcome(exposure = testexp, average_outcome = 22, cust_lambda_func = customlambda,
+#' cust_lambda_args = list(exposure = testexp$x, rr=1.02, constant = 4))
+#'
 #' @export
 
 sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
@@ -128,14 +138,14 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                                 average_outcome = average_outcome,
                                 trend = trend,
                                 amp = amp)
-    lambda <- create_lambda(baseline = baseline$baseline,
+    lambda <- create_lambda(baseline = baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- rpois(n = nrow(exposure), lambda = lambda)
   }
   else if (is.null(cust_lambda_func) & !is.null(cust_base_func)){
     baseline <- do.call(cust_base_func, cust_base_args)
-    lambda <- create_lambda(baseline = baseline$baseline,
+    lambda <- create_lambda(baseline = baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- rpois(n = nrow(exposure), lambda = lambda)

### 2016-08-23 R/customizable_functions.R
Adjust custom argument parameters in create_sims

@@ -179,7 +179,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #' for date, exposure, and outcomes, and estimates from fitting models
 #'
 #' @examples
-#' create_sims(n_reps=10, n=1000, central = 100, sd = 10, exposure_type="continuous",
+#' create_sims(n_reps=10, n=10, central = 100, sd = 10, exposure_type="continuous",
 #' exposure_trend = "cos1", exposure_amp = .6, average_outcome = 22, outcome_trend = "no trend",
 #' outcome_amp = .6, rr = 1.01)
 #'
@@ -187,7 +187,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 create_sims <- function(n_reps, n, central, sd, exposure_type, exposure_trend, exposure_amp,
                   average_outcome, outcome_trend, outcome_amp, rr, start.date = "2000-01-01",
                   cust_exp_func = NULL, cust_exp_args = NULL, cust_base_func = NULL,
-                  cust_lambda_func = NULL, cust_out_args = NULL){
+                  cust_lambda_func = NULL, cust_base_args = NULL, cust_lambda_args = NULL){
   exposure <- lapply(rep(n, times = n_reps), sim_exposure, central = central, sd = sd,
                      exposure_type = exposure_type, amp = exposure_amp, trend = exposure_trend,
                      start.date = start.date, cust_exp_func = cust_exp_func,
@@ -195,7 +195,7 @@ create_sims <- function(n_reps, n, central, sd, exposure_type, exposure_trend, e
   outcome <- lapply(exposure, sim_outcome, average_outcome = average_outcome,
                     trend = outcome_trend, amp = outcome_amp, rr = rr, start.date = start.date,
                     cust_base_func = cust_base_func, cust_lambda_func = cust_lambda_func,
-                    cust_args = cust_out_args)
+                    cust_base_args = cust_base_args, cust_lambda_args = cust_lambda_args)
   return(outcome)
 }
 

### 2016-08-30 R/customizable_functions.R
Adjust custom_exposure to allow user to put in different start date

@@ -1,10 +1,36 @@
-#' Pulls exposure series from Chicago NMMAPS data set
+#' Pull exposure series from data set
+#'
+#' By default, this function pulls exposure data from the Chicago NMMAPS data set in the dlnm
+#' package.  The user may specify a different data set from which to pull exposure
+#' values.
+#'
+#' @param n A numeric value specifying the number of days for which to obtain an exposure value.
+#' @param df Data frame from which to pull exposure values.
+#' @param metric A character string specifying the desired exposure metric.
+#'    Options are:
+#'    \itemize{
+#'      \item"cvd"
+#'      \item"resp"
+#'      \item"temp"
+#'      \item"dptp"
+#'      \item"rhum"
+#'      \item"pm10"
+#'      \item"o3"}
+#' @param start.date A date of the format "yyyy-mm-dd" from which to begin pulling exposure values
 #'
 #' @examples
-#' custom_exposure(n = 5, metric = "temp")
-custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, central = NA,
-                            metric = "temp"){
+#' custom_exposure(n = 5, metric = "temp", start.date = "1995-01-01")
+#'
+#' @export
+custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp", start.date = NULL){
+  if(!is.null(start.date)){
+    date1 <- which(df$date==start.date)
+    enddate <- date1+n
+    exposure <- df[date1:enddate, metric]
+  }
+  else{
   exposure <- df[1:n, metric]
+  }
   return(exposure)
 }
 
@@ -135,7 +161,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                                 average_outcome = average_outcome,
                                 trend = trend,
                                 amp = amp)
-    lambda <- create_lambda(baseline = baseline,
+    lambda <- create_lambda(baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- rpois(n = nrow(exposure), lambda = lambda)

### 2016-08-31 R/customizable_functions.R
Add documentation

@@ -16,10 +16,13 @@
 #'      \item"rhum"
 #'      \item"pm10"
 #'      \item"o3"}
-#' @param start.date A date of the format "yyyy-mm-dd" from which to begin pulling exposure values
+#' @param start.date A date of the format "yyyy-mm-dd" from which to begin pulling exposure
+#' values.  Dates in the Chicago NMMAPS data set are from 1987-01-01 to 2000-12-31.
+#'
+#' @return A numeric vector of exposure values
 #'
 #' @examples
-#' custom_exposure(n = 5, metric = "temp", start.date = "1995-01-01")
+#' custom_exposure(n = 5, metric = "temp", start.date = "2000-01-01")
 #'
 #' @export
 custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp", start.date = NULL){
@@ -36,14 +39,21 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp", start.
 
 #' Simulate random series of exposure values
 #'
+#' This function simulates binary or continuous exposure values with or without seasonal trends.
+#' It also allows for a custom function for exposure trend.
+#'
+#' @param cust_exp_function The name of a function from which to generate custom exposure values
 #' @param cust_exp_args A list of arguments used in the user-specified custom function
+#' @inheritParams std_exposure
+#'
+#' @return A data frame with two columns: date and exposure values
 #'
 #' @examples
 #' sim_exposure(n = 5, central = 0.25, exposure_type = "binary", amp = .02)
 #' sim_exposure(n = 5, central = 100, sd = 10, amp = .6,
 #'              exposure_type = "continuous")
 #' sim_exposure(n = 5, cust_exp_func = "custom_exposure",
-#'                     cust_exp_args = metric = "temp")
+#'                     cust_exp_args = list(metric = "temp"))
 #'
 #' @export
 sim_exposure <- function(n, central = NULL, sd=NULL, trend = "no trend", amp = .6,
@@ -85,6 +95,8 @@ sim_exposure <- function(n, central = NULL, sd=NULL, trend = "no trend", amp = .
 
 
 #' Pull smoothed Chicago NMMAPS mortality data
+#'
+#'
 #'
 #' @examples
 #' custom_baseline(n = 5)

### 2016-09-03 R/customizable_functions.R
Fix build error-- add title in documentation for `sim_outcome`
Turns out that all exported functions need a title line in their documentation. Adding that resolved the build error we were getting about documentation.

@@ -110,6 +110,9 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
   }
   return(lambda)
 }
+
+#' Simulate outcome
+#'
 #' @param cust_args A list of arguments and their values used in the user-specified custom functions
 #'
 #' @examples
@@ -119,7 +122,6 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
 #' cust_lambda_args = list(exposure = testexp$x, rr=1.02, constant = 4))
 #'
 #' @export
-
 sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                         amp = .6, rr = 1.01, start.date="2000-01-01",
                         cust_base_func = NULL, cust_lambda_func = NULL,

### 2016-09-03 R/customizable_functions.R
Fix error in vignette build
`baseline` is a dataframe, not a vector, now since we're passing date through with it, while `create_lambda` needs a vector for `baseline`. This change in the code puts in just the baseline values from the `baseline` dataframe into `create_lambda`

@@ -137,7 +137,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                                 average_outcome = average_outcome,
                                 trend = trend,
                                 amp = amp)
-    lambda <- create_lambda(baseline = baseline,
+    lambda <- create_lambda(baseline = baseline$baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- rpois(n = nrow(exposure), lambda = lambda)

### 2016-09-08 writing/DraftOne.Rmd
Make bindings visible

--- 2016-07-08 DraftOne.Rmd
@@
 
+##Stuff from .R Files
+
+```{r}
+sim_df2 <- function(n, central, amp, exposure_type, average_outcome, trend, rr, start.date = "2000-01-01"){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by = 1, length.out = n)
+  exposure = sim_exposure(n = n, amp = amp, central = central,
+                          exposure_type = exposure_type)
+
+  df <- data.frame(date, exposure) %>%
+    dplyr:: mutate(outcome = sim_outcome(exposure=exposure, average_outcome=average_outcome, trend = trend,
+                                        amp = amp, rr=rr, start.date=start.date))
+  return(df)
+}
+```
+
 ## References

### 2016-09-08 R/assess_performance.R

@@ -34,7 +34,7 @@ mean_beta <- function(df){
 #' @export
 #'
 beta_var <- function(df){
-  var_across_betas <- var(df$Estimate)
+  var_across_betas <- stats::var(df$Estimate)
   mean_beta_var <- mean(df$Std.Error^2)
   out <- data.frame(var_across_betas, mean_beta_var)
   return(out)
@@ -169,7 +169,7 @@ power_calc <- function(varying, values, plot = FALSE, ...){
   if(plot == TRUE){
     my_plot <- ggplot2::ggplot(out, ggplot2::aes(x = x, y = power)) +
       geom_line() + theme_minimal() +
-      xlab(varying)
+      ggplot2::xlab(varying)
     print(my_plot)
   }
 

### 2016-09-08 R/customizable_functions.R

@@ -234,14 +234,14 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
     lambda <- create_lambda(baseline = baseline$baseline,
                             exposure = exposure$x,
                             rr = rr)
-    outcome <- rpois(n = nrow(exposure), lambda = lambda)
+    outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
   }
   else if (is.null(cust_lambda_func) & !is.null(cust_base_func)){
     baseline <- do.call(cust_base_func, cust_base_args)
     lambda <- create_lambda(baseline = baseline,
                             exposure = exposure$x,
                             rr = rr)
-    outcome <- rpois(n = nrow(exposure), lambda = lambda)
+    outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
   }
   else if (is.null(cust_base_func) & !is.null(cust_lambda_func)){
     baseline <- create_baseline(n = nrow(exposure),
@@ -250,13 +250,13 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                                 amp = amp)
     cust_lambda_args$baseline <- baseline
     lambda <- do.call(cust_lambda_func, cust_lambda_args)
-    outcome <- rpois(n = nrow(exposure), lambda = lambda)
+    outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
   }
   else {
     baseline <- do.call(cust_base_func, cust_base_args)
     cust_lambda_args$baseline <- baseline
     lambda <- do.call(cust_lambda_func, cust_lambda_args)
-    outcome <- rpois(n = nrow(exposure), lambda = lambda)
+    outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
   }
   df <- data.frame(date, x = exposure$x, outcome)
   return(df)

### 2016-09-08 R/fit_models.R

@@ -22,7 +22,7 @@ spline_mod <- function(df, df_year = 7){
                    family = stats::quasipoisson(link = "log"))
 
   out_1 <- summary(mod)$coef[2, ]
-  out_2 <- confint.default(mod)[2, ]
+  out_2 <- stats::confint.default(mod)[2, ]
   out <- c(out_1, out_2)
   return(out)
 }
@@ -51,22 +51,22 @@ casecross_mod <- function(df){
     df <- subset(df, stratum %in% informative.strata)
 
     if(length(informative.strata) > 1){
-      mod <- glm(outcome ~ x + stratum,
+      mod <- stats::glm(outcome ~ x + stratum,
                  data = df,
-                 family = quasipoisson(link = "log"))
+                 family = stats::quasipoisson(link = "log"))
     } else {
-      mod <- glm(outcome ~ x,
+      mod <- stats::glm(outcome ~ x,
                  data = df,
-                 family = quasipoisson(link = "log"))
+                 family = stats::quasipoisson(link = "log"))
     }
   } else {
-    mod <- glm(outcome ~ x + stratum,
+    mod <- stats::glm(outcome ~ x + stratum,
                data = df,
-               family = quasipoisson(link = "log"))
+               family = stats::quasipoisson(link = "log"))
   }
 
   out_1 <- summary(mod)$coef[2, ]
-  out_2 <- confint.default(mod)[2, ]
+  out_2 <- stats::confint.default(mod)[2, ]
   out <- c(out_1, out_2)
   return(out)
 }
@@ -96,18 +96,18 @@ crossyear_mod <- function(df){
     df <- subset(df, stratum %in% informative.strata)
 
     if(length(informative.strata) > 1){
-      mod <- glm(y ~ x + stratum,
+      mod <- stats::glm(y ~ x + stratum,
                  data = df,
-                 family = quasipoisson(link = "log"))
+                 family = stats::quasipoisson(link = "log"))
     } else {
-      mod <- glm(y ~ x,
+      mod <- stats::glm(y ~ x,
                  data = df,
-                 family = quasipoisson(link = "log"))
+                 family = stats::quasipoisson(link = "log"))
     }
   } else {
-    mod <- glm(y ~ x + stratum,
+    mod <- stats::glm(y ~ x + stratum,
                data = df,
-               family = quasipoisson(link = "log"))
+               family = stats::quasipoisson(link = "log"))
   }
 
   out_1 <- summary(mod)$coef[2, ]

### 2016-09-08 R/sim_exposure.R

@@ -146,7 +146,7 @@ bin_t <- function(n, p, trend = "no trend", amp = .01, start.date = "2000-01-01"
 binary_exposure <- function(n, p, trend = "no trend", amp,
                             start.date = "2000-01-01", custom_func = NULL, ...){
   t <- bin_t(n=n, p=p, trend = trend, amp = amp, start.date = start.date, custom_func = custom_func)
-  x <- rbinom(length(t), size = 1, prob = t)
+  x <- stats::rbinom(length(t), size = 1, prob = t)
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
   df <- data.frame(date, x)
@@ -174,7 +174,7 @@ continuous_exposure <- function(n, mu, sd=1, trend = "no trend", amp = .6, start
   date <- seq(from = start.date, by = 1, length.out = n)
   t <- calc_t(n, trend, amp, start.date, ...)
   newmu <- mu * t
-  x <- rnorm(n, newmu, sd)
+  x <- stats::rnorm(n, newmu, sd)
   df <- data.frame(date, x)
   return(df)
 }

### 2016-09-14 R/customizable_functions.R
Margins and @s for sim_exposure and customizable_functions

@@ -1,10 +1,11 @@
 #' Pull exposure series from data set
 #'
-#' By default, this function pulls exposure data from the Chicago NMMAPS data set in the dlnm
-#' package.  The user may specify a different data set from which to pull exposure
-#' values.
+#' By default, this function pulls exposure data from the Chicago NMMAPS data
+#' set in the dlnm package.  The user may specify a different data set from
+#' which to pull exposure values.
 #'
-#' @param n A numeric value specifying the number of days for which to obtain an exposure value.
+#' @param n A numeric value specifying the number of days for which to obtain an
+#'    exposure value.
 #' @param df Data frame from which to pull exposure values.
 #' @param metric A character string specifying the desired exposure metric.
 #'    Options are:
@@ -14,8 +15,9 @@
 #'      \item"rhum"
 #'      \item"pm10"
 #'      \item"o3"}
-#' @param start.date A date of the format "yyyy-mm-dd" from which to begin pulling exposure
-#' values.  Dates in the Chicago NMMAPS data set are from 1987-01-01 to 2000-12-31.
+#' @param start.date A date of the format "yyyy-mm-dd" from which to begin
+#'    pulling exposure values. Dates in the Chicago NMMAPS data set are from
+#'    1987-01-01 to 2000-12-31.
 #'
 #' @return A numeric vector of exposure values
 #'
@@ -23,7 +25,8 @@
 #' custom_exposure(n = 5, metric = "temp", start.date = "2000-01-01")
 #'
 #' @export
-custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp", start.date = NULL){
+custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
+                            start.date = NULL){
   if(!is.null(start.date)){
     date1 <- which(df$date==start.date)
     enddate <- date1+n
@@ -37,11 +40,13 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp", start.
 
 #' Simulate random series of exposure values
 #'
-#' This function simulates binary or continuous exposure values with or without seasonal trends.
-#' It also allows for a custom function for exposure trend.
+#' This function simulates binary or continuous exposure values with or without
+#' seasonal trends. It also allows for a custom function for exposure trend.
 #'
-#' @param cust_exp_function The name of a function from which to generate custom exposure values
-#' @param cust_exp_args A list of arguments used in the user-specified custom function
[…]

### 2016-09-14 R/sim_exposure.R

@@ -88,8 +88,8 @@ calc_t <- function(n, trend = "no trend", amp = .6, custom_func = NULL, ...){
 #'
 #' @export
 #'
-bin_t <- function(n, p, trend = "no trend", amp = .01, start.date = "2000-01-01",
-                  custom_func = NULL,...){
+bin_t <- function(n, p, trend = "no trend", amp = .01,
+                  start.date = "2000-01-01", custom_func = NULL,...){
   day <- c(1:n)
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
@@ -138,14 +138,16 @@ bin_t <- function(n, p, trend = "no trend", amp = .01, start.date = "2000-01-01"
 #' @return A data frame with the dates and daily exposure values from n days
 #'
 #' @examples
-#' binary_exposure(n = 5, p = 0.1, trend = "cos1", amp = .02, start.date = "2001-02-01")
+#' binary_exposure(n = 5, p = 0.1, trend = "cos1", amp = .02,
+#'                 start.date = "2001-02-01")
 #'
 #'
 #' @export
 #'
 binary_exposure <- function(n, p, trend = "no trend", amp,
                             start.date = "2000-01-01", custom_func = NULL, ...){
-  t <- bin_t(n=n, p=p, trend = trend, amp = amp, start.date = start.date, custom_func = custom_func)
+  t <- bin_t(n=n, p=p, trend = trend, amp = amp, start.date = start.date,
+             custom_func = custom_func)
   x <- stats::rbinom(length(t), size = 1, prob = t)
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
@@ -159,7 +161,8 @@ binary_exposure <- function(n, p, trend = "no trend", amp,
 #' seasonal trend.
 #'
 #' @param mu A numeric value giving the mean exposure.
-#' @param sd A numeric value giving the standard deviation of the exposure values around the trend line.
+#' @param sd A numeric value giving the standard deviation of the exposure
+#'    values around the trend line.
 #' @inheritParams calc_t
 #'
 #' @return A data frame with the dates and daily exposure values from n days
@@ -168,7 +171,8 @@ binary_exposure <- function(n, p, trend = "no trend", amp,
 #' continuous_exposure(n = 5, mu = 100, sd = 10, trend = "cos1")
 #'
 #' @export
[…]

### 2016-09-14 writing/DraftOne.Rmd
fix examples in fit_models, delete rep_sims

@@
 
+```{r}
+#' Run simulations
+#'
+#' This function runs many simulations for a given dataset and model.
+#'
+#' @param n_sims A numeric value specifying the number of repetitions of the
+#'    simulation to run
+#' @param model A character string specifying which model to use.  Options
+#'    include "spline_mod" and "casecross_mod"
+#' @param n A numeric value specifying the number of days for which to simulate
+#'    data
+#' @param rr A numeric value specifying the relative risk for each 1-unit
+#'    increase in exposure
+#' @param x_type A character string specifying the type of exposure data.
+#'    Options are "binary" and "continuous".
+#' @inheritParams continuous_exposure
+#' @inheritParams binary_exposure
+#'
+#' @return A data frame with the summaries of the estimate of the log relative
+#'    risk for each simulation run
+#'
+#' @examples
+#' rep_sims(n_sims = 3, model = "spline_mod")
+#'
+#' @export
+#'
+rep_sims <- function(n_sims, model, n = 5 * 365, rr = 1.01,
+                     x_type = "continuous", mu = 10, sd = 2, lambda = 100){
+  out <- replicate(n_sims, eval(call(model,
+                                     sim_data(n = n, rr = rr,
+                                              x_type = x_type,
+                                              mu = mu, sd = sd,
+                                              lambda = lambda))))
+  out <- as.data.frame(t(out))
+  names(out) <- c("est", "se", "t", "p", "lower_ci", "upper_ci")
+  return(out)
+}
+
+```
+
 ## References

### 2016-09-14 R/fit_models.R

@@ -9,8 +9,10 @@
 #'    exposure
 #'
 #' @examples
-#' df <- constant_baseline(n = 600, lambda = 100)
-#' spline_mod(df)
+#' exp <- sim_exposure(n = 500, central = 100, sd = 10, trend = "cos1",
+#'                     amp = .6, exposure_type = "continuous")
+#' out <- sim_outcome(exposure = exp, average_outcome = 22, rr = 1.01)
+#' spline_mod(df = out)
 #'
 #' @export
 #'
@@ -37,7 +39,7 @@ spline_mod <- function(df, df_year = 7){
 #'    exposure
 #'
 #' @examples
-#' casecross_mod(df)
+#' casecross_mod(df = out)
 #'
 #' @export
 #'
@@ -81,7 +83,7 @@ casecross_mod <- function(df){
 #'    exposure
 #'
 #' @examples
-#' crossyear_mod(df)
+#' crossyear_mod(df = out)
 #'
 #' @export
 #'
@@ -113,42 +115,5 @@ crossyear_mod <- function(df){
   out_1 <- summary(mod)$coef[2, ]
   out_2 <- confint.default(mod)[2, ]
   out <- c(out_1, out_2)
-  return(out)
-}
-
-#' Run simulations
-#'
-#' This function runs many simulations for a given dataset and model.
-#'
-#' @param n_sims A numeric value specifying the number of repetitions of the
-#'    simulation to run
-#' @param model A character string specifying which model to use.  Options
-#'    include "spline_mod" and "casecross_mod"
-#' @param n A numeric value specifying the number of days for which to simulate
-#'    data
-#' @param rr A numeric value specifying the relative risk for each 1-unit
-#'    increase in exposure
-#' @param x_type A character string specifying the type of exposure data.
-#'    Options are "binary" and "continuous".
-#' @inheritParams continuous_exposure
-#' @inheritParams binary_exposure
-#'
-#' @return A data frame with the summaries of the estimate of the log relative
-#'    risk for each simulation run
-#'
-#' @examples
-#' rep_sims(n_sims = 3, model = "spline_mod")
-#'
-#' @export
-#'
-rep_sims <- function(n_sims, model, n = 5 * 365, rr = 1.01,
-                     x_type = "continuous", mu = 10, sd = 2, lambda = 100){
-  out <- replicate(n_sims, eval(call(model,
-                                     sim_data(n = n, rr = rr,
-                                              x_type = x_type,
[…]

### 2016-09-15 R/assess_performance.R
Revert to before previous commit, rechange things I wanted to keep, add to vignette

@@ -1,17 +1,15 @@
 #' Average Estimated Coefficient
 #'
-#' This function gives the mean value of the \eqn{\hat{\beta}}s and the mean
-#' estimated relative risk over the n simulations.
+#' This function gives the mean value of the \eqn{\hat{\beta}}s and the mean estimated
+#' relative risk over the n simulations.
 #'
-#' @param df A data frame of replicated simulations which must include a column
-#'    titled "Estimate"
+#' @param df A data frame of replicated simulations which must include a column titled "Estimate"
 #'
 #' @return A data frame with the mean estimated coefficient and mean estimated
 #'    relative risk
 #'
 #' @examples
-#' mod <- spline_mod(df = out)
-#' mean_beta(df=mod)
+#' mean_beta(df)
 #'
 #' @export
 #'

### 2016-09-15 R/customizable_functions.R

@@ -148,19 +148,22 @@ custom_baseline <- function(n, df = dlnm::chicagoNMMAPS, average_outcome = NA,
 #'
 #' @examples
 #' create_baseline(n = 5, average_outcome = 22, trend = "linear")
-#' create_baseline(n = 5, average_outcome = NA, trend = "no trend",
-#'                 custom_base_func = custom_baseline,
-#'                 cust_base_args = list(outcome_type = "death"))
+#' create_baseline(n = 5, average_outcome = NA, trend = NA, amp = NA,
+#'                 custom_func = "custom_baseline", outcome_type = "death")
 #'
 #' @export
 #'
-create_baseline <- function(n, average_outcome, trend = "no trend", amp = .6,
-                            cust_base_func = NULL, cust_base_args = list()){
+create_baseline <- function(n, average_outcome, trend, amp,
+                            cust_base_func = NULL, ...){
   if(is.null(cust_base_func)){
     lambda <- average_outcome
     baseline <- sim_baseline(n=n, lambda=lambda, trend=trend, amp=amp)
-  } else if(!is.null(cust_base_func)){
-    baseline <- do.call(cust_base_func, cust_base_args)
+  } else {
+    arguments <- list(...)
+    arguments$n <- n
+    arguments$average_outcome <- average_outcome
+    arguments$trend <- trend
+    baseline <- do.call(cust_base_func, arguments)
   }
   return(baseline)
 }
@@ -183,11 +186,7 @@ create_baseline <- function(n, average_outcome, trend = "no trend", amp = .6,
 #' @return A numeric vector of mean outcome values
 #'
 #' @examples
-#' base <- create_baseline(n=10, average_outcome = 22, trend = "no trend",
-#'                         amp = .6)
-#' exp <- sim_exposure(n = 10, central = 100, sd = 10, trend = "cos1",
-#'                     amp = .6, exposure_type = "continuous")
-#' create_lambda(baseline = base, exposure = exp$x, rr = 1.01)
+#' create_lambda(baseline, exposure, rr = 1.01)
 #'
 #' @export
 
@@ -216,17 +215,12 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
 #' @inheritParams create_baseline
 #' @inheritParams create_lambda
 #'
-#' @return A data frame with one column for date, one column for exposure values
-#'    called "x", and one column for outcome values called "outcome"
+#' @return
 #'
[…]

### 2016-09-15 R/assess_performance.R
Same commit message as last

@@ -1,15 +1,17 @@
 #' Average Estimated Coefficient
 #'
-#' This function gives the mean value of the \eqn{\hat{\beta}}s and the mean estimated
-#' relative risk over the n simulations.
+#' This function gives the mean value of the \eqn{\hat{\beta}}s and the mean
+#' estimated relative risk over the n simulations.
 #'
-#' @param df A data frame of replicated simulations which must include a column titled "Estimate"
+#' @param df A data frame of replicated simulations which must include a column
+#'    titled "Estimate"
 #'
 #' @return A data frame with the mean estimated coefficient and mean estimated
 #'    relative risk
 #'
 #' @examples
-#' mean_beta(df)
+#' mod <- spline_mod(df = out)
+#' mean_beta(df=mod)
 #'
 #' @export
 #'

### 2016-09-15 R/customizable_functions.R

@@ -285,6 +285,8 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #'             exposure_amp = .6, average_outcome = 22,
 #'             outcome_trend = "no trend", outcome_amp = .6, rr = 1.01)
 #'
+#' @export
+#'
 create_sims <- function(n_reps, n, central, sd, exposure_type, exposure_trend,
                         exposure_amp, average_outcome, outcome_trend,
                         outcome_amp, rr, start.date = "2000-01-01",
@@ -311,6 +313,18 @@ create_sims <- function(n_reps, n, central, sd, exposure_type, exposure_trend,
 #'    called "x" and "outcome"
 #' @param model A character string specifying model to be used. Choices are
 #'    "spline" and "casecrossover"
+#' @inheritParams spline_mod
+#'
+#' @return A data frame in which each row includes an estimate of beta hat,
+#'    standard error, t-value, p-value, and 2.5% and 97.5% confidence bounds
+#'    for each repetition of the simulation
+#'
+#' @examples
+#' sims <- create_sims(n_reps=10, n=50, central = 100, sd = 10,
+#'             exposure_type="continuous", exposure_trend = "cos1",
+#'             exposure_amp = .6, average_outcome = 22,
+#'             outcome_trend = "no trend", outcome_amp = .6, rr = 1.01)
+#' fit_mods(outcome = sims, model = "spline")
 #'
 #' @export
 fit_mods <- function(outcome, model, df_year = 7){
@@ -326,10 +340,16 @@ fit_mods <- function(outcome, model, df_year = 7){
   return(datframe)
 }
 
-#' Wrapper to do Everything
+#' Simulate data, fit models, and assess models
 #'
 #' This function generates exposures and outcomes, fits models, and evaluates
-#' them for many repetitions
+#' them for many simulation repetitions
+#'
+#' @inheritParams create_sims
+#' @inheritParams sim_outcome
+#'
+#' @return A list object with summaries of each model fitted on the simulated
+#'    data sets and measures of model evaluation including coverage and power
 #'
 #' @examples
 #' eesim(n_reps = 3, n = 50, central = 100, sd = 10,
@@ -358,7 +378,7 @@ eesim <- function(n_reps, n, central, sd, exposure_type, exposure_trend,
                           cust_base_args = NULL, cust_lambda_args = NULL)
   mods <- fit_mods(datasims, model, df_year)
[…]

### 2016-09-15 R/customizable_functions.R
Add baseline$baseline lots of places, fix baseline issue

@@ -192,7 +192,7 @@ create_baseline <- function(n, average_outcome, trend, amp,
 
 create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
   if(is.null(cust_lambda_func)){
-    log_lambda <- log(baseline) + log(rr) * exposure
+    log_lambda <- log(baseline$baseline) + log(rr) * exposure
     lambda <- exp(log_lambda)
   } else {
     arguments <- list(...)
@@ -219,7 +219,7 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
 #'
 #' @examples
 #' sim_outcome(exposure, cust_base_func = custombase,
-#' cust_base_args = list(n=nrow(exposure), slope = .2, intercept = 55))
+#'    cust_base_args = list(n=nrow(exposure), slope = .2, intercept = 55))
 #' sim_outcome(exposure, p, average_outcome = 22, cust_lambda_func =
 #'             customlambda, cust_lambda_args = list(exposure = testexp$x,
 #'             rr=1.02, constant = 4))
@@ -247,7 +247,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
   }
   else if (is.null(cust_lambda_func) & !is.null(cust_base_func)){
     baseline <- do.call(cust_base_func, cust_base_args)
-    lambda <- create_lambda(baseline = baseline,
+    lambda <- create_lambda(baseline = baseline$baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
@@ -257,7 +257,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                                 average_outcome = average_outcome,
                                 trend = trend,
                                 amp = amp)
-    cust_lambda_args$baseline <- baseline
+    cust_lambda_args$baseline <- baseline$baseline
     lambda <- do.call(cust_lambda_func, cust_lambda_args)
     outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
   }

### 2016-09-15 R/customizable_functions.R
One to many baseline$baseline - only need when not using a custom baseline function

@@ -247,7 +247,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
   }
   else if (is.null(cust_lambda_func) & !is.null(cust_base_func)){
     baseline <- do.call(cust_base_func, cust_base_args)
-    lambda <- create_lambda(baseline = baseline$baseline,
+    lambda <- create_lambda(baseline = baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)

### 2016-09-20 R/customizable_functions.R
Check examples (weird baseline problems), add plot to vignette

@@ -148,7 +148,7 @@ custom_baseline <- function(n, df = dlnm::chicagoNMMAPS, average_outcome = NA,
 #'
 #' @examples
 #' create_baseline(n = 5, average_outcome = 22, trend = "linear")
-#' create_baseline(n = 5, average_outcome = NA, trend = NA, amp = NA,
+#' create_baseline(n = 5, average_outcome = NA, trend = NA,
 #'                 custom_func = "custom_baseline", outcome_type = "death")
 #'
 #' @export
@@ -186,7 +186,10 @@ create_baseline <- function(n, average_outcome, trend, amp,
 #' @return A numeric vector of mean outcome values
 #'
 #' @examples
-#' create_lambda(baseline, exposure, rr = 1.01)
+#' base <- create_baseline(n = 5, average_outcome = 22, trend = "linear")
+#' exp <- sim_exposure(n = 5, central = 100, sd = 10, amp = .6,
+#'                     exposure_type = "continuous")
+#' create_lambda(baseline = base, exposure = exp, rr = 1.01)
 #'
 #' @export
 
@@ -218,11 +221,9 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
 #' @return
 #'
 #' @examples
-#' sim_outcome(exposure, cust_base_func = custombase,
-#'    cust_base_args = list(n=nrow(exposure), slope = .2, intercept = 55))
-#' sim_outcome(exposure, p, average_outcome = 22, cust_lambda_func =
-#'             customlambda, cust_lambda_args = list(exposure = testexp$x,
-#'             rr=1.02, constant = 4))
+#' exp <- sim_exposure(n = 5, central = 100, sd = 10, amp = .6,
+#'                     exposure_type = "continuous")
+#' sim_outcome(exposure = exp, average_outcome = 22, trend = "linear")
 #'
 #' @export
 sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
@@ -240,7 +241,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                                 average_outcome = average_outcome,
                                 trend = trend,
                                 amp = amp)
-    lambda <- create_lambda(baseline = baseline$baseline,
+    lambda <- create_lambda(baseline = baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)

### 2016-09-26 R/customizable_functions.R
Add outcome and models to vignette

@@ -63,6 +63,8 @@ sim_exposure <- function(n, central = NULL, sd=NULL, trend = "no trend",
                          amp = .6, exposure_type = NULL,
                          start.date = "2001-01-01", cust_exp_func = NULL,
                          cust_exp_args = NULL){
+  start.date <- as.Date(start.date)
+  date <- seq(from = start.date, by = 1, length.out = n)
   if(is.null(cust_exp_args)){
     arguments <- vector(mode = "list")
   } else {
@@ -85,8 +87,6 @@ sim_exposure <- function(n, central = NULL, sd=NULL, trend = "no trend",
     if (!is.null(central)){
       arguments$central <- central
     }
-    start.date <- as.Date(start.date)
-    date <- seq(from = start.date, by = 1, length.out = n)
     x <- do.call(cust_exp_func, arguments)
     exposure <- data.frame(date, x)
   } else {

### 2016-09-26 R/assess_performance.R
Add model performance to vignette

@@ -27,14 +27,14 @@ mean_beta <- function(df){
 #'
 #' Standard Deviation of Estimated Coefficients
 #'
-#' This function gives the variance of the point estimates of $\hat{\beta}$ over
-#' the n simulations and the mean of the variances of each $\hat{\beta}$.
+#' This function gives the variance of the point estimates of beta hat over
+#' the n simulations and the mean of the variances of each beta hat.
 #'
 #' @param df A data frame of replicated simulations which must include columns
 #'    titled "Estimate" and "Std.Error".
 #'
-#' @return A data frame of the variance across all values of \eqn{\hat{\beta}}
-#'    and the mean variance of the $\hat{\beta}$s
+#' @return A data frame of the variance across all values of beta hat
+#'    and the mean variance of the beta hats
 #'
 #' @examples
 #' beta_var(df)

### 2016-09-26 writing/custom_examples.Rmd
Add customization examples to vignette

--- 2016-08-23 custom_examples.Rmd
@@
 binary_exposure(n=5, p = .2, y = .1, trend = "custom", custom_func = sintrend)
-plot(binary_exposure(n=400, p = .2, y = .1, trend = "custom", custom_func = sintrend))
+plot(binary_exposure(n=400, p = .2, y = .1, amp = .01, trend = "custom", custom_func = sintrend))
 

### 2016-09-26 R/customizable_functions.R
Make sure `start.date` passes through to `std_exposure` when running `sim_exposure`

@@ -82,6 +82,7 @@ sim_exposure <- function(n, central = NULL, sd=NULL, trend = "no trend",
     arguments$amp <- amp
     arguments$exposure_type <- exposure_type
     arguments$sd <- sd
+    arguments$start.date <- start.date
     exposure <- do.call("std_exposure", arguments)
   } else if (!(is.null(cust_exp_func))){
     if (!is.null(central)){

### 2016-09-29 R/customizable_functions.R
Add custom examples

@@ -361,9 +361,10 @@ fit_mods <- function(outcome, model, df_year = 7){
 #'
 #' @export
 #'
-eesim <- function(n_reps, n, central, sd, exposure_type, exposure_trend,
-                  exposure_amp, average_outcome, outcome_trend = "no trend",
-                  outcome_amp, rr, start.date = "2000-01-01",
+eesim <- function(n_reps, n, central, sd, exposure_type, exposure_trend = NULL,
+                  exposure_amp = NULL, average_outcome = NULL,
+                  outcome_trend = "no trend",
+                  outcome_amp = NULL, rr, start.date = "2000-01-01",
                   cust_exp_func = NULL, cust_exp_args = NULL,
                   cust_base_func = NULL, cust_lambda_func = NULL,
                   cust_base_args = NULL, cust_lambda_args = NULL, model,

### 2016-10-26 R/assess_performance.R
Make examples self-sufficient

@@ -36,7 +36,15 @@ mean_beta <- function(df){
 #'    and the mean variance of the beta hats
 #'
 #' @examples
-#' beta_var(df)
+#' sims <- create_sims(n_reps = 10, n = 600, central = 100,
+#'                     sd = 10, exposure_type = "continuous",
+#'                     exposure_trend = "cos1",
+#'                     exposure_amp = 0.6,
+#'                     average_outcome = 20,
+#'                     outcome_trend = "no trend",
+#'                     rr = 1.01)
+#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' beta_var(fits)
 #'
 #' @export
 beta_var <- function(df){
@@ -57,7 +65,15 @@ beta_var <- function(df){
 #'    simulations
 #'
 #' @examples
-#' beta_bias(df, true_rr = 1.02)
+#' sims <- create_sims(n_reps = 10, n = 600, central = 100,
+#'                     sd = 10, exposure_type = "continuous",
+#'                     exposure_trend = "cos1",
+#'                     exposure_amp = 0.6,
+#'                     average_outcome = 20,
+#'                     outcome_trend = "no trend",
+#'                     rr = 1.01)
+#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' beta_bias(fits, true_rr = 1.02)
 #'
 #' @export
 beta_bias <- function(df, true_rr){
@@ -97,7 +113,15 @@ coverage_beta <- function(df, true_rr){
 #' @return Power at the 5% significance level
 #'
 #' @examples
-#' power_beta(df)
+#' sims <- create_sims(n_reps = 10, n = 600, central = 100,
+#'                     sd = 10, exposure_type = "continuous",
+#'                     exposure_trend = "cos1",
+#'                     exposure_amp = 0.6,
+#'                     average_outcome = 20,
+#'                     outcome_trend = "no trend",
+#'                     rr = 1.01)
+#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' power_beta(fits)
 #'
 #' @export
 power_beta <- function(df){

### 2016-10-26 R/customizable_functions.R
Inherit parameters from `continuous_exposure` for `custom_exposure`

@@ -48,6 +48,7 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
 #' @param cust_exp_args A list of arguments used in the user-specified custom
 #'    function
 #' @inheritParams std_exposure
+#' @inheritParams continuous_exposure
 #'
 #' @return A data frame with two columns: date and exposure values
 #'
@@ -59,7 +60,7 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
 #'                     cust_exp_args = list(metric = "temp"))
 #'
 #' @export
-sim_exposure <- function(n, central = NULL, sd=NULL, trend = "no trend",
+sim_exposure <- function(n, central = NULL, sd = NULL, trend = "no trend",
                          amp = .6, exposure_type = NULL,
                          start.date = "2001-01-01", cust_exp_func = NULL,
                          cust_exp_args = NULL){

### 2016-10-26 R/sim_exposure.R
Take dots out of `std_exposure`

@@ -202,16 +202,15 @@ continuous_exposure <- function(n, mu, sd=1, trend = "no trend", amp = .6,
 #' @export
 std_exposure <- function(n, central, sd = NULL, trend = "no trend",
                          exposure_type = "binary", amp,
-                         start.date = "2000-01-01", ...){
+                         start.date = "2000-01-01"){
   if(exposure_type=="binary"){
     p <- central
-    df <- binary_exposure(n=n, p=p, trend=trend, amp=amp, start.date=start.date,
-                          ...)
+    df <- binary_exposure(n=n, p=p, trend=trend, amp=amp, start.date=start.date)
   }
   else if(exposure_type == "continuous"){
     mu <- central
     df <- continuous_exposure(n=n, mu=mu, sd =sd, trend=trend, amp=amp,
-                              start.date=start.date, ...)
+                              start.date=start.date)
   }
   return(df)
 }

### 2016-10-26 R/sim_exposure.R
Undo last commit

@@ -202,15 +202,16 @@ continuous_exposure <- function(n, mu, sd=1, trend = "no trend", amp = .6,
 #' @export
 std_exposure <- function(n, central, sd = NULL, trend = "no trend",
                          exposure_type = "binary", amp,
-                         start.date = "2000-01-01"){
+                         start.date = "2000-01-01", ...){
   if(exposure_type=="binary"){
     p <- central
-    df <- binary_exposure(n=n, p=p, trend=trend, amp=amp, start.date=start.date)
+    df <- binary_exposure(n=n, p=p, trend=trend, amp=amp, start.date=start.date,
+                          ...)
   }
   else if(exposure_type == "continuous"){
     mu <- central
     df <- continuous_exposure(n=n, mu=mu, sd =sd, trend=trend, amp=amp,
-                              start.date=start.date)
+                              start.date=start.date, ...)
   }
   return(df)
 }

### 2016-10-26 R/sim_exposure.R
Don't call `amp` error for `bin_t` for "no trend"

@@ -94,10 +94,10 @@ bin_t <- function(n, p, trend = "no trend", amp = .01,
   date <- seq(from = start.date, by = 1, length.out = n)
   if (trend == "monthly"){
   }
-  else if (p > .5 & amp >1-p){
+  else if (p > .5 & amp >1-p & !(trend == "no trend")){
     stop(paste0("For p>.5, amp must be between 0 and 1-p."))
   }
-  else if (p<.5 & amp >p){
+  else if (p<.5 & amp >p & !(trend == "no trend")){
     stop(paste0("For p<.5, amp must be between 0 and p."))
   }
   if (trend == "cos1"){

### 2016-10-26 R/customizable_functions.R
Update documentation to inherit parameters

@@ -43,12 +43,13 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
 #' This function simulates binary or continuous exposure values with or without
 #' seasonal trends. It also allows for a custom function for exposure trend.
 #'
-#' @param cust_exp_function The name of a function from which to generate custom
+#' @param cust_exp_func The name of a function from which to generate custom
 #'    exposure values
 #' @param cust_exp_args A list of arguments used in the user-specified custom
 #'    function
 #' @inheritParams std_exposure
 #' @inheritParams continuous_exposure
+#' @inheritParams calc_t
 #'
 #' @return A data frame with two columns: date and exposure values
 #'

### 2016-10-26 R/sim_exposure.R
Undo default amp in `std_exposure`

@@ -205,7 +205,7 @@ continuous_exposure <- function(n, mu, sd = 1, trend = "no trend", amp = .6,
 #'
 #' @export
 std_exposure <- function(n, central, sd = NULL, trend = "no trend",
-                         exposure_type = "binary", amp = 0.6,
+                         exposure_type = "binary", amp,
                          start.date = "2000-01-01", ...){
   if(exposure_type=="binary"){
     p <- central

### 2016-11-16 R/plots.R
Add calendar plot to plots.R

@@ -21,6 +21,8 @@
 #'           model = "spline", df_year = 1)
 #' coverage_plot(ex_sim[[1]], true_param = 1.10)
 #'
+#' @export
+#'
 coverage_plot <- function(summarystats, true_param){
 summarystats %>%
   arrange(Estimate) %>%
@@ -41,7 +43,56 @@ summarystats %>%
   scale_x_discrete(breaks = NULL) + xlab("")
 }
 #'
+#' Calendar Plot
 #'
+#' This plot displays daily values of a binary or continuous variable on a
+#' calendar via color.
+#'
+#' @param df Data frame with one column for date with entries in the format
+#' "yyyy-mm-dd" and one column for the daily values of the variable.
+#' @param type Character string specifying whether the exposure is continuous or
+#' discrete
+#' @param labels Vector of character strings naming the levels of a discrete
+#' variable
+#'
+#' @examples
+#' testdat <- sim_exposure(n = 1000, central = 0.1,
+#'            exposure_type = "binary")
+#' testdat$x[c(89,101,367,500,502,598,678,700,895)] <- 3
+#' calendar_plot(testdat, type = "discrete", labels = c("no", "yes", "maybe"))
+#'
+#'
+calendar_plot <- function(df, type = "continuous", labels = NULL){
+  names(df) <- c("date", "x")
+  if(type=="continuous"){
+    Exposure<- df$x
+  }
+  else if(type=="discrete"){
+    Exposure <- factor(df$x, levels = levels(factor(df$x)), labels = labels)
+  }
+  plot <- df %>%
+  mutate(Weekday = lubridate::wday(date),
+         Month = lubridate::month(date, label = TRUE),
+         Year = lubridate::year(date),
+         Exposure) %>%
+  group_by(Year, Month) %>%
+  dplyr::mutate(saturday = lag(Weekday) == 7,
+                saturday = ifelse(is.na(saturday), 0, saturday),
+                Week = 1 + cumsum(saturday)) %>%
+  ungroup() %>%
+  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) +
+  geom_tile(colour = "white") +
+  facet_grid(Year ~ Month, scales = "free")
+  if(type=="continuous"){
+    newplot <- plot + scale_fill_gradientn(colours = viridis(256)) +
+      scale_y_reverse() + theme_void()
+  }
+  else if(type=="discrete"){
[…]

### 2016-11-17 README.Rmd
Add README. Close #12.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r, echo = FALSE, message = FALSE}
library(eesim)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridisLite)
library(gridExtra)
```

## Overview of package

This package allows you to simulate time series of environmental health data and perform simulation-based power analyses and other measures of model performance. The package includes four main parts:

1. Generation of exposure data; 
2. Generation of outcome data;
3. Fitting models to simulated data; and 
4. Evaluating model performance on simulated data.

The user has the option to customize different aspects of the simulation at each of these steps.

The package creates time series that are relevant for environmental epidemiology studies of ambient exposures (e.g., studies of acute mortality risks associated with daily air pollution concentration, daily temperature, or occurance of a community-wide extreme event like a heat wave).

## Basic example of using the package

The main function of this package is the `eesim` function. You can use the `eesim` function to conduct all four steps of the simulation process at once (simulate exposure data, simulate outcome data, fit models to simulated data, and evaluate model performance).

The `eesim` function requires [required inputs]. The function returns a list with two elements. The first element gives simulation-specific results: the estimated effect, standard error, t- and p-values, and upper and lower 95% confidence bounds when a model was applied to each set of simulated data. The second element gives some measures of model assessment, assessed over all simulations, including the mean beta and relative risk estimates across simulations.

For example, you can use the following call to (1) generate 10 observations of a continuous exposure with mean 100 and standard deviation 10, with a seasonal trend in expected value; (2) generate 10 associated outcome values, where the outcome has an average value of 20 and a relative risk of 1.10 per one-unit increase in the exposure; (3) fit a generalized linear model that controls for long-term and seasonal trends with a natural cubic spline with 1 degree of freedom per year; and (4) evaluate the performace of that model on the simulated data:

[To do-- increase `n` and `n_reps`, but wait until we've edited the vignette, because that will increase time to render. We could change `df` then, too (and maybe change to case-crossover? That may be easier for a first example)]

```{r, warning = FALSE, message = FALSE}
ex_sim <- eesim(n_reps = 100, n = 10, central = 100, sd = 10,
                exposure_type = "continuous", 
                average_outcome = 20, rr = 1.10,
                model = "spline", df_year = 1)
head(ex_sim[[1]])
ex_sim[[2]]
```

The first element of the returned object can be used to explore the behavior of individual simulations. For example, to look at the relative risk point estimate and 95% confidence interval from each of the 100 simulations, as well as which 95% confidence intervals include the true relative risk using the following call:

```{r fig.width = 4, fig.height = 5, fig.align = "center"}
ex_sim[[1]] %>%
  arrange(Estimate) %>%
  mutate(index = 1:n(),
         rr = exp(Estimate),
         lower_rr = exp(lower_ci),
         upper_rr = exp(upper_ci),
         includes_true = lower_rr < 1.10 & 1.10 < upper_rr) %>%
  ggplot(aes(x = index, y = rr, color = includes_true)) + 
  coord_flip() + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_rr, ymax = upper_rr)) + 
  geom_hline(yintercept = 1.10, linetype = 2) + 
  scale_color_manual(values = c("red", "darkgray")) +
  theme(legend.position="none",
        panel.background = element_rect(fill='white', colour='white')) + 
  ylab("Relative risk") +
  scale_x_discrete(breaks = NULL) + xlab("") 
```

The second element of the returned object gives the following overall summaries of model performance across all simulations:

```{r echo = FALSE}
eesim_overall_output <- data_frame(element = paste0("`",
                                                    colnames(ex_sim[[2]]),
                                                    "`"),
                                   desc = c("The mean estimated effect (log relative risk) over all simulations",
                                            "The mean relative risk over all simulations",
                                            "Variance of the point estimates of the estimated effect over all simulations",
                                            "The mean across simulations of the variances of each estimated effect",
                                            "Relative bias of the mean of the estimated coefficients",
                                            "Percent of simulations for which the 95% confidence interval estimate includes the true value",
                                            NA))
knitr::kable(eesim_overall_output, col.names = c("Variable", "Description"))
```

##Piece-by-piece breakdown of package utility

To demonstrate how the `eesim` function works, here is a breakdown of each of the four main parts: generating exposure data, generating outcome data, fitting models, and evaluating models. The helper functions used for each step are described in detail in this section.

### Generating exposure data

The first task of the package is generating exposure data. This can be dones with the `sim_exposure` function. In this function, the user can specify whether they would like the exposure data to be binary or continuous (`exposure_type`). For continuous exposure data, the user can specify the mean (`central`) and standard deviation (`sd`) of the exposure data:

```{r}
x_cont <- sim_exposure(n = 1000, central = 50, sd = 5,
                      exposure_type = "continuous") 
x_cont %>% slice(1:5)
```
```{r fig.width = 4, fig.height = 2, fig.align = "center"}
ggplot(x_cont, aes(x = date, y = x)) + geom_point(alpha = 0.2)
```

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
x_cont %>% 
  mutate(Weekday = lubridate::wday(date), 
         Month = lubridate::month(date, label = TRUE),
         Year = lubridate::year(date),
         Exposure = x) %>% 
  group_by(Year, Month) %>%
  dplyr::mutate(saturday = lag(Weekday) == 7,
                saturday = ifelse(is.na(saturday), 0, saturday),
                Week = 1 + cumsum(saturday)) %>%
  ungroup() %>%
  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
   geom_tile(colour = "white") + 
   facet_grid(Year ~ Month, scales = "free") + 
  scale_fill_gradientn(colours = viridis(256)) + 
  scale_y_reverse() + theme_void()
```

For binary exposure data, the `central` argument gives the probability of exposure on a study day:

```{r}
x_bin <- sim_exposure(n = 1000, central = 0.1,
                      exposure_type = "binary")
x_bin %>% slice(1:5)
```
```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
x_bin %>% 
  mutate(Weekday = lubridate::wday(date), 
         Month = lubridate::month(date, label = TRUE),
         Year = lubridate::year(date),
         Exposure = factor(x, levels = c(0, 1), 
                    labels = c("Not exposed", "Exposed"))) %>% 
  group_by(Year, Month) %>%
  dplyr::mutate(saturday = lag(Weekday) == 7,
                saturday = ifelse(is.na(saturday), 0, saturday),
                Week = 1 + cumsum(saturday)) %>%
  ungroup() %>%
  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
   geom_tile(colour = "white") + 
   facet_grid(Year ~ Month, scales = "free") + 
  scale_fill_manual(values = c("#414487FF", "#7AD151FF")) + 
  scale_y_reverse() + theme_void()
```

For environmental epidemiology applications, exposure data often has a seasonal trend and / or long-term trend. The default for `sim_exposure` is to simulate the exposure data without a time trend (`trend = "no trend"`). However, we have also built in several time trends from which a user can to choose to simulate exposure data with a time trend, either seasonal or long-term or both. These trend patterns differ slightly depending on whether the user is simulating binary or continuous data. [add reference for Schwartz paper where we got some of the trend equations] Below are plots of the built-in trends for continuous exposure data from which the user may choose.

```{r fig.width = 7, fig.height = 4, echo = FALSE, message = FALSE}
data.frame(day = 1:1000) %>%
  mutate(`"no trend"` = calc_t(n = 1000, trend = "no trend"),
         `"cos1"` = calc_t(n = 1000, trend = "cos1"),
         `"cos2"` = calc_t(n = 1000, trend = "cos2"),
         `"cos3"` = calc_t(n = 1000, trend = "cos3"),
         `"linear"` = calc_t(n = 1000, trend = "linear"),
         `"curvilinear"` = calc_t(n = 1000, trend = "curvilinear"),
         `"cos1linear"` = calc_t(n = 1000, trend = "cos1linear")) %>%
  gather(trend_method, trend_value, -day) %>%
  mutate(trend_method = factor(trend_method,
                               levels = c('"no trend"', 
                                          '"linear"',
                                          '"curvilinear"',
                                          '"cos1"',
                                          '"cos1linear"',
                                          '"cos2"',
                                          '"cos3"'))) %>%
  ggplot(aes(x = day, y = trend_value)) + 
  geom_line() + facet_wrap(~ trend_method, ncol = 4) + 
  theme_bw()
```

The trend options are similar for binary exposure, but exclude  "curvilinear" and "cos1linear" and include a "monthly" trend, for which the user enters a vector of 12 probabilities of exposure, one for each month.

Here is an example of generating continuous exposure data with a "cos1linear"" 
trend:

```{r, fig.width = 5, fig.height = 3, fig.align = "center"}
testexp <- sim_exposure(n=1000, central = 50, sd = 5, trend = "cos1linear",
                        amp = .6, exposure_type = "continuous")
head(testexp)
qplot(testexp$date, testexp$x)+geom_point()+coord_cartesian(ylim = c(0,110)) + labs(title = "Exposure with a cos1linear trend", x = "Date", y="Exposure")+theme_bw()
```

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
testexp %>% 
  mutate(Weekday = lubridate::wday(date), 
         Month = lubridate::month(date, label = TRUE),
         Year = lubridate::year(date),
         Exposure = x) %>% 
  group_by(Year, Month) %>%
  dplyr::mutate(saturday = lag(Weekday) == 7,
                saturday = ifelse(is.na(saturday), 0, saturday),
                Week = 1 + cumsum(saturday)) %>%
  ungroup() %>%
  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
   geom_tile(colour = "white") + 
   facet_grid(Year ~ Month, scales = "free") + 
  scale_fill_gradientn(colours = viridis(256)) + 
  scale_y_reverse() + theme_void()
```

The default amplitude for the trend is .6, but we can adjust the "amp" parameter
for a smaller or larger trend amplitude:

```{r, fig.width = 5, fig.height = 3, fig.align = "center"}
smallamp <- sim_exposure(n=1000, central = 50, sd = 5, trend = "cos1linear",
                        amp = .2, exposure_type = "continuous")
qplot(smallamp$date, smallamp$x)+geom_point()+coord_cartesian(ylim = c(0,110))+labs(title = "Cos1linear exposure with smaller amplitude", x="Date", y="Exposure") + theme_bw()
```

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
smallamp %>% 
  mutate(Weekday = lubridate::wday(date), 
         Month = lubridate::month(date, label = TRUE),
         Year = lubridate::year(date),
         Exposure = x) %>% 
  group_by(Year, Month) %>%
  dplyr::mutate(saturday = lag(Weekday) == 7,
                saturday = ifelse(is.na(saturday), 0, saturday),
                Week = 1 + cumsum(saturday)) %>%
  ungroup() %>%
  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
   geom_tile(colour = "white") + 
   facet_grid(Year ~ Month, scales = "free") + 
  scale_fill_gradientn(colours = viridis(256)) + 
  scale_y_reverse() + theme_void()
```

Here is an example of generating binary exposure data with a "monthly" trend 
starting from June 1, 2002.  The probabilities for the central parameter are for each month of the year, starting in January.

```{r, message = FALSE, fig.width = 5, fig.height = 3, fig.align = "center"}
testbin <- sim_exposure(n=1000, central = c(.1,.1,.2,.3,.4,.4,.5,.6,.5,.3,.2,.1),
                        trend = "monthly", exposure_type = "binary", 
                        start.date = "2002-06-01")
head(testbin)
qplot(testbin$date, testbin$x)+geom_point()+labs(title = "Binary exposure data with monthly trend", x="Date", y="Exposure")+theme_bw()
```

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
testbin %>% 
  mutate(Weekday = lubridate::wday(date), 
         Month = lubridate::month(date, label = TRUE),
         Year = lubridate::year(date),
         Exposure = factor(x, levels = c(0, 1), 
                    labels = c("Not exposed", "Exposed"))) %>% 
  group_by(Year, Month) %>%
  dplyr::mutate(saturday = lag(Weekday) == 7,
                saturday = ifelse(is.na(saturday), 0, saturday),
                Week = 1 + cumsum(saturday)) %>%
  ungroup() %>%
  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
   geom_tile(colour = "white") + 
   facet_grid(Year ~ Month, scales = "free") + 
  scale_fill_manual(values = c("#414487FF", "#7AD151FF")) + 
  scale_y_reverse() + theme_void()
```

The following figure shows the steps taken in simulating exposure ...

```{r echo = FALSE, fig.width = 8, fig.height = 3}
# binary_trend 
# binary_continuous
continuous_trend <- data_frame(t = calc_t(n = 1000, trend = "cos1")) %>%
  mutate(x = 100 * t,
         date = seq(from = as.Date("2000-01-01"),
                    by = 1, length.out = 1000)) %>%
  ggplot(aes(x = date, y = x)) + geom_point() + theme_minimal() + 
  ylim(c(0, 200))
continuous_simulated <- continuous_exposure(n = 1000, mu = 100,
                                            sd = 10, trend = "cos1") %>%
  ggplot(aes(x = date, y = x)) + geom_point() + theme_minimal() + 
  ylim(c(0, 200))
grid.arrange(continuous_trend, continuous_simulated, ncol = 2)
```

### Generating outcome data

Next, the eesim package uses the exposure data to generate health outcome data with or without seasonal trends. eesim creates "baseline" values for outcome based on the user-specified trend and average outcome. Next it uses the following function to relate exposure and relative risk to outcomes:
$$\lambda = e^{\log{(baseline)}+\log{(relative risk)}*exposure}$$
The values of lambda are then used to randomize the outcome values around the baseline with a Poisson($\lambda\$) distribution.

Here is an example of generating health outcome data with an upward linear trend using exposure data with a "cos1" trend:

```{r fig.width = 5, fig.height = 3, fig.align = "center"}
testexp2 <- sim_exposure(n=1000, central = 100, sd = 10, trend = "cos1", exposure_type = "continuous")
testout <- sim_outcome(exposure = testexp2, average_outcome = 22, trend = "linear", rr = 1.01)
qplot(testout$date, testout$outcome)+geom_point()+labs(title = "Health outcomes with a linear trend", x="Date", y="Outcome") + theme_bw()
```

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
testout %>% 
  mutate(Weekday = lubridate::wday(date), 
         Month = lubridate::month(date, label = TRUE),
         Year = lubridate::year(date),
         Exposure = outcome) %>% 
  group_by(Year, Month) %>%
  dplyr::mutate(saturday = lag(Weekday) == 7,
                saturday = ifelse(is.na(saturday), 0, saturday),
                Week = 1 + cumsum(saturday)) %>%
  ungroup() %>%
  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
   geom_tile(colour = "white") + 
   facet_grid(Year ~ Month, scales = "free") + 
  scale_fill_gradientn(colours = viridis(256)) + 
  scale_y_reverse() + theme_void()
```

###Fitting models

eesim next uses many generated data sets to fit statistical models to relate exposure to outcome and estimate relative risk.  The built-in model choices are a spline model and a case-crossover model. The fit_mods function outputs a data frame with estimates of the log relative risk, p-values, and upper and lower 95% confidence bounds for each simulated data set.

Here is an example of fitting the spline model with 7 degrees of freedom:

```{r}
sims <- create_sims(n_reps=10, n=100, central = 100, sd = 10,
             exposure_type="continuous", exposure_trend = "cos1",
             exposure_amp = .6, average_outcome = 22,
             outcome_trend = "no trend", outcome_amp = .6, rr = 1.01)
fits <- fit_mods(outcome = sims, model = "spline", df_year = 7)
fits
```

###Evaluating the models

Lastly, eesim evaluates model performance with several different measures.  The check_sims function takes the true relative risk as input and returns mean beta and relative risk estimates across all simulated data sets, variance of the estimates of beta, the mean of the variances of each beta hat, the relative bias of the mean of the beta hats, the percent coverage of the true beta, and the power of the test at the 5% significance level.

Here is an example of the use of the check_sims function:

```{r}
check_sims(fits, true_rr = 1.01)
```

##Using custom functions

An important feature of eesim is that the user may input custom functions for any part of the simulation process.  For example, the user may wish to generate exposure data with a custom trend, then automate the process of generating outcomes, fitting models, and evaluating performance using the built-in features of eesim. Functions the user has the option to customize within the eesim framework are exposure trend, outcome baseline, and outcome lambda. A custom model may also be used as long as returns the necessary output if the model assessmen tools in eesim are to be used.

To use custom functions within eesim, the user must input the name of the custom function as well as a list of all arguments for the custom function and their values. Many inputs for the eesim functions may no longer be necessary when a custom function is used, in which case they can simply be left out.

###Custom exposure trend

Here is an example of using a custom function to specify an exposure trend that is not built in to eesim:

```{r, fig.width = 5, fig.height = 3, fig.align = "center"}
sintrend <- function(n, mu, y){
  day <- c(1:n)
  base <- mu + y * sin(2 * pi * (day / 365))
  rnorm(n, mean = base, sd = 1)
}
customexp <- sim_exposure(n=1000, exposure_type = "continuous", cust_exp_func = sintrend, cust_exp_args = list(n=1000, mu = 75, y=3))
head(customexp)
qplot(customexp$date, customexp$x)+geom_point()+labs(title = "Exposure values with a custom trend", x="Date", y="Exposure")+theme_bw()
```

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
customexp %>% 
  mutate(Weekday = lubridate::wday(date), 
         Month = lubridate::month(date, label = TRUE),
         Year = lubridate::year(date),
         Exposure = x) %>% 
  group_by(Year, Month) %>%
  dplyr::mutate(saturday = lag(Weekday) == 7,
                saturday = ifelse(is.na(saturday), 0, saturday),
                Week = 1 + cumsum(saturday)) %>%
  ungroup() %>%
  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
   geom_tile(colour = "white") + 
   facet_grid(Year ~ Month, scales = "free") + 
  scale_fill_gradientn(colours = viridis(256)) + 
  scale_y_reverse() + theme_void()
```

This custom exposure function can be called with eesim:

```{r, warning=F, error = TRUE}
ex_sim2 <- eesim(n_reps = 3, n = 10, central = 100, sd = 10,
                exposure_type = "continuous", exposure_trend = "custom", 
                exposure_amp = .6, cust_exp_func = sintrend,
                cust_exp_args = list(n=10, mu=75,y=3),
                average_outcome = 22, rr = 1.01, 
                model = "spline", df_year = 1)
ex_sim2
```

###Custom outcomes

There are two ways to customize the simulated outcome data: creating a custom baseline for outcome values or customizing the relationship between outcome and exposure.

The outcome baseline is comprised of the values the user expects the outcomes to take on without exposure factored in. The user may write a function to specify the trend of the baseline, then use it as an input in sim\_outcome or eesim. Here is an example of creating a custom baseline function and using it in the eesim function:

```{r, warning = F}
custombase <- function(n, slope, intercept){
  day <- c(1:n)
  baseline <- day*slope + intercept
  return(baseline)
}

#Example:
custombase(n=5, slope = .3, intercept = 55)

ex_sim3 <- eesim(n_reps = 3, n = 10, central = 100, sd = 10,
                exposure_type = "continuous", exposure_trend = "cos1",
                exposure_amp = .6, average_outcome = 22, rr = 1.01, 
                cust_base_func = custombase, cust_base_args = list(n=10, slope = .5,                        intercept = 12), model = "spline", df_year = 1)
ex_sim3
```

Here is an example of creating a custom lambda, meaning a custom function relating relative risk and exposure to outcomes, and using it in eesim with the custom baseline function created above:

```{r, warning = F}
customlambda <- function(exposure, rr, constant, baseline){
  log_lambda <- log(baseline) + log(rr) * exposure + constant
  lambda <- exp(log_lambda)
  return(lambda)
}
```

### 2016-12-05 R/plots.R
Import pipe from dplyr for `coverage_plot`

@@ -21,6 +21,8 @@
 #'           model = "spline", df_year = 1)
 #' coverage_plot(ex_sim[[1]], true_param = 1.05)
 #'
+#' @importFrom dplyr %>%
+#'
 #' @export
 #'
 coverage_plot <- function(summarystats, true_param){

### 2016-12-05 R/plots.R
Use package::function syntax for `coverage_plot`

@@ -26,23 +26,23 @@
 #' @export
 #'
 coverage_plot <- function(summarystats, true_param){
-summarystats %>%
-  arrange(Estimate) %>%
-  mutate(index = 1:n(),
-         rr = exp(Estimate),
-         lower_rr = exp(lower_ci),
-         upper_rr = exp(upper_ci),
-         includes_true = lower_rr < true_param & true_param < upper_rr) %>%
-  ggplot(aes(x = index, y = rr, color = includes_true)) +
-  coord_flip() +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower_rr, ymax = upper_rr)) +
-  geom_hline(yintercept = true_param, linetype = 2) +
-  scale_color_manual(values = c("red", "darkgray")) +
-  theme(legend.position="none",
-        panel.background = element_rect(fill='white', colour='white')) +
-  ylab("Relative risk") +
-  scale_x_discrete(breaks = NULL) + xlab("")
+  summarystats %>%
+    dplyr::arrange(Estimate) %>%
+    dplyr::mutate(index = 1:n(),
+                  rr = exp(Estimate),
+                  lower_rr = exp(lower_ci),
+                  upper_rr = exp(upper_ci),
+                  includes_true = lower_rr < true_param & true_param < upper_rr) %>%
+    ggplot2::ggplot(ggplot2::aes(x = index, y = rr, color = includes_true)) +
+    ggplot2::coord_flip() +
+    ggplot2::geom_point() +
+    ggplot2::geom_errorbar(ggplot2::aes(ymin = lower_rr, ymax = upper_rr)) +
+    ggplot2::geom_hline(yintercept = true_param, linetype = 2) +
+    ggplot2::scale_color_manual(values = c("red", "darkgray")) +
+    ggplot2::theme(legend.position="none",
+                   panel.background = ggplot2::element_rect(fill = 'white', colour = 'white')) +
+    ggplot2::ylab("Relative risk") +
+    ggplot2::scale_x_discrete(name = "", breaks = NULL)
 }
 #'
 #' Calendar Plot

### 2016-12-05 R/plots.R
Have `coverage_plot` return ggplot object explicitly

@@ -26,7 +26,7 @@
 #' @export
 #'
 coverage_plot <- function(summarystats, true_param){
-  summarystats %>%
+  out <- summarystats %>%
     dplyr::arrange(Estimate) %>%
     dplyr::mutate(index = 1:n(),
                   rr = exp(Estimate),
@@ -43,6 +43,7 @@ coverage_plot <- function(summarystats, true_param){
                    panel.background = ggplot2::element_rect(fill = 'white', colour = 'white')) +
     ggplot2::ylab("Relative risk") +
     ggplot2::scale_x_discrete(name = "", breaks = NULL)
+  return(out)
 }
 #'
 #' Calendar Plot

### 2016-12-05 R/plots.R
Change `coverage_plot` to use standard evaluation

@@ -27,16 +27,16 @@
 #'
 coverage_plot <- function(summarystats, true_param){
   out <- summarystats %>%
-    dplyr::arrange(Estimate) %>%
-    dplyr::mutate(index = 1:n(),
-                  rr = exp(Estimate),
-                  lower_rr = exp(lower_ci),
-                  upper_rr = exp(upper_ci),
-                  includes_true = lower_rr < true_param & true_param < upper_rr) %>%
-    ggplot2::ggplot(ggplot2::aes(x = index, y = rr, color = includes_true)) +
+    dplyr::arrange_(~ Estimate) %>%
+    dplyr::mutate_(index = ~ 1:length(Estimate),
+                   rr = ~ exp(Estimate),
+                   lower_rr = ~ exp(lower_ci),
+                   upper_rr = ~ exp(upper_ci),
+                   includes_true = ~ lower_rr < true_param & true_param < upper_rr) %>%
+    ggplot2::ggplot(ggplot2::aes_(x = ~ index, y = ~ rr, color = ~ includes_true)) +
     ggplot2::coord_flip() +
     ggplot2::geom_point() +
-    ggplot2::geom_errorbar(ggplot2::aes(ymin = lower_rr, ymax = upper_rr)) +
+    ggplot2::geom_errorbar(ggplot2::aes_(ymin = ~ lower_rr, ymax = ~ upper_rr)) +
     ggplot2::geom_hline(yintercept = true_param, linetype = 2) +
     ggplot2::scale_color_manual(values = c("red", "darkgray")) +
     ggplot2::theme(legend.position="none",

### 2016-12-05 R/plots.R
Export `calendar_plot` function

@@ -64,7 +64,7 @@ coverage_plot <- function(summarystats, true_param){
 #' testdat$x[c(89,101,367,500,502,598,678,700,895)] <- 3
 #' calendar_plot(testdat, type = "discrete", labels = c("no", "yes", "maybe"))
 #'
-#'
+#' @export
 calendar_plot <- function(df, type = "continuous", labels = NULL){
   names(df) <- c("date", "x")
   if(type=="continuous"){

### 2016-12-05 R/plots.R
Make some style changes.

@@ -68,7 +68,7 @@ coverage_plot <- function(summarystats, true_param){
 calendar_plot <- function(df, type = "continuous", labels = NULL){
   names(df) <- c("date", "x")
   if(type=="continuous"){
-    Exposure<- df$x
+    Exposure <- df$x
   }
   else if(type=="discrete"){
     Exposure <- factor(df$x, levels = levels(factor(df$x)), labels = labels)
@@ -91,7 +91,7 @@ calendar_plot <- function(df, type = "continuous", labels = NULL){
       scale_y_reverse() + theme_void()
   }
   else if(type=="discrete"){
-    newplot <- plot + viridis::scale_color_viridis(discrete=TRUE) +
+    newplot <- plot + viridis::scale_color_viridis(discrete = TRUE) +
       scale_y_reverse() + theme_void()
   }
   return(newplot)

### 2016-12-05 R/plots.R
Import pipe from `dplyr` into `calendar_plot`

@@ -64,6 +64,8 @@ coverage_plot <- function(summarystats, true_param){
 #' testdat$x[c(89,101,367,500,502,598,678,700,895)] <- 3
 #' calendar_plot(testdat, type = "discrete", labels = c("no", "yes", "maybe"))
 #'
+#' @importFrom dplyr %>%
+#'
 #' @export
 calendar_plot <- function(df, type = "continuous", labels = NULL){
   names(df) <- c("date", "x")

### 2016-12-05 R/plots.R
Make some style changes and add error is user supplies something other than "discrete" or "continuous" as "type"

@@ -69,32 +69,34 @@ coverage_plot <- function(summarystats, true_param){
 #' @export
 calendar_plot <- function(df, type = "continuous", labels = NULL){
   names(df) <- c("date", "x")
+
   if(type=="continuous"){
     Exposure <- df$x
-  }
-  else if(type=="discrete"){
+  } else if(type=="discrete"){
     Exposure <- factor(df$x, levels = levels(factor(df$x)), labels = labels)
   }
+
   plot <- df %>%
-  mutate(Weekday = lubridate::wday(date),
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure) %>%
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) +
-  geom_tile(colour = "white") +
-  facet_grid(Year ~ Month, scales = "free")
+    mutate(Weekday = lubridate::wday(date),
+           Month = lubridate::month(date, label = TRUE),
+           Year = lubridate::year(date),
+           Exposure) %>%
+    group_by(Year, Month) %>%
+    dplyr::mutate(saturday = lag(Weekday) == 7,
+                  saturday = ifelse(is.na(saturday), 0, saturday),
+                  Week = 1 + cumsum(saturday)) %>%
+    ungroup() %>%
+    ggplot(aes(x = Weekday, y = Week, fill = Exposure)) +
+    geom_tile(colour = "white") +
+    facet_grid(Year ~ Month, scales = "free")
   if(type=="continuous"){
     newplot <- plot + scale_fill_gradientn(colours = viridis(256)) +
       scale_y_reverse() + theme_void()
-  }
-  else if(type=="discrete"){
+  } else if(type=="discrete"){
     newplot <- plot + viridis::scale_color_viridis(discrete = TRUE) +
       scale_y_reverse() + theme_void()
+  } else {
+    stop('The parameter `type` must be "continuous" or "discrete".')
   }
   return(newplot)
 }

### 2016-12-05 R/plots.R
Close #15
Fix viridis palette for discrete calendar plot. The main problem was we needed `scale_fill_viridis` rather than `scale_color_viridis`.

@@ -95,7 +95,7 @@ calendar_plot <- function(df, type = "continuous", labels = NULL){
     newplot <- plot + viridis::scale_fill_viridis() +
       ggplot2::scale_y_reverse() + ggplot2::theme_void()
   } else {
-    newplot <- plot + viridis::scale_color_viridis(discrete = TRUE) +
+    newplot <- plot + viridis::scale_fill_viridis(discrete = TRUE, begin = 0.2, end = 0.9) +
       ggplot2::scale_y_reverse() + ggplot2::theme_void()
   }
 

### 2016-12-05 README.Rmd
Update README to match vignette

@@
 The first element of the returned object can be used to explore the behavior of individual simulations.
-For example, to look at the relative risk point estimate and 95% confidence interval from each of the 100 simulations, as well as which 95% confidence intervals include the true relative risk using the following call:
+For example, to look at the relative risk point estimate and 95% confidence interval from each of the 100 simulations, as well as which 95% confidence intervals include the true relative risk using the `coverage_plot` function that comes with the package:
 
 ```{r fig.width = 4, fig.height = 5, fig.align = "center"}
-ex_sim[[1]] %>%
-  arrange(Estimate) %>%
-  mutate(index = 1:n(),
-         rr = exp(Estimate),
-         lower_rr = exp(lower_ci),
-         upper_rr = exp(upper_ci),
-         includes_true = lower_rr < 1.10 & 1.10 < upper_rr) %>%
-  ggplot(aes(x = index, y = rr, color = includes_true)) + 
-  coord_flip() + 
-  geom_point() + 
-  geom_errorbar(aes(ymin = lower_rr, ymax = upper_rr)) + 
-  geom_hline(yintercept = 1.10, linetype = 2) + 
-  scale_color_manual(values = c("red", "darkgray")) +
-  theme(legend.position="none",
-        panel.background = element_rect(fill='white', colour='white')) + 
-  ylab("Relative risk") +
-  scale_x_discrete(breaks = NULL) + xlab("") 
+coverage_plot(ex_sim[[1]], true_param = 1.10)
 ```
@@
 
+You can plot a calendar plot of this simulated exposure time series using the `calendar_plot` function that comes with the package:
+
 ```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-x_cont %>% 
-  mutate(Weekday = lubridate::wday(date), 
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure = x) %>% 
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
-   geom_tile(colour = "white") + 
-   facet_grid(Year ~ Month, scales = "free") + 
-  scale_fill_gradientn(colours = viridis(256)) + 
-  scale_y_reverse() + theme_void()
+calendar_plot(x_cont, type = "continuous")
 ```
@@
 ```
+
 ```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-x_bin %>% 
-  mutate(Weekday = lubridate::wday(date), 
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure = factor(x, levels = c(0, 1), 
-                    labels = c("Not exposed", "Exposed"))) %>% 
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
-   geom_tile(colour = "white") + 
-   facet_grid(Year ~ Month, scales = "free") + 
-  scale_fill_manual(values = c("#414487FF", "#7AD151FF")) + 
-  scale_y_reverse() + theme_void()
+calendar_plot(x_bin, type = "discrete", labels = c("Not exposed", "Exposed"))
 ```
@@
 head(testexp)
-qplot(testexp$date, testexp$x)+geom_point()+coord_cartesian(ylim = c(0,110)) + labs(title = "Exposure with a cos1linear trend", x = "Date", y="Exposure")+theme_bw()
+qplot(testexp$date, testexp$x)+geom_point()+coord_cartesian(ylim = c(0,110)) + 
+  labs(title = "Exposure with a cos1linear trend", x = "Date", y="Exposure")+theme_bw()
 ```
@@
 ```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-testexp %>% 
-  mutate(Weekday = lubridate::wday(date), 
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure = x) %>% 
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
-   geom_tile(colour = "white") + 
-   facet_grid(Year ~ Month, scales = "free") + 
-  scale_fill_gradientn(colours = viridis(256)) + 
-  scale_y_reverse() + theme_void()
+calendar_plot(testexp, type = "continuous")
 ```
@@
                         amp = .2, exposure_type = "continuous")
-qplot(smallamp$date, smallamp$x)+geom_point()+coord_cartesian(ylim = c(0,110))+labs(title = "Cos1linear exposure with smaller amplitude", x="Date", y="Exposure") + theme_bw()
+qplot(smallamp$date, smallamp$x) +
+  geom_point() + coord_cartesian(ylim = c(0,110)) +
+  labs(title = "Cos1linear exposure with smaller amplitude", x="Date", y="Exposure") + theme_bw()
 ```
@@
 ```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-smallamp %>% 
-  mutate(Weekday = lubridate::wday(date), 
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure = x) %>% 
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
-   geom_tile(colour = "white") + 
-   facet_grid(Year ~ Month, scales = "free") + 
-  scale_fill_gradientn(colours = viridis(256)) + 
-  scale_y_reverse() + theme_void()
+calendar_plot(smallamp, type = "continuous")
 ```
@@
 ```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-testbin %>% 
-  mutate(Weekday = lubridate::wday(date), 
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure = factor(x, levels = c(0, 1), 
-                    labels = c("Not exposed", "Exposed"))) %>% 
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
-   geom_tile(colour = "white") + 
-   facet_grid(Year ~ Month, scales = "free") + 
-  scale_fill_manual(values = c("#414487FF", "#7AD151FF")) + 
-  scale_y_reverse() + theme_void()
+calendar_plot(testbin, type = "discrete", labels = c("Not exposed", "Exposed"))
 ```
@@
 testout <- sim_outcome(exposure = testexp2, average_outcome = 22, trend = "linear", rr = 1.01)
-qplot(testout$date, testout$outcome)+geom_point()+labs(title = "Health outcomes with a linear trend", x="Date", y="Outcome") + theme_bw()
+qplot(testout$date, testout$outcome) + geom_point() +
+  labs(title = "Health outcomes with a linear trend", x="Date", y="Outcome") + theme_bw()
 ```
@@
 ```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-testout %>% 
-  mutate(Weekday = lubridate::wday(date), 
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure = outcome) %>% 
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
-   geom_tile(colour = "white") + 
-   facet_grid(Year ~ Month, scales = "free") + 
-  scale_fill_gradientn(colours = viridis(256)) + 
-  scale_y_reverse() + theme_void()
+calendar_plot(testout %>% select(date, outcome), type = "continuous")
 ```
@@
 ```{r}
-sims <- create_sims(n_reps=10, n=100, central = 100, sd = 10,
+sims <- create_sims(n_reps = 10, n = 100, central = 100, sd = 10,
              exposure_type="continuous", exposure_trend = "cos1",
@@
 ```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-customexp %>% 
-  mutate(Weekday = lubridate::wday(date), 
-         Month = lubridate::month(date, label = TRUE),
-         Year = lubridate::year(date),
-         Exposure = x) %>% 
-  group_by(Year, Month) %>%
-  dplyr::mutate(saturday = lag(Weekday) == 7,
-                saturday = ifelse(is.na(saturday), 0, saturday),
-                Week = 1 + cumsum(saturday)) %>%
-  ungroup() %>%
-  ggplot(aes(x = Weekday, y = Week, fill = Exposure)) + 
-   geom_tile(colour = "white") + 
-   facet_grid(Year ~ Month, scales = "free") + 
-  scale_fill_gradientn(colours = viridis(256)) + 
-  scale_y_reverse() + theme_void()
+calendar_plot(customexp, type = "continuous")
 ```

### 2016-12-05 R/customizable_functions.R
Close #14, close #13.
A few things here. First, `eesim` wasn't passing all the custom arguments through to later functions, which is why there was an error for the custom exposure example. For the `baseline` bug, the problem was that `baseline` came through as a numerical vector in this case, which the `create_lambda` is expecting a dataframe with a column also named `baseline`. I added a line to the function to check if `baseline` is a dataframe and, if so, to pull out just the `baseline` column for the `baseline` object. This isn't terribly elegant, but it should get the job done.

@@ -217,7 +217,8 @@ create_baseline <- function(n, average_outcome, trend, amp,
 
 create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
   if(is.null(cust_lambda_func)){
-    log_lambda <- log(baseline$baseline) + log(rr) * exposure
+    if("data.frame" %in% class(baseline)){ baseline <- baseline$baseline }
+    log_lambda <- log(baseline) + log(rr) * exposure
     lambda <- exp(log_lambda)
   } else {
     arguments <- list(...)
@@ -268,15 +269,13 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                             exposure = exposure$x,
                             rr = rr)
     outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
-  }
-  else if (is.null(cust_lambda_func) & !is.null(cust_base_func)){
+  } else if (is.null(cust_lambda_func) & !is.null(cust_base_func)){
     baseline <- do.call(cust_base_func, cust_base_args)
     lambda <- create_lambda(baseline = baseline,
                             exposure = exposure$x,
                             rr = rr)
     outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
-  }
-  else if (is.null(cust_base_func) & !is.null(cust_lambda_func)){
+  } else if (is.null(cust_base_func) & !is.null(cust_lambda_func)){
     baseline <- create_baseline(n = nrow(exposure),
                                 average_outcome = average_outcome,
                                 trend = trend,
@@ -284,8 +283,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
     cust_lambda_args$baseline <- baseline$baseline
     lambda <- do.call(cust_lambda_func, cust_lambda_args)
     outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
-  }
-  else {
+  } else {
     baseline <- do.call(cust_base_func, cust_base_args)
     cust_lambda_args$baseline <- baseline
     lambda <- do.call(cust_lambda_func, cust_lambda_args)
@@ -358,6 +356,7 @@ create_sims <- function(n_reps, n, central, sd, exposure_type, exposure_trend,
                         cust_exp_func = NULL, cust_exp_args = NULL,
                         cust_base_func = NULL, cust_lambda_func = NULL,
                         cust_base_args = NULL, cust_lambda_args = NULL){
+
[…]

### 2017-02-21 writing/EpiDraft.Rmd
Write epi intro

---
title: "Outstanding Title"
author: "Sarah Koehler"
date: "February 21, 2017"
output:
  pdf_document: default
  html_document: default
---

#Introduction

Simulation studies are important in environmental epidemiology research on air pollution, temperature, and other exposures. For example, simulated data can be used to test new statistical models and perform power analyses. Developing the code to simulate environmental time series for these types of studies can be time consuming, and simulation methods have varied across previous simulation studies [cite studies], complicating the comparison of results across different studies.

Here, we present an R package that automates simulation of environmental time series for such simulations, offering both sensible defaults for each stage of the simulation process and also the option to customize any of the simulation steps while using defaults for any steps that do not require customization. This package can be used to aid simulation studies to explore environmental epidemiology models relating acute health outcomes and daily environmental exposures; it can also be used to estimate the power of studies when planning or proposing future research.

#Previous Simulation Studies

#Simulating Data and Testing Models

#Power Analysis

### 2017-02-21 writing/JSSDraft/JSSDraft.Rmd
Write epi intro

@@
 Here, we present an R package that automates simulation of environmental time series for such simulations, offering both sensible defaults for each stage of the simulation process and also the option to customize any of the simulation steps while using defaults for any steps that do not require customization.
-This packge can be used to aid simulation studies to explore environmental epidemiology models relating acute health outcomes and daily environmental exposures; it can also be used to estimate the power of studies when planning or proposing future research.
+This package can be used to aid simulation studies to explore environmental epidemiology models relating acute health outcomes and daily environmental exposures; it can also be used to estimate the power of studies when planning or proposing future research.
 keywords:

### 2017-02-23 writing/EpiDraft.Rmd
Add to simulating data section

@@
 For example, simulated data can be used to test new statistical models and perform power analyses.
-Developing the code to simulate environmental time series for these types of studies can be time consuming, and simulation methods have varied across previous simulation studies [cite studies], complicating the comparison of results across different studies.
+Two current challenges of simulating time series data for epidemiology studies are (1) methods for simulating are inconsistent, making it difficult to compare results from different studies, and (2) developing the code to simulate environmental time series for these types of studies can be time consuming.
+We introduce eesim, an open-source software package in R that offers a solution to these challenges.
+
+Paragraph about previous simulation studies
 
@@
 
-#Previous Simulation Studies
-
 #Simulating Data and Testing Models
 
+[Chart / Figure of functions included and simulation process]
+
+This package allows you to simulate time series of environmental health data and perform simulation-based power analyses and other measures of model performance. The package includes four main parts:
+
+1. Generation of exposure data; 
+2. Generation of outcome data;
+3. Fitting models to simulated data; and 
+4. Evaluating model performance on simulated data.
+
+The user has the option to customize different aspects of the simulation at each of these steps.
+
+The main function of this package is the `eesim` function. You can use the `eesim` function to conduct all four steps of the simulation process outlined above at once.
+
+The `eesim` function requires several arguments, both numeric and character, including the desired number of repetitions of the simulation, the sample size for each repetiton, the type of exposure values (binary or continous), the expected average value of the response variable, the relative risk, and the type of model desired.
+The user has the option to input more arguments specifying trends or custom functions for the exposure, the outcome, or both.
+
+The `eesim` function returns a list with two elements.
+The first element gives simulation-specific results: the estimated effect size, standard error, t- and p-values, and upper and lower 95% confidence bounds when a model was applied to each set of simulated data.
+The second element gives some measures of model assessment, assessed over all simulations, including the mean effect size and relative risk estimates across simulations.
+
+[insert equation here?]
+
+For example, you can use the following call to (1) generate 3 repetitions of 100 observations of a continuous exposure with mean 50 and standard deviation 5, with a seasonal trend in expected value; (2) generate 10 associated outcome values, where the outcome has an average value of 20 and a relative risk of 1.10 per one-unit increase in the exposure; (3) fit a generalized linear model that controls for long-term and seasonal trends with a natural cubic spline with 2 degrees of freedom per year; and (4) evaluate the performace of that model on the simulated data:
+
+```{r}
+library(eesim)
+eesim(n_reps = 3, n = 100, central = 50, sd = 5, exposure_type = "continuous", 
+      outcome_trend = "curvilinear", average_outcome = 20, rr = 1.10, model = "spline", df_year = 2)
+```
+
+The above output gives the estimated effect sizes for each of the three repetitions along with their standard errors, test statistics, p-values, and confidence intervals.
+It also summarizes the outcomes of all three repetitions with a mean estimated beta (which equals the log of the estimated relative risk), estimated relative risk, the variance of the estimated betas, and the mean variance of the individual betas.
+
 #Power Analysis

### 2017-02-23 writing/JSSDraft/JSSDraft.Rmd
Add to simulating data section

@@
 For example, \citet{Bellan2015} used a simulation study to determine the power and statistical validity of two potential Ebola vaccine trials, a randomized controlled trial and a stepped-wedge cluster trial.
-Simulation was useful in this case because the trial would be conducting in a population with clustering by geographical district and with a temporally changing hazard of Ebola infection over the trial period \citep{Bellan2015}.
+Simulation was useful in this case because the trial would be conducted in a population with clustering by geographical district and with a temporally changing hazard of Ebola infection over the trial period \citep{Bellan2015}.
 

### 2017-02-23 writing/EpiDraft.Rmd
Complete Simulating Data section.

@@
 
+The `eesim` output also gives the relative parameter estimate bias, according to the following equation  from (citet{Beaujean2014}), given $\theta_{H}$ (true value of the parameter, as set when creating the simulated data) and $\hat{\theta}$ (value of the parameter estimated from applying a model to the simulated data):
+
+\begin{equation}
+\theta_{bias} = \frac{\hat{\theta} - \theta_{H}}{\theta_{H}}
+\end{equation}
+
+Finally, the `eesim` output gives the coverage and power of the model.
+
 #Power Analysis

### 2017-02-24 writing/EpiDraft.Rmd
Add previous studies to intro

@@
 Two current challenges of simulating time series data for epidemiology studies are (1) methods for simulating are inconsistent, making it difficult to compare results from different studies, and (2) developing the code to simulate environmental time series for these types of studies can be time consuming.
-We introduce eesim, an open-source software package in R that offers a solution to these challenges.
+We introduce `eesim`, an open-source software package in R that offers a solution to these challenges.
 
-Paragraph about previous simulation studies
+Simulated data has been used in a number of environmental epidemiology studies on various kinds of exposures and outcomes.
+(citet{Armstrong2014}) used a simulation study to investigate short-term mortality displacement following heat waves.
+(citet{Bateson1999}), (citet{Bateson2001}), and (citet{Peng2006}) used simulated data with confounding seasonal trends to compare experimental designs and model choice.
+(citet{Strickland2015}), (citet{Gryparis2009}), and (citet{Butland2013}) investigated spatial misalignment and geographic heterogeneity of exposures using simulated data.
+(citet{Roberts2006}) used a simulation study to test the reliability of using AIC to determine the shape of the relationship between particulate matter and mortality.
+(citet{Bobb2015}) used simulation to assess the performance of a proposed method for estimating the health effects of multi-pollutant exposures called Bayesian kernel machine regression (BKMR).
 
 Here, we present an R package that automates simulation of environmental time series for such simulations, offering both sensible defaults for each stage of the simulation process and also the option to customize any of the simulation steps while using defaults for any steps that do not require customization.
+There are a few important advantages to having an open-source software package that can simulate this type of environmental time series for simulation studies.
+First, current methods of simulating environmental time series to use to assess modeling choices vary across different studies.
+If results from a simulation study are not robust to different choices in how to simulate the data used in the study, results across studies using different methods of simulating data would not be comparable \citep{Alfons2010}.
+
+This package offers several methods of simulating time series so researchers can check if study results are robust to data simulation method.
+This software also allows for consistency and comparison of data simulation across studies and will save researchers time and effort in developing study-specific code.
 This package can be used to aid simulation studies to explore environmental epidemiology models relating acute health outcomes and daily environmental exposures; it can also be used to estimate the power of studies when planning or proposing future research.
@@
 
-[Chart / Figure of functions included and simulation process]
-
-This package allows you to simulate time series of environmental health data and perform simulation-based power analyses and other measures of model performance. The package includes four main parts:
+Simulation studies can be used to assess different models and modeling choices.
+This way of assessing models can be particularly useful for complex statistical methods, for which assessing performance metrics can often prove difficult (citep{Alfons2010}).
+The `eesim` package allows you to simulate time series of environmental health data and perform simulation-based power analyses and other measures of model performance.
+The package includes four main parts:
 
@@
 
-The main function of this package is the `eesim` function. You can use the `eesim` function to conduct all four steps of the simulation process outlined above at once.
-
+The main function of this package is the `eesim` function.
+You can use the `eesim` function to conduct all four steps of the simulation process outlined above at once.
 The `eesim` function requires several arguments, both numeric and character, including the desired number of repetitions of the simulation, the sample size for each repetiton, the type of exposure values (binary or continous), the expected average value of the response variable, the relative risk, and the type of model desired.
@@
 #Power Analysis
+
+Simulation studies can be used to determine effective sample size or power when planning or proposing future research studies.
+This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable (citep{Bellan2015, Johnson2015, Burton2009, Alfons2010, Beaujean2014}).
+Simulation studies can often be used to generate power analyses under a more realistic set of assumptions (citep{Burton2009, Beaujean2014}).

### 2017-02-24 writing/EpiDraft.Rmd
Add to Power Analysis section

@@
 
-Here, we present an R package that automates simulation of environmental time series for such simulations, offering both sensible defaults for each stage of the simulation process and also the option to customize any of the simulation steps while using defaults for any steps that do not require customization.
+Here, we present an R package that automates simulation of environmental time series for such studies, offering both sensible defaults for each stage of the simulation process and also the option to customize any of the simulation steps while using defaults for any steps that do not require customization.
 There are a few important advantages to having an open-source software package that can simulate this type of environmental time series for simulation studies.
-First, current methods of simulating environmental time series to use to assess modeling choices vary across different studies.
-If results from a simulation study are not robust to different choices in how to simulate the data used in the study, results across studies using different methods of simulating data would not be comparable \citep{Alfons2010}.
+This package offers several methods of simulating time series so researchers can check if study results are robust to the data simulation method (citep{Alfons2010}).
+This software also allows for consistency and comparison of data simulation across studies and will save researchers time and effort in developing study-specific code.
 
-This package offers several methods of simulating time series so researchers can check if study results are robust to data simulation method.
-This software also allows for consistency and comparison of data simulation across studies and will save researchers time and effort in developing study-specific code.
 This package can be used to aid simulation studies to explore environmental epidemiology models relating acute health outcomes and daily environmental exposures; it can also be used to estimate the power of studies when planning or proposing future research.
@@
 Simulation studies can be used to determine effective sample size or power when planning or proposing future research studies.
-This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable (citep{Bellan2015, Johnson2015, Burton2009, Alfons2010, Beaujean2014}).
+This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable (citep{Bellan2015, Johnson2015, Burton2009, Alfons2010, Beaujean2014, Schoemann2014}).
 Simulation studies can often be used to generate power analyses under a more realistic set of assumptions (citep{Burton2009, Beaujean2014}).
+
+In `eesim`, the function `power_calc` can be used to investigate the effects of various sample sizes or relative risks on power.
+This function requires arguments including the parameter to be varied (sample size or relative risk), the desired values of that parameter, and lists of the same arguments as used in the `eesim` function.
+The `power_calc` function outputs a table of the values of the varied parameter along with their corresponding power.
+Here is an example of using `power_calc` to test power for different sample sizes:
+
+```{r}
+power_calc(varying = "n", values = c(5, 50, 500), simargs = list(n_reps = 10,
+           rr = 1.02, central = 50, sd = 5, exposure_trend = "no trend",
+           exposure_amp = .6, exposure_type = "continuous",
+           outcome_trend = "cos1", outcome_amp = .6, average_outcome = 100),
+           fitargs = list(model = "spline"))
+```
+
+The `power_calc` function will also return a plot if `plot=TRUE` is specified.

### 2017-02-24 writing/EpiDraft.Rmd
Fix bugs in meeting

@@
 library(eesim)
-eesim(n_reps = 3, n = 100, central = 50, sd = 5, exposure_type = "continuous", 
-      outcome_trend = "curvilinear", average_outcome = 20, rr = 1.10, model = "spline", df_year = 2)
+eesim(n_reps = 3, n = 100, central = 50, sd = 5, outcome_amp = .6, exposure_type = "continuous", outcome_trend = "cos1", average_outcome = 20, rr = 1.10, model = "spline", df_year = 2)
 ```
@@
 ```{r}
-power_calc(varying = "n", values = c(5, 50, 500), simargs = list(n_reps = 10,
-           rr = 1.02, central = 50, sd = 5, exposure_trend = "no trend",
+power_calc(varying = "n", values = c(365*1:3), simargs = list(n_reps = 10,
+           rr = 1.5, central = 50, sd = 5, exposure_trend = "no trend",
            exposure_amp = .6, exposure_type = "continuous",

### 2017-03-20 writing/EpiDraft.Rmd
mystery leftovers

@@
 date: "February 21, 2017"
+bibliography: mybibfile.bib
 output:

### 2017-03-20 R/assess_performance.R

@@ -203,8 +203,8 @@ power_calc <- function(varying, values, simargs = list(), fitargs = list(),
   out <- data.frame(x = values, power = NA)
   simargs$n <- out$x
   if(varying == "n"){
-    simargs$n <- out$x
     for(i in 1:nrow(out)){
+      simargs$n <- out$x[i]
       rep_df[[i]] <- do.call(create_sims, simargs) #needs to output a list of data frames with date, x, outcome
       #How do I make it cycle through values of n and put each output from create_sims into a list?
       fitargs$outcome <- rep_df #this is not the right input for fit_mods

### 2017-03-20 writing/DraftOne.Rmd
Attempt at power_calc using purrr in Draft1 - not working :(

@@
 
+##2nd attempt at power_calc
+
+```{r}
+power_calc <- function(varying, values, n_reps, n=NULL, central, sd, exposure_type,
+                       exposure_trend, exposure_amp, average_outcome, outcome_trend,
+                        outcome_amp, rr=NULL, start.date = "2000-01-01",
+                        cust_exp_func = NULL, cust_exp_args = NULL,
+                        cust_base_func = NULL, cust_lambda_func = NULL,
+                        cust_base_args = NULL, cust_lambda_args = NULL, 
+                       model, df_year=7){
+  if(varying == "n"){
+    rep_df <- n %>% map(create_sims, n_reps=n_reps, central=central, sd=sd, 
+                        exposure_type = exposure_type,
+                        exposure_trend=exposure_trend,exposure_amp=exposure_amp,
+                        average_outcome=average_outcome,outcome_trend=outcome_trend,
+                        outcome_amp = outcome_amp, rr=rr, start.date = start.date)
+      fits <- rep_df %>% map(fit_mods, model=model)
+      power <- fits %>% map(power_beta) #makes a list, want to extract the values of power and put in a data frame with values of n.
+    }
+  else if(varying == "rr"){
+    rep_df <- rr %>% map(create_sims, n=n, n_reps=n_reps, central=central, sd=sd, 
+                        exposure_type = exposure_type,
+                        exposure_trend=exposure_trend,exposure_amp=exposure_amp,
+                        average_outcome=average_outcome,outcome_trend=outcome_trend,
+                        outcome_amp = outcome_amp, start.date = start.date)
+      fits <- rep_df %>% map(fit_mods, model=model)
+      power <- fits %>% map(power_beta) #makes a list, want to extract the values of power and put in a data frame with values of rr.
+    }
+  if(plot == TRUE){
+    my_plot <- ggplot2::ggplot(out, ggplot2::aes_(x = ~ x, y = ~ power)) +
+      ggplot2::geom_line() + ggplot2::theme_minimal() +
+      ggplot2::xlab(varying)
+    print(my_plot)
+  }
+  return(power)
+}
+
+#Example that doesn't work: 
+power_calc(varying = "n", values = c(5, 10, 25), n_reps = 3, central = 100, sd=10, exposure_type = "continuous", exposure_trend = "cos1", exposure_amp = .6, average_outcome=22, outcome_trend = "no trend", outcome_amp = .6, rr=1.1, start.date = "2000-01-01")
+```
+
 ## References

### 2017-03-21 writing/DraftOne.Rmd
power_calc - What even??

@@
 
-##2nd attempt at power_calc
+##2nd attempt at power_calc using purrr
 
@@
 #Example that doesn't work: 
-power_calc(varying = "n", values = c(5, 10, 25), n_reps = 3, central = 100, sd=10, exposure_type = "continuous", exposure_trend = "cos1", exposure_amp = .6, average_outcome=22, outcome_trend = "no trend", outcome_amp = .6, rr=1.1, start.date = "2000-01-01")
+power_calc(varying = "n", values = c(5, 10, 25), n_reps = 3, central = 100, sd=10, exposure_type = "continuous", exposure_trend = "cos1", exposure_amp = .6, average_outcome=22, outcome_trend = "no trend", outcome_amp = .6, rr=1.1, start.date = "2000-01-01", model = "spline", df_year = 1)
+```
+
+Not working, so step by step breakdown:
+
+```{r}
+#Establishing values for arguments
+varying = "n"
+values = c(5, 10, 25)
+n_reps = 3
+central = 100
+sd=10
+exposure_type = "continuous"
+exposure_trend = "cos1"
+exposure_amp = .6
+average_outcome=22
+outcome_trend = "no trend"
+outcome_amp = .6
+rr=1.1
+start.date = "2000-01-01"
+model = "spline"
+df_year = 1
+
+n<- c(5, 10, 20)
+
+#Getting list of data frames with date, x, outcome for each run
+rep_df <- n %>% map(create_sims, n_reps=n_reps, central=central, sd=sd, 
+                        exposure_type=exposure_type,
+                        exposure_trend=exposure_trend,exposure_amp=exposure_amp,
+                        average_outcome=average_outcome,outcome_trend=outcome_trend,
+                        outcome_amp = outcome_amp, rr=rr, start.date = start.date)
+rep_df
+
+#Try with actual values:
+
+rep_df <- n %>% map(create_sims, n_reps=3, central=100, sd=10, 
+                        exposure_type = "continuous",
+                        exposure_trend= "cos1",exposure_amp=.6,
+                        average_outcome=22,outcome_trend="no trend",
+                        outcome_amp = .6, rr=1.1, start.date = "2000-01-01")
+rep_df
+#This works, but why are the outcomes crazy? Oh noooooooo :( 
 ```

### 2017-03-21 R/customizable_functions.R

@@ -343,7 +343,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #'    for date, exposure, and outcomes, and estimates from fitting models
 #'
 #' @examples
-#' create_sims(n_reps=10, n=10, central = 100, sd = 10,
+#' create_sims(n_reps=3, n=100, central = 100, sd = 10,
 #'             exposure_type="continuous", exposure_trend = "cos1",
 #'             exposure_amp = .6, average_outcome = 22,
 #'             outcome_trend = "no trend", outcome_amp = .6, rr = 1.01)

### 2017-03-21 writing/EpiDraft.Rmd
Fix in-line citations into Rmd format

@@
 Simulated data has been used in a number of environmental epidemiology studies on various kinds of exposures and outcomes.
-(citet{Armstrong2014}) used a simulation study to investigate short-term mortality displacement following heat waves.
-(citet{Bateson1999}), (citet{Bateson2001}), and (citet{Peng2006}) used simulated data with confounding seasonal trends to compare experimental designs and model choice.
-(citet{Strickland2015}), (citet{Gryparis2009}), and (citet{Butland2013}) investigated spatial misalignment and geographic heterogeneity of exposures using simulated data.
-(citet{Roberts2006}) used a simulation study to test the reliability of using AIC to determine the shape of the relationship between particulate matter and mortality.
-(citet{Bobb2015}) used simulation to assess the performance of a proposed method for estimating the health effects of multi-pollutant exposures called Bayesian kernel machine regression (BKMR).
+[@Armstrong2014] used a simulation study to investigate short-term mortality displacement following heat waves.
+[@Bateson1999], [@Bateson2001], and [@Peng2006] used simulated data with confounding seasonal trends to compare experimental designs and model choice.
+[@Strickland2015], [@Gryparis2009], and [@Butland2013] investigated spatial misalignment and geographic heterogeneity of exposures using simulated data.
+[@Roberts2006] used a simulation study to test the reliability of using AIC to determine the shape of the relationship between particulate matter and mortality.
+[@Bobb2015] used simulation to assess the performance of a proposed method for estimating the health effects of multi-pollutant exposures called Bayesian kernel machine regression (BKMR).
 
@@
 There are a few important advantages to having an open-source software package that can simulate this type of environmental time series for simulation studies.
-This package offers several methods of simulating time series so researchers can check if study results are robust to the data simulation method (citep{Alfons2010}).
+This package offers several methods of simulating time series so researchers can check if study results are robust to the data simulation method [@Alfons2010].
 This software also allows for consistency and comparison of data simulation across studies and will save researchers time and effort in developing study-specific code.
@@
 Simulation studies can be used to assess different models and modeling choices.
-This way of assessing models can be particularly useful for complex statistical methods, for which assessing performance metrics can often prove difficult (citep{Alfons2010}).
+This way of assessing models can be particularly useful for complex statistical methods, for which assessing performance metrics can often prove difficult [@Alfons2010].
 The `eesim` package allows you to simulate time series of environmental health data and perform simulation-based power analyses and other measures of model performance.
@@
 
-The `eesim` output also gives the relative parameter estimate bias, according to the following equation  from (citet{Beaujean2014}), given $\theta_{H}$ (true value of the parameter, as set when creating the simulated data) and $\hat{\theta}$ (value of the parameter estimated from applying a model to the simulated data):
+The `eesim` output also gives the relative parameter estimate bias, according to the following equation  from [@Beaujean2014], given $\theta_{H}$ (true value of the parameter, as set when creating the simulated data) and $\hat{\theta}$ (value of the parameter estimated from applying a model to the simulated data):
 
@@
 Simulation studies can be used to determine effective sample size or power when planning or proposing future research studies.
-This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable (citep{Bellan2015, Johnson2015, Burton2009, Alfons2010, Beaujean2014, Schoemann2014}).
-Simulation studies can often be used to generate power analyses under a more realistic set of assumptions (citep{Burton2009, Beaujean2014}).
+This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable [@Bellan2015; @Johnson2015; @Burton2009; @Alfons2010; @Beaujean2014; @Schoemann2014].
+Simulation studies can often be used to generate power analyses under a more realistic set of assumptions [@Burton2009; @Beaujean2014].
 
@@
 ```{r}
-power_calc(varying = "n", values = c(365*1:3), simargs = list(n_reps = 10,
-           rr = 1.5, central = 50, sd = 5, exposure_trend = "no trend",
-           exposure_amp = .6, exposure_type = "continuous",
-           outcome_trend = "cos1", outcome_amp = .6, average_outcome = 100),
-           fitargs = list(model = "spline"))
+#Power_calc momentarily out of commission :(
 ```

### 2017-03-22 writing/DraftOne.Rmd
Meeting 3/21

@@
 
-n<- c(5, 10, 20)
+n<- 100
 
 #Getting list of data frames with date, x, outcome for each run
-rep_df <- n %>% map(create_sims, n_reps=n_reps, central=central, sd=sd, 
+rep_df <- n %>% purrr::map(create_sims, n_reps=n_reps, central=central, sd=sd, 
                         exposure_type=exposure_type,
@@
 
-rep_df <- n %>% map(create_sims, n_reps=3, central=100, sd=10, 
+rep_df <- n %>% purrr::map(create_sims, n_reps=3, central=100, sd=10, 
                         exposure_type = "continuous",
@@
                         average_outcome=22,outcome_trend="no trend",
-                        outcome_amp = .6, rr=1.1, start.date = "2000-01-01")
+                        outcome_amp = .6, rr=1.01, start.date = "2000-01-01")
 rep_df
+
 #This works, but why are the outcomes crazy? Oh noooooooo :( 

### 2017-04-04 writing/DraftOne.Rmd
power_calc returns power

@@
                         cust_base_args = NULL, cust_lambda_args = NULL, 
-                       model, df_year=7){
+                       model, df_year=7, plot=FALSE){
   if(varying == "n"){
-    rep_df <- n %>% map(create_sims, n_reps=n_reps, central=central, sd=sd, 
+    rep_df <- values %>% purrr::map(create_sims, n_reps=n_reps, central=central, sd=sd, 
                         exposure_type = exposure_type,
@@
                         outcome_amp = outcome_amp, rr=rr, start.date = start.date)
-      fits <- rep_df %>% map(fit_mods, model=model)
-      power <- fits %>% map(power_beta) #makes a list, want to extract the values of power and put in a data frame with values of n.
+      fits <- rep_df %>% purrr::map(fit_mods, model=model)
+      power <- fits %>% purrr::map(power_beta) #makes a list, want to extract the values of power and put in a data frame with values of n.
     }
   else if(varying == "rr"){
-    rep_df <- rr %>% map(create_sims, n=n, n_reps=n_reps, central=central, sd=sd, 
-                        exposure_type = exposure_type,
+    rep_df <- values %>% purrr::map(create_sims, n=n, n_reps=n_reps, central=central, sd=sd,                         exposure_type = exposure_type,
                         exposure_trend=exposure_trend,exposure_amp=exposure_amp,
@@
                         outcome_amp = outcome_amp, start.date = start.date)
-      fits <- rep_df %>% map(fit_mods, model=model)
-      power <- fits %>% map(power_beta) #makes a list, want to extract the values of power and put in a data frame with values of rr.
+      fits <- rep_df %>% purrr::map(fit_mods, model=model)
+      power <- fits %>% purrr::map(power_beta) #makes a list, want to extract the values of power and put in a data frame with values of rr.
+  }
+  else if(varying=="average_outcome"){
+    rep_df <- values %>% purrr::map(create_sims, n=n, n_reps=n_reps,
+                        central=central,
+                        sd=sd, exposure_type = exposure_type, exposure_trend=exposure_trend,
+                        exposure_amp=exposure_amp, outcome_trend=outcome_trend,
+                        outcome_amp = outcome_amp, rr=rr, start.date = start.date)
+      fits <- rep_df %>% purrr::map(fit_mods, model=model)
+      power <- fits %>% purrr::map(power_beta)
   }
@@
 
-#Example that doesn't work: 
-power_calc(varying = "n", values = c(5, 10, 25), n_reps = 3, central = 100, sd=10, exposure_type = "continuous", exposure_trend = "cos1", exposure_amp = .6, average_outcome=22, outcome_trend = "no trend", outcome_amp = .6, rr=1.1, start.date = "2000-01-01", model = "spline", df_year = 1)
+#Example: 
+pow <- power_calc(varying = "rr", values = c(1.2,1.1,1.01), n=50, n_reps = 3, central = 100, sd=10, exposure_type = "continuous", exposure_trend = "cos1", exposure_amp = .6, average_outcome=22, outcome_trend = "no trend", outcome_amp = .6, start.date = "2000-01-01", model = "spline", df_year = 1)
 ```
@@
 
-rep_df <- n %>% purrr::map(create_sims, n_reps=3, central=100, sd=10, 
+rep_df <- n %>% purrr::map(create_sims, n_reps=10, central=100, sd=10, 
                         exposure_type = "continuous",
@@
                         outcome_amp = .6, rr=1.01, start.date = "2000-01-01")
-rep_df
-
-#This works, but why are the outcomes crazy? Oh noooooooo :( 
+fits <- rep_df %>% map(fit_mods, model="spline")
+power <- fits %>% map(power_beta)
+power
+#This works!!!!!
 ```

### 2017-04-04 R/assess_performance.R

@@ -182,43 +182,54 @@ check_sims <- function(df, true_rr){
 #' @param varying A character string of the parameter to be varied.  Choices are
 #'    "n" or "rr"
 #' @param values A numeric vector of the chosen values of the varying parameters
-#' @param simargs A list of arguments to be passed to the create_sims function
-#' @param fitargs A list of arguments to be passed to the fit_mods function
 #' @param plot "TRUE" or "FALSE" for whether to produce a plot
 #' @inheritParams power_beta
+#' @inheritParams create_sims
+#' @inheritParams fit_mods
 #'
 #' @return Data frame with the value of the varying parameter and its
 #'    corresponding power
 #'
 #' @examples
-#' power_calc(varying = "n", values = c(5 * (1:5)), simargs = list(n_reps = 50,
-#'            rr = 1.02, central = 50, sd = 5, exposure_trend = "no trend",
-#'            exposure_amp = .6, exposure_type = "continuous",
-#'            outcome_trend = "cos1", outcome_amp = .6, average_outcome = 100),
-#'            fitargs = list(model = "spline"))
+#' power_calc(varying = "n", values = c(15, 50, 100), n_reps = 20,
+#'            central = 100, sd=10, rr = 1.01, exposure_type = "continuous",
+#'            exposure_trend = "cos1", exposure_amp = .6, average_outcome=22,
+#'            outcome_trend = "no trend", outcome_amp = .6,
+#'            start.date = "2000-01-01", model = "spline", df_year = 1, plot=TRUE)
 #'
 #' @export
-power_calc <- function(varying, values, simargs = list(), fitargs = list(),
-                       plot = FALSE){
-  out <- data.frame(x = values, power = NA)
-  simargs$n <- out$x
+power_calc <- function(varying, values, n_reps, n=NULL, central, sd, exposure_type,
+                       exposure_trend, exposure_amp, average_outcome, outcome_trend,
+                       outcome_amp, rr=NULL, start.date = "2000-01-01",
+                       cust_exp_func = NULL, cust_exp_args = NULL,
+                       cust_base_func = NULL, cust_lambda_func = NULL,
+                       cust_base_args = NULL, cust_lambda_args = NULL,
+                       model, df_year=7, plot=FALSE){
   if(varying == "n"){
-    for(i in 1:nrow(out)){
-      simargs$n <- out$x[i]
-      rep_df[[i]] <- do.call(create_sims, simargs) #needs to output a list of data frames with date, x, outcome
[…]

### 2017-04-04 R/assess_performance.R
power_calc returns data frame

@@ -187,15 +187,15 @@ check_sims <- function(df, true_rr){
 #' @inheritParams create_sims
 #' @inheritParams fit_mods
 #'
-#' @return Data frame with the value of the varying parameter and its
+#' @return Data frame with the values of the varying parameter and their
 #'    corresponding power
 #'
 #' @examples
-#' power_calc(varying = "n", values = c(15, 50, 100), n_reps = 20,
+#' pow <- power_calc(varying = "n", values = c(15, 50, 100), n_reps = 20,
 #'            central = 100, sd=10, rr = 1.01, exposure_type = "continuous",
 #'            exposure_trend = "cos1", exposure_amp = .6, average_outcome=22,
 #'            outcome_trend = "no trend", outcome_amp = .6,
-#'            start.date = "2000-01-01", model = "spline", df_year = 1, plot=TRUE)
+#'            start.date = "2000-01-01", model = "spline", df_year = 1)
 #'
 #' @export
 power_calc <- function(varying, values, n_reps, n=NULL, central, sd, exposure_type,
@@ -231,11 +231,16 @@ power_calc <- function(varying, values, n_reps, n=NULL, central, sd, exposure_ty
     fits <- rep_df %>% purrr::map(fit_mods, model=model)
     power <- fits %>% purrr::map(power_beta)
   }
+  powervec <- rep(0, length(power))
+  for (i in 1:length(power)){
+  powervec[i] <- power[[i]][,1]
+  }
+  dat <- data.frame(values=values, power=powervec)
   if(plot == TRUE){
     my_plot <- ggplot2::ggplot(out, ggplot2::aes_(x = ~ x, y = ~ power)) +
       ggplot2::geom_line() + ggplot2::theme_minimal() +
       ggplot2::xlab(varying)
     print(my_plot)
   }
-  return(power)
+  return(dat)
 }

### 2017-04-04 R/assess_performance.R
Fix power_calc plot option

@@ -191,11 +191,11 @@ check_sims <- function(df, true_rr){
 #'    corresponding power
 #'
 #' @examples
-#' pow <- power_calc(varying = "n", values = c(15, 50, 100), n_reps = 20,
+#' pow <- power_calc(varying = "n", values = c(15, 25, 50, 80, 95, 100), n_reps = 20,
 #'            central = 100, sd=10, rr = 1.01, exposure_type = "continuous",
 #'            exposure_trend = "cos1", exposure_amp = .6, average_outcome=22,
 #'            outcome_trend = "no trend", outcome_amp = .6,
-#'            start.date = "2000-01-01", model = "spline", df_year = 1)
+#'            start.date = "2000-01-01", model = "spline", df_year = 1, plot=TRUE)
 #'
 #' @export
 power_calc <- function(varying, values, n_reps, n=NULL, central, sd, exposure_type,
@@ -237,9 +237,9 @@ power_calc <- function(varying, values, n_reps, n=NULL, central, sd, exposure_ty
   }
   dat <- data.frame(values=values, power=powervec)
   if(plot == TRUE){
-    my_plot <- ggplot2::ggplot(out, ggplot2::aes_(x = ~ x, y = ~ power)) +
+    my_plot <- ggplot2::ggplot(dat, ggplot2::aes_(x = ~ dat$values, y = ~ dat$power)) +
       ggplot2::geom_line() + ggplot2::theme_minimal() +
-      ggplot2::xlab(varying)
+      ggplot2::xlab(varying) + ggplot2::ylab("power")
     print(my_plot)
   }
   return(dat)

### 2017-04-05 writing/EpiDraft.Rmd
Obtain and tidy Legionnaire's data

@@
 The `power_calc` function will also return a plot if `plot=TRUE` is specified.
+
+##Legionnaire's Example
+
+Here is a plot of weekly counts of Legionnaire's disease in the United States from 2014 through 2016. This data was obtained from the CDC website.
+
+```{r, echo=FALSE}
+#I used the variable "Legionellosis, Current week" so hopefully that's right
+leg1 <- read.csv("~/eesim/Legionnaires data/2014Leg.csv")[,c(2,3,24)]
+leg2 <- read.csv("~/eesim/Legionnaires data/2015Leg.csv")[,c(2,3,24)]
+leg3 <- read.csv("~/eesim/Legionnaires data/2016Leg.csv")[,c(2,3,24)]
+leg <- data.frame(year = c(leg1$MMWR.Year, leg2$MMWR.Year, leg3$MMWR.Year), week = c(leg1$MMWR.Week, leg2$MMWR.Week, leg3$MMWR.Week), cases = c(leg1$Legionellosis..Current.week, leg2$Legionellosis..Current.week, leg3$Legionellosis..Current.week))
+plot(leg$cases, xlab = "Week", ylab = "Number of Cases", main = "Legionnaire's Disease 2014-2016")
+```
+
+We can pull the trend from this data:
+
+```{r}
+custom
+```

### 2017-04-17 writing/EpiDraft.Rmd
Scrape one value from CDC webpage like a boss

@@
 
+```{r}
+#scrape just the New York City value for week 14 of 2017 from cdc website:
+library(rvest)
+url <- read_html("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=14&mmwr_table=2J&request=Submit&mmwr_location=")
+nyc <- html_nodes(url, 'tr:nth-child(15) td:nth-child(2)')
+as.numeric(html_text(nyc))
+```
+
 We can pull the trend from this data:

### 2017-04-17 writing/EpiDraft.Rmd
Use purrr to scrape 2017 weeks 1-9

--- 2017-04-05 writing/EpiDraft.Rmd
@@
 
+```{r}
+#scrape just the New York City value for week 14 of 2017 from cdc website:
+library(rvest)
+url <- read_html("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=14&mmwr_table=2J&request=Submit&mmwr_location=")
+#Used Selector Gadget to get css selector
+nyc <- html_nodes(url, 'tr:nth-child(15) td:nth-child(2)')
+as.numeric(html_text(nyc))
+
+#Use purrr to get values for 2017 weeks 1-9
+i <- 1:9
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=0", as.character(.x), "&mmwr_table=2J&request=Submit&mmwr_location=", sep=""))
+html <- url %>% map(~read_html(.x))
+nodes <- html %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)'))
+nycdat <- nodes %>% map(~html_text(.x))
+as.numeric(nycdat)
+
+#There was no table for week 2 so we have an NA.
+```
+
 We can pull the trend from this data:

### 2017-04-17 writing/EpiDraft.Rmd
Simulate Legionnaire-like data

@@
 
+###Actual Data from the CDC
+
 Here is a plot of weekly counts of Legionnaire's disease in the United States from 2014 through 2016. This data was obtained from the CDC website.
@@
 ```
+
+### Simulated Data with Similar Trend
+
+Many epidemiologic data sets are of a sensitive nature and not publicly available.  We can use eesim to simulate data with a similar seasonal trend to fit and assess models.
+
+```{r, echo=F}
+exposure <- continuous_exposure(n=1001, mu = 65, sd = 3, trend = "cos1", amp=.4)
+#Make this into weekly data instead of daily.
+exposuremat <- matrix(data=exposure$x, ncol = 7, nrow=143, byrow=T)
+exposurex <- apply(exposuremat, 1, mean)
+newdate <- seq.int(from=exposure$date[1], to=exposure$date[1001], by=7)
+weeklyexp <- data.frame(date=newdate, x=exposurex)
+plot(weeklyexp, main = "Plot of Exposure values")
+outcome <- sim_outcome(weeklyexp, average_outcome = 10, trend = "linear")
+plot(outcome$outcome)
+
+#Try with eesim function
+
+```

### 2017-04-19 writing/EpiDraft.Rmd
Print data from eesim, legionnaire simulated data

--- 2017-04-17 writing/EpiDraft.Rmd
@@
 ```
+
+### Simulated Data with Similar Trend
+
+Many epidemiologic data sets are of a sensitive nature and not publicly available.  We can use eesim to simulate data with a similar seasonal trend to fit and assess models.
+
+```{r, echo=F}
+#Question: does exposure need to have trend or is it enough to just make outcome look like real data?
+exposure <- continuous_exposure(n=1001, mu = 65, sd = 3, trend = "no trend", amp=.4)
+#Make this into weekly data instead of daily.
+exposuremat <- matrix(data=exposure$x, ncol = 7, nrow=143, byrow=T)
+exposurex <- apply(exposuremat, 1, mean)
+newdate <- seq.int(from=exposure$date[1], to=exposure$date[1001], by=7)
+weeklyexp <- data.frame(date=newdate, x=exposurex)
+plot(weeklyexp, main = "Plot of Exposure values")
+outcome <- sim_outcome(weeklyexp, average_outcome = 10, trend = "cos1linear")
+plot(outcome$outcome)
+
+#Try with eesim function
+leg <- eesim(n_reps=1, n=800, central = 65, sd = 3, exposure_type = "continuous", exposure_trend = "no trend", exposure_amp=.6,average_outcome = 10, outcome_trend = "cos1linear", outcome_amp = .6, rr= 1.02, model = "spline", df_year = 4)
+plot(leg[[3]]$outcome)
+```

### 2017-04-19 R/customizable_functions.R

@@ -447,6 +447,6 @@ eesim <- function(n_reps, n, central = NULL, sd = NULL, exposure_type, exposure_
   mods <- fit_mods(datasims, model, df_year)
   check <- check_sims(df = mods, true_rr = rr)
   totalsims <- list(mods, check)
-  return(totalsims)
+  return(c(totalsims, datasims))
 }
 

### 2017-04-19 writing/EpiDraft.Rmd
Code data scrape

@@
 
-```{r}
+```{r, cache=T}
 #scrape just the New York City value for week 14 of 2017 from cdc website:
@@
 
+#All in one?
+allinone <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=0", as.character(.x), "&mmwr_table=2J&request=Submit&mmwr_location=", sep="")) %>% map(~read_html(.x)) %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x))
+
 #There was no table for week 2 so we have an NA.
+
+#Get ALL THE DATA for NYC 1998-2017 (nothing for 1996 or 7)
+#1998:
+i <- 1:9
+url1to9 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=1998&mmwr_week=0", as.character(.x), "&mmwr_table=2B&request=Submit&mmwr_location=", sep=""))
+i <- 10:53
+url10to53 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=1998&mmwr_week=", as.character(.x), "&mmwr_table=2B&request=Submit&mmwr_location=", sep=""))
+url1998 <- c(url1to9, url10to53)
+nyc1998 <- url1998 %>% map(~read_html(.x)) %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(2)')) %>% map(~html_text(.x))
+as.numeric(nyc1998)
+
+#For the earlier years we only have cumulative data.  Could take differences via code? 2006 is the first year with variable for current week.
+
 ```

### 2017-04-19 writing/EpiDraft.Rmd
Scrape scrape

@@
 
+#MidAtlantic 2006:
+i <- 1:9
+url1to9 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2006&mmwr_week=0", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
+i <- 10:53
+url10to53 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2006&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
+url2006 <- c(url1to9, url10to53)
+ma2006 <- url2006 %>% map(~read_html(.x)) %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x))
+as.numeric(ma2006)
+
 ```

### 2017-04-19 writing/EpiDraft.Rmd
Add changes for script for pulling data

--- 2017-04-19 writing/EpiDraft.Rmd
@@
 
+```{r}
+library(rvest)
+library(purrr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(tidyr)
+library(ggplot2)
+
+## For 2017
+
+i <- str_pad(1:14, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=", as.character(.x), "&mmwr_table=2J&request=Submit&mmwr_location=", sep=""))
+html_2017 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2017 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+nodes_midatlantic <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)'))
+madat <- nodes_midatlantic %>% map(~html_text(.x))
+as.numeric(madat)
+
+nodes_nyc <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)'))
+nycdat <- nodes_nyc %>% map(~html_text(.x))
+as.numeric(nycdat)
+
+nodes_pa <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)'))
+padat <- nodes_pa %>% map(~html_text(.x))
+as.numeric(padat)
+
+leg_cases_2017 <- data_frame(week_end_date = as.character(week_end_date),
+                        midatlantic = as.numeric(madat),
+                        nyc = as.numeric(nycdat),
+                        penn = as.numeric(padat)) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+## For 2016
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2016.asp?mmwr_year=2016&mmwr_week=", as.character(.x), "&mmwr_table=2G&request=Submit&mmwr_location=", sep=""))
+html_2016 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2016 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+nodes_midatlantic <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)'))
+madat <- nodes_midatlantic %>% map(~html_text(.x))
+as.numeric(madat)
+
+nodes_nyc <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)'))
+nycdat <- nodes_nyc %>% map(~html_text(.x))
+as.numeric(nycdat)
+
+nodes_pa <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)'))
+padat <- nodes_pa %>% map(~html_text(.x))
+as.numeric(padat)
+
+leg_cases_2016 <- data_frame(week_end_date = as.character(week_end_date),
+                        midatlantic = as.numeric(madat),
+                        nyc = as.numeric(nycdat),
+                        penn = as.numeric(padat)) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+## For 2015
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2015.asp?mmwr_year=2015&mmwr_week=", as.character(.x), "&mmwr_table=2G&request=Submit&mmwr_location=", sep=""))
+html_2015 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2015 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+nodes_midatlantic <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)'))
+madat <- nodes_midatlantic %>% map(~html_text(.x))
+as.numeric(madat)
+
+nodes_nyc <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)'))
+nycdat <- nodes_nyc %>% map(~html_text(.x))
+as.numeric(nycdat)
+
+nodes_pa <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)'))
+padat <- nodes_pa %>% map(~html_text(.x))
+as.numeric(padat)
+
+leg_cases_2015 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = as.numeric(madat),
+                             nyc = as.numeric(nycdat),
+                             penn = as.numeric(padat)) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+# Putting two years together
+
+leg_cases <- bind_rows(leg_cases_2015, leg_cases_2016, leg_cases_2017)
+
+leg_cases %>% 
+  gather(key = location, value = cases, - week_end_date) %>% 
+  ggplot(aes(x = week_end_date, y = cases, color = location)) + 
+  #geom_point() + 
+  geom_line() + 
+  theme_classic() + 
+  theme(legend.position = "top")
+```
+
 ```{r, cache=T}

### 2017-04-25 writing/EpiDraft.Rmd
Meeting 4/19, only text changes
Scraped legionnaire's data 2006-2017

--- 2017-04-19 writing/EpiDraft.Rmd
@@
 
-```{r, cache=T}
-#scrape just the New York City value for week 14 of 2017 from cdc website:
+##Scraping the data
+
+```{r, cache=TRUE}
 library(rvest)
-url <- read_html("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=14&mmwr_table=2J&request=Submit&mmwr_location=")
-#Used Selector Gadget to get css selector
-nyc <- html_nodes(url, 'tr:nth-child(15) td:nth-child(2)')
-as.numeric(html_text(nyc))
+library(purrr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(tidyr)
+library(ggplot2)
 
-#Use purrr to get values for 2017 weeks 1-9
-i <- 1:9
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=0", as.character(.x), "&mmwr_table=2J&request=Submit&mmwr_location=", sep=""))
-html <- url %>% map(~read_html(.x))
-nodes <- html %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)'))
-nycdat <- nodes %>% map(~html_text(.x))
+## For 2017
+
+i <- str_pad(1:14, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=", as.character(.x), "&mmwr_table=2J&request=Submit&mmwr_location=", sep=""))
+html_2017 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2017 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+nodes_midatlantic <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)'))
+madat <- nodes_midatlantic %>% map(~html_text(.x))
+as.numeric(madat)
+checkma <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x))
+
+nodes_nyc <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)'))
+nycdat <- nodes_nyc %>% map(~html_text(.x))
+as.numeric(nycdat)
+checknyc <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x))
+
+nodes_pa <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)'))
+padat <- nodes_pa %>% map(~html_text(.x))
+as.numeric(padat)
+checkpa <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x))
+
+leg_cases_2017 <- data_frame(week_end_date = as.character(week_end_date),
+                        midatlantic = as.numeric(madat),
+                        nyc = as.numeric(nycdat),
+                        penn = as.numeric(padat), 
+                        checkma = as.character(checkma),
+                        checknyc=as.character(checknyc), 
+                        checkpa=as.character(checkpa)) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+## For 2016
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2016.asp?mmwr_year=2016&mmwr_week=", as.character(.x), "&mmwr_table=2G&request=Submit&mmwr_location=", sep=""))
+html_2016 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2016 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+nodes_midatlantic <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)'))
+madat <- nodes_midatlantic %>% map(~html_text(.x))
+as.numeric(madat)
+checkma <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>%  map(~html_text(.x)) %>% map(~str_trim(.x))
+
+nodes_nyc <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)'))
+nycdat <- nodes_nyc %>% map(~html_text(.x))
+checknyc <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>%  map(~html_text(.x)) %>% map(~str_trim(.x))
 as.numeric(nycdat)
 
-#All in one?
-allinone <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=0", as.character(.x), "&mmwr_table=2J&request=Submit&mmwr_location=", sep="")) %>% map(~read_html(.x)) %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x))
+nodes_pa <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)'))
+padat <- nodes_pa %>% map(~html_text(.x))
+checkpa <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>%  map(~html_text(.x)) %>% map(~str_trim(.x))
+as.numeric(padat)
 
-#There was no table for week 2 so we have an NA.
+leg_cases_2016 <- data_frame(week_end_date = as.character(week_end_date),
+                        midatlantic = as.numeric(madat),
+                        nyc = as.numeric(nycdat),
+                        penn = as.numeric(padat),
+                        checkma = as.character(checkma),
+                        checknyc=as.character(checknyc), 
+                        checkpa=as.character(checkpa)) %>% 
+  mutate(week_end_date = mdy(week_end_date))
 
-#Get ALL THE DATA for NYC 1998-2017 (nothing for 1996 or 7)
-#1998:
-i <- 1:9
-url1to9 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=1998&mmwr_week=0", as.character(.x), "&mmwr_table=2B&request=Submit&mmwr_location=", sep=""))
-i <- 10:53
-url10to53 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=1998&mmwr_week=", as.character(.x), "&mmwr_table=2B&request=Submit&mmwr_location=", sep=""))
-url1998 <- c(url1to9, url10to53)
-nyc1998 <- url1998 %>% map(~read_html(.x)) %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(2)')) %>% map(~html_text(.x))
-as.numeric(nyc1998)
+## For 2015
 
-#For the earlier years we only have cumulative data.  Could take differences via code? 2006 is the first year with variable for current week.
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2015.asp?mmwr_year=2015&mmwr_week=", as.character(.x), "&mmwr_table=2G&request=Submit&mmwr_location=", sep=""))
+html_2015 <- url %>% map(~read_html(.x))
 
-#MidAtlantic 2006:
-i <- 1:9
-url1to9 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2006&mmwr_week=0", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
-i <- 10:53
-url10to53 <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2006&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
-url2006 <- c(url1to9, url10to53)
-ma2006 <- url2006 %>% map(~read_html(.x)) %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x))
-as.numeric(ma2006)
+week_end_date <- html_2015 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2015 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2014
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2014&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
+html_2014 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2014 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2014 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##2013
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2013&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
+html_2013 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2013 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2013 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2012
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2012&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
+html_2012 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2012 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2012 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2011
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2011&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
+html_2011 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2011 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2011 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2010
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2010&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
+html_2010 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2010 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2010 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2009
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2009&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
+html_2009 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2009 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2009 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2008
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2008&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
+html_2008 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2008 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2008 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2007
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2007&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
+html_2007 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2007 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2007 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+##For 2006
+
+i <- str_pad(1:52, pad = "0", width = 2)
+url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2006&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
+html_2006 <- url %>% map(~read_html(.x))
+
+week_end_date <- html_2006 %>% 
+  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
+  map(~html_text(.x)) %>% 
+  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
+  map(~ str_replace(.x[1], "week ending", "")) %>% 
+  map(~ str_replace(.x[1], "\\(", "")) %>% 
+  map(~ str_trim(.x[1]))
+
+madat <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkma <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
+
+nycdat <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checknyc <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+padat <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
+checkpa <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
+
+leg_cases_2006 <- data_frame(week_end_date = as.character(week_end_date),
+                             midatlantic = madat,
+                             nyc = nycdat,
+                             penn = padat, 
+                             checkma = checkma,
+                             checknyc = checknyc,
+                             checkpa = checkpa) %>% 
+  mutate(week_end_date = mdy(week_end_date))
+
+# Putting three years together
+
+leg_cases <- bind_rows(leg_cases_2006, leg_cases_2007, leg_cases_2008, leg_cases_2009, leg_cases_2010, leg_cases_2011, leg_cases_2012, leg_cases_2013, leg_cases_2014, leg_cases_2015, leg_cases_2016, leg_cases_2017)
+
+leg_cases[,1:4] %>% 
+  gather(key = location, value = cases, - week_end_date) %>% 
+  ggplot(aes(x = week_end_date, y = cases, color = location)) + 
+  #geom_point() + 
+  geom_line() + 
+  theme_classic() + 
+  theme(legend.position = "top")
 

### 2017-04-26 writing/EpiDraft.Rmd
Add text to legionnaire's example

@@
 ```{r}
-#Power_calc momentarily out of commission :(
+#Power_calc works except the casecrossover option
 ```
@@
 
-Here is a plot of weekly counts of Legionnaire's disease in the United States from 2014 through 2016. This data was obtained from the CDC website.
+We can perform power analyses on real data to determine whether we have a large enough sample size, relative risk, or average outcome value to provide sufficient power for testing hypotheses.
+Presented here is weekly data from the CDC on mortality related to Legionnaire's disease (citation?).
+We focus on data from New York City, Pennsylvania, and the Midatlantic region from 2006 through 2017.
 
@@
 
-##Scraping the data
+```{r, echo=F, warning=F, message=F}
+#Scraping the data from CDC
 
-```{r, cache=TRUE}
 library(rvest)
@@
 
-# Putting three years together
+# Putting all years together
 
@@
   theme(legend.position = "top")
+
+```
 

### 2017-04-26 writing/EpiDraft.Rmd
Run checks for correct location
Adjusting Leg simulation

@@
 
+```{r}
+#Checking for anomalies:
+leg_cases[which(leg_cases$checkma != "MID. ATLANTIC"),]
+leg_cases[which(leg_cases$checknyc != "N.Y. City"),]
+leg_cases[which(leg_cases$checkpa != "Pa."),]
+
+#Row 48 is the only row with different states. 569 and 574 have NA in every column.
 ```
@@
 ```{r, echo=F}
-#Question: does exposure need to have trend or is it enough to just make outcome look like real data?
-exposure <- continuous_exposure(n=1001, mu = 65, sd = 3, trend = "no trend", amp=.4)
+#Means of real data:
+map(leg_cases[,2:4], mean, na.rm=T)
+#(ma=14.633, nyc=4.69, pa=5.314)
+exposure <- continuous_exposure(n=1001, mu = 65, sd = 5, trend = "cos1", amp=.4)
 #Make this into weekly data instead of daily.
-exposuremat <- matrix(data=exposure$x, ncol = 7, nrow=143, byrow=T)
-exposurex <- apply(exposuremat, 1, mean)
-newdate <- seq.int(from=exposure$date[1], to=exposure$date[1001], by=7)
-weeklyexp <- data.frame(date=newdate, x=exposurex)
+x <- seq.int(from=1, to=1001, by=7)
+weeklyexp <- exposure[x,]
 plot(weeklyexp, main = "Plot of Exposure values")
-outcome <- sim_outcome(weeklyexp, average_outcome = 10, trend = "cos1linear")
+outcome <- sim_outcome(weeklyexp, average_outcome = 4.69, trend = "no trend")
 plot(outcome$outcome)
 
-#Try with eesim function
-leg <- eesim(n_reps=1, n=800, central = 65, sd = 3, exposure_type = "continuous", exposure_trend = "no trend", exposure_amp=.6,average_outcome = 10, outcome_trend = "cos1linear", outcome_amp = .6, rr= 1.02, model = "spline", df_year = 4)
-plot(leg[[3]]$outcome)
 ```

### 2017-05-02 writing/EpiDraft.Rmd
How to get exposure data?
Set all code to not evaluate while working on the text of the Epi draft
Work some on section on Legionnaires' disease

--- 2017-04-25 writing/EpiDraft.Rmd
@@
 
-#Introduction
+```{r echo = FALSE}
+knitr::opts_chunk$set(eval = FALSE)
+```
+
+# Introduction
 
@@
 
-#Simulating Data and Testing Models
+# Simulating Data and Testing Models
 
@@
 
-#Power Analysis
+# Power Analysis
 
@@
 
-##Legionnaire's Example
+## Legionnaire's Example
 
-###Actual Data from the CDC
+As a case study, we give an example of conducting a power analysis for a planned study of the relationship between Legionnaires' disease (LD) and extreme precipitation in a community.
+
+[Description of LD]
+The bacteria that causes LD, *Legionella*, is often present in ... . 
+[Where Lp live-- water systems, systems for A/C, ...]
+[What Lp need to grow-- protozoa (must grow inside them?), temperature conditions, water, water temperature?]
+[What Lp need to spread-- survival on aerosols, lower UV, enough wind to spread aerosols but not to break up aerosols, high relative humidity]
+The incubation period for LD is 2--10 days (this is relevant for the time between dispersal of Lp on aerosols and case onset) [@Halsby2014] [also could reference Campese et al.
+2011, Farnham et al.
+2014, Graham et al.
+2011, Whiley et al.
+2014, Phin et al.
+2014, Brandsema et al.
+2014, Dunn et al.
+2012, Fisman et al.
+2005, Halsby et al.
+2013].
+[Destinction between community-acquired LD and other LD cases; sporadic vs. epidemic]
+Based on previous studies, approximately 2--6% of LD cases are outbreak-associated, rather than sporadic [Campese et al.
+2011, Graham et al.
+2011, Fisman et al.
+2005, Halsby et al.
+2013, Phin et al.
+2014].
+Community counts of sporadic, community-acquired LD cases have been found to display a strong seasonal pattern, with most cases in summer to early fall [@Halsby2014; Campese et al.
+2011; Falkinham et al.
+2015; Farnham et al.
+2014; Phin et al.
+2014; Brandsema et al.
+2014].
+This seasonal trend may stem in part from seasonal variaition in weather factors, but may also be related to other seasonally-varying factors.
+Counts of LD cases have also shown a strong long-term trend in many locations, including in England and Wales between 1993 and 2008 [@Halsby2014], in France between 1998 and 2008 [Campese et al.
+2011], and in the United States between 2000 and 2011 [Falkinham et al.
+2015], ...
+.
+This long-term trend may in part reflect a growing improvement in reporting and testing [?] of potential cases [ref]; in fact, one study in Australia notes that the total incidence of LD cases might be a factor of 20 higher than the number of reported cases because of severe underreporting [Whiley et al.
+2014].
+
+[Why we think LD might be linked with weather] 
+[Theoretical reasons]
+Weather could affect risk of LD by affecting either the growth or the spread, via aerosol dispersion [?], of the *Legionella* bacteria that causes the disease.
+[Why weather factor might affect growth of Lp] [Temperature could affect growth of Lp, including through a chain that increases growth of what protozoa eat, since Lp need to grow in protozoa?] A number of weather-related factors, including higher relative humidity and lower exposure to sunlight [?], might aid in survival of *Legionella* on aerosols [check Berendt 1980, Survival of Legionella pneumophila in aerosols...; Hambleton et al.
+1983, Susceptibility of Lp to ultraviolet radiation; check for anything more current about survival of Lp on aerosols]
+[Evidenced on sporadic cases from epi studies]
+Several studies have investigated the association between weather conditions and daily counts of community-acquired LD cases using time series models.
+Many of these have found an increased risk of community-acquired LD following certain weather conditions; weather variables implicated include rainfall, wind speed, sunlight, relative humidity.
+For example, a study of community-acquired LD and weather in England and Wales found increased risk of LD associated with higher temperatures (at lags of 1--9 weeks; estimated odds ratio for 95th percentile versus 75th percentile temperature of 3.91), higher relative humidity (at lags of 2--10 days; estimated odds ratio for 75th versus 50th percentile relative humidity of 1.52), and higher rainfall (at lags of 2--10 days; estimated odds ratio for 75th versus 50th percentile rainfall of 1.78) [@Halsby2014].
+A study of sporadic LD cases in the Netherlands between 2003 and 2011 found increased risk associated with warm temperature (4-week mean) and increased rainfall (2-week duration and intensity) [Brandsema et al.
+2014].
+A study in Taiwan between 1995 and 2011 found an increase in LD cases associated with increased temperature (lag day 10-11) and increased precipitation (lag day 11), with an increased risk of LD cases of 2.6 associated with heavy versus no rainfall eleven days prior [Nai-Tzu et al.
+2014].
+A study of LD cases in the greater Philadelphia metropolitan area between 1995 and 2003 found increased risk of LD following about eight days following precipitation (compared to no precipitation) as well as increased risk associated with higher relative humidity and higher temperature and decreased risk associated with a higher wind speed (at a similar lag?) [double-check.
+Fisman et al.
+2005].
+[Evidence on outbreak cases from epi studies]
+The risk of outbreaks of LD might also be related to weather conditions, including extreme precipitation (e.g., [check Hoyle et al.
+1985 Legionnaires' disease: seeking a wider source; Thacker et al.
+1978, An outbreak in 1965 of severe respiratory illness caused by ...]).
+
+[Why power might be an issue in a study of LD and weather] 
+While the epidemiological estimates of the association between LD and weather variables could help in designing strategies to prevent cases of the disease or identify periods of heightened risk, the power to do such an analysis may be low in many communities because LD is typically a fairly rare disease.
+For example, a study of all sporadic community-acquired LD cases throughout England and Wales between 1993 and 2008 included only 1,676 cases after excluding cases that were possible travel-acquired and other data quality-related exclusions [@Halsby2014].
+Several studies have used time series [Dunn et al.
+2012] or case-crossover [@Halsby2014, Nai-Tzu et al.
+2014, Fisman et al.
+2005] designs to investigate the association between weather variables and community-wide risk of LD.
+These, however, have mostly been conducted on very large study populations (e.g., England and Wales[@Halsby2014]; Netherlands [Brandsema et al.
+2014]; Taiwan [Nai-Tzu et al.
+2014]).
+Even in these large populations, associations with some weather variables were suggestive but not statistically significant (e.g., for windspeed and UV in England and Wales; @Halsby2014).
+One study was conducting in a smaller population (the greater Philadelphia metropolitan area; Fisman et al.
+2005).
+Some of these studies have controlled for season [@Halsby2014; Brandsema et al.
+2014, Nai-Tzu et al.
+2014, Fisman et al.
+2005] while some have not [?].
+
+Since it is unclear if the relationship between LD risk and weather variables is linear, many of these studies have investigated either categorical versions of the weather variables [e.g., Brandsema et al.
+2014, Nai-Tzu et al.
+2014] or estimated risks at two percentiles along the weather variable's full distribution (e.g., @Halsby2014).
+
+### Actual Data from the CDC
 

### 2017-05-02 R/customizable_functions.R
sd=1 default in create_sims

@@ -351,7 +351,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #'
 #' @export
 #'
-create_sims <- function(n_reps, n, central, sd, exposure_type, exposure_trend,
+create_sims <- function(n_reps, n, central, sd=1, exposure_type, exposure_trend,
                         exposure_amp, average_outcome, outcome_trend,
                         outcome_amp, rr, start.date = "2000-01-01",
                         cust_exp_func = NULL, cust_exp_args = NULL,

### 2017-05-02 R/customizable_functions.R
Add slope option for linear trends

@@ -69,7 +69,7 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
 #' ggplot(x_cust, aes(x = date, y = x)) + geom_point()
 #' @export
 sim_exposure <- function(n, central = NULL, sd = NULL, trend = "no trend",
-                         amp = .6, exposure_type = NULL,
+                         slope = 1, amp = .6, exposure_type = NULL,
                          start.date = "2001-01-01", cust_exp_func = NULL,
                          cust_exp_args = NULL){
   start.date <- as.Date(start.date)
@@ -88,6 +88,7 @@ sim_exposure <- function(n, central = NULL, sd = NULL, trend = "no trend",
     }
     arguments$central <- central
     arguments$trend <- trend
+    arguments$slope <- slope
     arguments$amp <- amp
     arguments$exposure_type <- exposure_type
     arguments$sd <- sd
@@ -159,6 +160,8 @@ custom_baseline <- function(n, df = dlnm::chicagoNMMAPS, outcome_type = "cvd",
 #'      \item{"no trend"}
 #'      \item{"custom"}
 #'    }
+#' @param slope slope A numeric value specifying the slope of the trend, to be used
+#'    with trend = "linear" or trend = "cos1linear".
 #' @param amp A numeric value specifying the amplitude of the seasonal trend.
 #'    Must be between 0 and 1.
 #' @param ... optional arguments to a custom baseline function
@@ -175,11 +178,12 @@ custom_baseline <- function(n, df = dlnm::chicagoNMMAPS, outcome_type = "cvd",
 #'
 #' @export
 #'
-create_baseline <- function(n, average_baseline, trend, amp,
+create_baseline <- function(n, average_baseline, trend, slope, amp,
                             cust_base_func = NULL, ...){
   if(is.null(cust_base_func)){
     lambda <- average_baseline
-    baseline <- sim_baseline(n=n, lambda=lambda, trend=trend, amp=amp)
+    baseline <- sim_baseline(n=n, lambda=lambda, trend=trend, slope=slope,
+                             amp=amp)
   } else {
     arguments <- list(...)
     arguments$n <- n
@@ -208,7 +212,8 @@ create_baseline <- function(n, average_baseline, trend, amp,
 #' @return A numeric vector of mean outcome values
 #'
 #' @examples
-#' base <- create_baseline(n = 5, average_outcome = 22, trend = "linear")
[…]

### 2017-05-02 R/sim_exposure.R

@@ -17,6 +17,8 @@
 #'    }
 #'    See the package vignette for examples of the shapes of these trends.
 #'    Default is "no trend".
+#' @param slope A numeric value specifying the slope of the trend, to be used
+#'    with trend = "linear" or trend = "cos1linear".
 #' @param amp A numeric value specifying the amplitude of the seasonal trend.
 #'    Must be between 0 and 1.
 #' @param custom_func A character string specifying a customized function from
@@ -29,7 +31,7 @@
 #' calc_t(5, "cos3", amp = .5)
 #'
 #' @export
-calc_t <- function(n, trend = "no trend", amp = .6, custom_func = NULL, ...){
+calc_t <- function(n, trend = "no trend", slope=1, amp = .6, custom_func = NULL, ...){
   day <- c(1:n)
   if (trend == "cos1"){
     seasont <- 1 + amp * cos(2 * pi * (day / 365))
@@ -39,11 +41,11 @@ calc_t <- function(n, trend = "no trend", amp = .6, custom_func = NULL, ...){
     } else if (trend == "cos3"){
       seasont <- 1 + .75 ^ (day / 365) * amp * cos(2 * pi * (day / 365))
     } else if (trend == "linear"){
-      seasont <- 1 + (day / n)
+      seasont <- 1 + slope*(day / n)
     } else if (trend == "curvilinear"){
       seasont <- 1 + day * (2 / n) + day^2 * (-1 / n^2)
     } else if (trend == "cos1linear"){
-      seasont <- (1 + (day / n)) * (1 + amp * cos(2 * pi * (day / 365)))
+      seasont <- (1 + slope*(day / n)) * (1 + amp * cos(2 * pi * (day / 365)))
     } else if (trend == "no trend"){
       seasont <- rep(1, n)
     } else if (trend == "custom" & !is.null(custom_func)) {
@@ -75,6 +77,8 @@ calc_t <- function(n, trend = "no trend", amp = .6, custom_func = NULL, ...){
 #'    }
 #'
 #' @param p A numeric value giving the mean probability of exposure
+#' @param slope A numeric value specifying the slope of the trend, to be used
+#'    with trend = "linear" or trend = "cos1linear".
 #' @param amp A numeric value specifying the amplitude of the seasonal trend.
 #'    Must be between 0 and .5.
 #' @param start.date A date of the format "yyyy-mm-dd" from which to begin
[…]

### 2017-05-02 writing/EpiDraft.Rmd
Turn code chunk evaluation back on for the moment
sd=1 default in create_sims

@@
 ```{r echo = FALSE}
-knitr::opts_chunk$set(eval = FALSE)
+# knitr::opts_chunk$set(eval = FALSE)
 ```
@@
   theme(legend.position = "top")
+
+save(leg_cases, file="data/leg_cases.rda")
+
+```
+
+```{r}
+#Checking for anomalies:
+leg_cases[which(leg_cases$checkma != "MID. ATLANTIC"),]
+leg_cases[which(leg_cases$checknyc != "N.Y. City"),]
+leg_cases[which(leg_cases$checkpa != "Pa."),]
+
+#Row 48 is the only row with different states. 569 and 574 have NA in every column.
+```
+
+```{r}
+#Power analysis using eesim
 

### 2017-05-02 R/customizable_functions.R
Clarify documentation of sd argument

@@ -305,7 +305,8 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #'        to be performed
 #' @param n A numeric value specifying the number of days to simulate.
 #' @param sd A numeric value giving the standard deviation of the exposure
-#'    values around the trend line.
+#'    values from the exposure trend line (not the total standard deviation of
+#'    the exposure values).
 #' @param exposure_trend A character string specifying the seasonal trend for
 #'        exposure. Options for continuous exposure are:
 #'        \itemize{

### 2017-05-02 R/sim_exposure.R

@@ -165,7 +165,8 @@ binary_exposure <- function(n, p, trend = "no trend", slope, amp,
 #'
 #' @param mu A numeric value giving the mean exposure.
 #' @param sd A numeric value giving the standard deviation of the exposure
-#'    values around the trend line.
+#'    values from the exposure trend line (not the total standard deviation of
+#'    the exposure values).
 #' @inheritParams calc_t
 #' @inheritParams binary_exposure
 #'

### 2017-05-02 R/assess_performance.R
Add custom model option in fit_mods and eesim

@@ -76,7 +76,7 @@ beta_var <- function(df){
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.01)
-#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, model = "spline", df_year = 1)
 #' beta_bias(fits, true_rr = 1.02)
 #'
 #' @export
@@ -100,12 +100,12 @@ beta_bias <- function(df, true_rr){
 #' @examples
 #' sims <- create_sims(n_reps = 10, n = 600, central = 100,
 #'                     sd = 10, exposure_type = "continuous",
-#'                     exposure_trend = "cos1",
+#'                     exposure_trend = "cos1", exposure_slope=1,
 #'                     exposure_amp = 0.6,
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.01)
-#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, model = "spline", df_year = 1)
 #' coverage_beta(df=fits, true_rr = 1.02)
 #'
 #' @export
@@ -132,7 +132,7 @@ coverage_beta <- function(df, true_rr){
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.01)
-#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, model = "spline", df_year = 1)
 #' power_beta(fits)
 #'
 #' @export
@@ -159,7 +159,7 @@ power_beta <- function(df){
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.02)
-#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, model = "spline", df_year = 1)
 #' check_sims(df = fits, true_rr = 1.02)
 #'
 #' @export

### 2017-05-02 R/customizable_functions.R

@@ -363,8 +363,8 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #' @export
 #'
 create_sims <- function(n_reps, n, central, sd=1, exposure_type, exposure_trend,
-                        exposure_slope, exposure_amp, average_outcome,
-                        outcome_trend, outcome_slope,
+                        exposure_slope=1, exposure_amp, average_outcome,
+                        outcome_trend, outcome_slope=1,
                         outcome_amp, rr, start.date = "2000-01-01",
                         cust_exp_func = NULL, cust_exp_args = NULL,
                         cust_base_func = NULL, cust_lambda_func = NULL,
@@ -388,10 +388,17 @@ create_sims <- function(n_reps, n, central, sd=1, exposure_type, exposure_trend,
 
 #' Fit models
 #'
-#' @param outcome A list of simulated data sets which each include columns
-#'    called "x" and "outcome"
+#' @param data A list of simulated data sets which each include columns
+#'    called "x" for exposure values and "outcome" for outcome values
 #' @param model A character string specifying model to be used. Choices are
 #'    "spline" and "casecrossover"
+#' @param custom_model A character string for the name of the custom model. This
+#'    is a user-created function which takes a data frame with columns "x" for
+#'    exposure values and "outcome" for outcome values. It must output a data
+#'    frame with columns called "Estimate", "Std. Error", "t value", "Pr(>|t|)",
+#'    "2.5 %", and "97.5 %", the output from summary() and confint()
+#' @param custom_model_args A list of arguments and their values for a custom
+#'    model.
 #' @inheritParams spline_mod
 #'
 #' @return A data frame in which each row includes an estimate of beta hat,
@@ -403,15 +410,20 @@ create_sims <- function(n_reps, n, central, sd=1, exposure_type, exposure_trend,
 #'             exposure_type="continuous", exposure_trend = "cos1",
 #'             exposure_amp = .6, average_outcome = 22,
 #'             outcome_trend = "no trend", outcome_amp = .6, rr = 1.01)
-#' fit_mods(outcome = sims, model = "spline")
+#' fit_mods(data = sims, model = "spline")
 #'
 #' @export
-fit_mods <- function(outcome, model, df_year = 7){
+fit_mods <- function(data, model=NULL, df_year = 7, custom_model = NULL,
+                     custom_model_args = list()){
[…]

### 2017-05-02 R/fit_models.R

@@ -2,7 +2,8 @@
 #'
 #' This function fits a GLM with splines to your data.
 #'
-#' @param df A data frame with exposure and outcome data
+#' @param df A data frame with exposure and outcome data called "x" and "outcome",
+#'    respectively
 #' @param df_year A numeric value specifying the degrees of freedom per year
 #'
 #' @return Summary of the estimated log relative risk for a 1-unit increase in

### 2017-05-03 writing/EpiDraft.Rmd
Upload precipitation data from Rachel
Aggregate rainfall data

@@
 ```{r, echo=F, warning=F, message=F}
-#Scraping the data from CDC
-
-library(rvest)
-library(purrr)
-library(dplyr)
-library(stringr)
-library(lubridate)
-library(tidyr)
+#plot of CDC data
 library(ggplot2)
-
-## For 2017
-
-i <- str_pad(1:14, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2017.asp?mmwr_year=2017&mmwr_week=", as.character(.x), "&mmwr_table=2J&request=Submit&mmwr_location=", sep=""))
-html_2017 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2017 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-nodes_midatlantic <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)'))
-madat <- nodes_midatlantic %>% map(~html_text(.x))
-as.numeric(madat)
-checkma <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x))
-
-nodes_nyc <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)'))
-nycdat <- nodes_nyc %>% map(~html_text(.x))
-as.numeric(nycdat)
-checknyc <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x))
-
-nodes_pa <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)'))
-padat <- nodes_pa %>% map(~html_text(.x))
-as.numeric(padat)
-checkpa <- html_2017 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x))
-
-leg_cases_2017 <- data_frame(week_end_date = as.character(week_end_date),
-                        midatlantic = as.numeric(madat),
-                        nyc = as.numeric(nycdat),
-                        penn = as.numeric(padat), 
-                        checkma = as.character(checkma),
-                        checknyc=as.character(checknyc), 
-                        checkpa=as.character(checkpa)) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-## For 2016
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2016.asp?mmwr_year=2016&mmwr_week=", as.character(.x), "&mmwr_table=2G&request=Submit&mmwr_location=", sep=""))
-html_2016 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2016 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-nodes_midatlantic <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)'))
-madat <- nodes_midatlantic %>% map(~html_text(.x))
-as.numeric(madat)
-checkma <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>%  map(~html_text(.x)) %>% map(~str_trim(.x))
-
-nodes_nyc <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)'))
-nycdat <- nodes_nyc %>% map(~html_text(.x))
-checknyc <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>%  map(~html_text(.x)) %>% map(~str_trim(.x))
-as.numeric(nycdat)
-
-nodes_pa <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)'))
-padat <- nodes_pa %>% map(~html_text(.x))
-checkpa <- html_2016 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>%  map(~html_text(.x)) %>% map(~str_trim(.x))
-as.numeric(padat)
-
-leg_cases_2016 <- data_frame(week_end_date = as.character(week_end_date),
-                        midatlantic = as.numeric(madat),
-                        nyc = as.numeric(nycdat),
-                        penn = as.numeric(padat),
-                        checkma = as.character(checkma),
-                        checknyc=as.character(checknyc), 
-                        checkpa=as.character(checkpa)) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-## For 2015
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_2015.asp?mmwr_year=2015&mmwr_week=", as.character(.x), "&mmwr_table=2G&request=Submit&mmwr_location=", sep=""))
-html_2015 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2015 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2015 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2015 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2014
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2014&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
-html_2014 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2014 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2014 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2014 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##2013
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2013&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
-html_2013 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2013 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2013 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2013 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2012
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2012&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
-html_2012 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2012 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2012 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2012 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2011
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2011&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
-html_2011 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2011 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2011 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2011 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2010
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2010&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
-html_2010 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2010 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2010 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2010 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2009
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2009&mmwr_week=", as.character(.x), "&mmwr_table=2F&request=Submit&mmwr_location=", sep=""))
-html_2009 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2009 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(12) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(15) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(2)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2009 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2009 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2008
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2008&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
-html_2008 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2008 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2008 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2008 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2007
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2007&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
-html_2007 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2007 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2007 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2007 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-##For 2006
-
-i <- str_pad(1:52, pad = "0", width = 2)
-url <- i %>% map(~paste("https://wonder.cdc.gov/mmwr/mmwr_1995_2014.asp?mmwr_year=2006&mmwr_week=", as.character(.x), "&mmwr_table=2C&request=Submit&mmwr_location=", sep=""))
-html_2006 <- url %>% map(~read_html(.x))
-
-week_end_date <- html_2006 %>% 
-  map(~html_nodes(.x, 'tr:nth-child(1) td:nth-child(1)')) %>% 
-  map(~html_text(.x)) %>% 
-  map(~ str_extract(.x[1], "week ending .+ \\(")) %>% 
-  map(~ str_replace(.x[1], "week ending", "")) %>% 
-  map(~ str_replace(.x[1], "\\(", "")) %>% 
-  map(~ str_trim(.x[1]))
-
-madat <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkma <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(13) td:nth-child(1)')) %>% map(~html_text(.x)) %>% map(~str_trim(.x)) %>% as.character()
-
-nycdat <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checknyc <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(16) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-padat <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(12)')) %>% map(~html_text(.x)) %>% as.numeric()
-checkpa <- html_2006 %>% map(~html_nodes(.x, 'tr:nth-child(17) td:nth-child(1)')) %>% map(~html_text(.x))%>% map(~str_trim(.x)) %>% as.character()
-
-leg_cases_2006 <- data_frame(week_end_date = as.character(week_end_date),
-                             midatlantic = madat,
-                             nyc = nycdat,
-                             penn = padat, 
-                             checkma = checkma,
-                             checknyc = checknyc,
-                             checkpa = checkpa) %>% 
-  mutate(week_end_date = mdy(week_end_date))
-
-# Putting all years together
-
-leg_cases <- bind_rows(leg_cases_2006, leg_cases_2007, leg_cases_2008, leg_cases_2009, leg_cases_2010, leg_cases_2011, leg_cases_2012, leg_cases_2013, leg_cases_2014, leg_cases_2015, leg_cases_2016, leg_cases_2017)
 
@@
   theme(legend.position = "top")
-
-save(leg_cases, file="data/leg_cases.rda")
 
@@
 #Row 48 is the only row with different states. 569 and 574 have NA in every column.
+```
+
+```{r}
+library(lubridate)
+library(dplyr)
+#Precipitation data for New York City and Philadelphia:
+parain <- readRDS("~/eesim/writing/data/pa.rds")
+nyrain <- readRDS("~/eesim/writing/data/pa.rds")
+paquant <- quantile(parain$prcp, probs = c(.75, .95, .99), na.rm=T)
+nyquant <- quantile(nyrain$prcp, probs = c(.75, .95, .99), na.rm=T)
+parain$quant75 <- ifelse(parain$prcp>paquant[1], 1,0)
+parain$quant95 <- ifelse(parain$prcp>paquant[2], 1,0)
+parain$quant99 <- ifelse(parain$prcp>paquant[3], 1,0)
+nyrain$quant75 <- ifelse(nyrain$prcp>nyquant[1], 1,0)
+nyrain$quant95 <- ifelse(nyrain$prcp>nyquant[2], 1,0)
+nyrain$quant99 <- ifelse(nyrain$prcp>nyquant[3], 1,0)
+parain$month <- month(parain$date)
+parain$year <- year(parain$date)
+nyrain$month <- month(nyrain$date)
+nyrain$year <- year(nyrain$date)
+#Monthly rainfall above 75th, 95th, and 99th percentiles for Philly and NYC
+pamonthly <- group_by(parain, month, year) %>% summarise(quant75 =mean(quant75), quant95=mean(quant95), quant99=mean(quant99))
+nymonthly <- group_by(nyrain, month, year) %>% summarise(mean(quant75), mean(quant95), mean(quant99))
+ggplot(pamonthly, aes(x=1:136, y=quant95))+geom_line()
 ```

### 2017-05-03 R/customizable_functions.R
Make some edits to helpfile documentation

@@ -301,47 +301,69 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 
 #' Create simulated data for many repetitions
 #'
-#' @param n_reps A numeric value specifying the number of simulation repititions
-#'        to be performed
-#' @param n A numeric value specifying the number of days to simulate.
+#' @param n_reps A numeric value specifying the number of datasets to simulate (e.g.,
+#'        \code{n_reps = 1000} would simulate one thousand time series datasets with the
+#'        specified characteristics, which can be used for a power analysis or to investigate
+#'        the performance of a proposed model).
+#' @param n A numeric value specifying the number of days to simulate (e.g., \code{n = 365}
+#'        would simulate a dataset with a year's worth of data).
 #' @param sd A numeric value giving the standard deviation of the exposure
-#'    values from the exposure trend line (not the total standard deviation of
-#'    the exposure values).
-#' @param exposure_trend A character string specifying the seasonal trend for
-#'        exposure. Options for continuous exposure are:
+#'        values from the exposure trend line (not the total standard deviation of
+#'        the exposure values).
+#' @param exposure_trend A character string specifying a seasonal and / or long-term trend for
+#'        expected mean exposure. See the vignette for \code{eesim} for examples of each option.
+#'        The shapes are based on those used in Bateson and Schwartz (1999).
+#'        For trends with a seasonal component, the amplitude of the seasonal trend can be
+#'        customized using the \code{exposure\_amp} argument. For trends with a long-term
+#'        pattern, the slope of the long-term trend can be set using the \code{exposure\_slope}
+#'        argument. If using the "custom" option, you must ... .
+#'        If using the "monthly" option for a binary exposure, you must input a numeric
+#'        vector of length 12 for the \code{central} argument that gives the probability of
+#'        exposure for each month, starting in January and ending in December.
+#'        Options for continuous exposure are:
 #'        \itemize{
-#'      \item{"cos1"}
-#'      \item{"cos2"}
-#'      \item{"cos3"}
-#'      \item{"linear"}
-#'      \item{"curvilinear"}
-#'      \item{"cos1linear"}
-#'      \item{"no trend"}
-#'      \item{"custom"}
+#'      \item{"cos1": A seasonal trend only.}
+#'      \item{"cos2": A seasonal trend with variable amplitude across years.}
[…]

### 2017-05-03 writing/EpiDraft.Rmd

@@
 #plot of CDC data
+library(rvest)
+library(purrr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(tidyr)
 library(ggplot2)
 
+load("~/eesim/writing/data/leg_cases.rda")
 leg_cases[,1:4] %>% 
@@
 
-```{r}
+```{r, echo=F, eval=F}
 #Checking for anomalies:
@@
 nymonthly <- group_by(nyrain, month, year) %>% summarise(mean(quant75), mean(quant95), mean(quant99))
-ggplot(pamonthly, aes(x=1:136, y=quant95))+geom_line()
+#What pattern is this?
+ggplot(pamonthly, aes(x=1:136, y=quant95))+geom_line()+geom_vline(xintercept=seq(from=1, to=136, by=12), col="red") + ggtitle("Monthly rainfall for Philadelphia 2006-2017")+ ylab("Proportion of days over 95th percentile")
 ```
@@
 ```{r}
-custom
 ```

### 2017-05-13 R/customizable_functions.R
Work some on `eesim` documentation, give names to output, change order of output in list, and make simulated datasets a list within the list (rather than adding each as a separate list element)

@@ -301,6 +301,12 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 
 #' Create simulated data for many repetitions
 #'
+#' Creates a collection of synthetic datasets that follow a set of user-specified
+#' conditions (e.g., exposure mean and variance, average daily outcome count,
+#' long-term and seasonal trends in exposure and outcome, association between exposure
+#' and outcome). These synthetic datasets can be used to investigate performance of a specific
+#' model or to estimate power or required sample size for a hypothetical study.
+#'
 #' @param n_reps A numeric value specifying the number of datasets to simulate (e.g.,
 #'        \code{n_reps = 1000} would simulate one thousand time series datasets with the
 #'        specified characteristics, which can be used for a power analysis or to investigate
@@ -373,8 +379,9 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #' @inheritParams std_exposure
 #' @inheritParams sim_outcome
 #'
-#' @return A list resulting from repetitions of simulations with data frames
-#'    for date, exposure, and outcomes, and estimates from fitting models
+#' @return A list object, in which each list element is one of the synthetic datasets
+#'    simulated under the input conditions. Each synthetic dataset includes columns for
+#'    for date, daily exposure (\code{x}), and daily outcome count.
 #'
 #' @references
 #'
@@ -474,8 +481,28 @@ fit_mods <- function(data, model=NULL, df_year = 7, custom_model = NULL,
 #' @inheritParams std_exposure
 #' @inheritParams fit_mods
 #'
-#' @return A list object with summaries of each model fitted on the simulated
-#'    data sets and measures of model evaluation including coverage and power
+#' @return A list object with three elements:
+#' \itemize{
+#'   \item{\code{simulated_datasets}: }{A list in which each element is a data frame with
+#'     one of the simulated time series datasets, created according to the specifications
+#'     set by the user.}
+#'   \item{\code{indiv_performance}: }{A dataframe with one row per simulated dataset (i.e.,
+#'     total number of rows equal to \code{n_reps}). Each row gives the results of fitting the
+#'     specified model to one of the simulated datasets and includes, for the model fit to that
[…]

### 2017-05-14 writing/EpiDraft.Rmd

@@
 #I used the variable "Legionellosis, Current week" so hopefully that's right
-leg1 <- read.csv("~/eesim/Legionnaires data/2014Leg.csv")[,c(2,3,24)]
-leg2 <- read.csv("~/eesim/Legionnaires data/2015Leg.csv")[,c(2,3,24)]
-leg3 <- read.csv("~/eesim/Legionnaires data/2016Leg.csv")[,c(2,3,24)]
+leg1 <- read.csv("../Legionnaires data/2014Leg.csv")[,c(2,3,24)]
+leg2 <- read.csv("../Legionnaires data/2015Leg.csv")[,c(2,3,24)]
+leg3 <- read.csv("../Legionnaires data/2016Leg.csv")[,c(2,3,24)]
 leg <- data.frame(year = c(leg1$MMWR.Year, leg2$MMWR.Year, leg3$MMWR.Year), week = c(leg1$MMWR.Week, leg2$MMWR.Week, leg3$MMWR.Week), cases = c(leg1$Legionellosis..Current.week, leg2$Legionellosis..Current.week, leg3$Legionellosis..Current.week))
@@
 
-load("~/eesim/writing/data/leg_cases.rda")
+load("../writing/data/leg_cases.rda")
 leg_cases[,1:4] %>% 
@@
 #Precipitation data for New York City and Philadelphia:
-parain <- readRDS("~/eesim/writing/data/pa.rds")
-nyrain <- readRDS("~/eesim/writing/data/pa.rds")
+parain <- readRDS("data/pa.rds")
+nyrain <- readRDS("data/pa.rds")
 paquant <- quantile(parain$prcp, probs = c(.75, .95, .99), na.rm=T)

### 2017-05-14 R/sim_exposure.R
Fix error message for bin_t

@@ -80,7 +80,7 @@ calc_t <- function(n, trend = "no trend", slope=1, amp = .6, custom_func = NULL,
 #' @param slope A numeric value specifying the slope of the trend, to be used
 #'    with trend = "linear" or trend = "cos1linear".
 #' @param amp A numeric value specifying the amplitude of the seasonal trend.
-#'    Must be between 0 and .5.
+#'    Must be between -.5 and .5.
 #' @param start.date A date of the format "yyyy-mm-dd" from which to begin
 #'    simulating values
 #' @inheritParams calc_t
@@ -99,11 +99,11 @@ bin_t <- function(n, p, trend = "no trend", slope = 1, amp = .01,
   date <- seq(from = start.date, by = 1, length.out = n)
   if (trend == "monthly"){
   }
-  else if (p > .5 & amp >1-p & !(trend == "no trend")){
-    stop(paste0("For p>.5, amp must be between 0 and 1-p."))
+  else if (abs(p) > .5 & abs(amp) >1-p & !(trend == "no trend")){
+    stop(paste0("For abs(p)>.5, amp must be between -(1-p) and 1-p."))
   }
-  else if (p<.5 & amp >p & !(trend == "no trend")){
-    stop(paste0("For p<.5, amp must be between 0 and p."))
+  else if (abs(p)<.5 & abs(amp) >p & !(trend == "no trend")){
+    stop(paste0("For abs(p)<.5, amp must be between -p and p."))
   }
   if (trend == "cos1"){
     seasont <- p + amp * cos(2 * pi * (day / 365))

### 2017-05-15 writing/custom_examples.Rmd
Add custom model option in fit_mods and eesim
Get rid of "custom" trend option in calc_t and bin_t

--- 2016-08-23 custom_examples.Rmd
@@
   day <- c(1:n)
-  base <- 1 + y * sin(2 * pi * (day / 365))
-  rnorm(n, mean = base, sd = 1)
+  1 + y * sin(2 * pi * (day / 365))
 }
-calc_t(n = 5, y = .3, trend = "custom", custom_func = sintrend)
-plot(calc_t(1000, y = 2, trend = "custom", custom_func = sintrend))
+binsin <- function(n, y, p){
+  day <- c(1:n)
+  seasont <- p + y * sin(2 * pi * (day / 365))
+}
+calc_t(n = 5, y = .3, custom_func = sintrend)
+plot(calc_t(1000, y = 2, custom_func = sintrend))
+plot(bin_t(1000, p=.3, trend="cos1"))
+plot(bin_t(1000, p=.3, custom_func=binsin, y=.1))
 ```
@@
 ```{r}
-binary_exposure(n=5, p = .2, y = .1, trend = "custom", custom_func = sintrend)
-plot(binary_exposure(n=400, p = .2, y = .1, trend = "custom", custom_func = sintrend))
+binary_exposure(n=5, p = .2, y = .1, amp=.1, custom_func = sintrend)
+plot(binary_exposure(n=1000, p = .1, y = .1, amp = .1, custom_func = sintrend))
 
-continuous_exposure(n = 5, y = .6, mu = 100, sd = 10, trend = "custom", custom_func = sintrend)
-plot(continuous_exposure(n=1000, y = .6, mu = 5, sd = .5, trend = "custom", custom_func = sintrend))
+continuous_exposure(n = 5, y = .6, mu = 100, sd = 10, custom_func = sintrend)
+plot(continuous_exposure(n=1000, y = 1, mu = 200, sd = 50, custom_func = sintrend))
 ```
@@
 ```
+
+##Custom function for model
+
+```{r}
+trivial_mod <- function(df, constant){
+  newx <- constant*df$x
+  mod <- lm(outcome~newx, data=df)
+  out_1 <- summary(mod)$coef[2, ]
+  out_2 <- stats::confint.default(mod)[2, ]
+  out <- c(out_1, out_2)
+  return(out)
+}
+```

### 2017-05-15 R/sim_exposure.R

@@ -13,14 +13,13 @@
 #'      \item{"curvilinear"}
 #'      \item{"cos1linear"}
 #'      \item{"no trend"}
-#'      \item{"custom"}
 #'    }
 #'    See the package vignette for examples of the shapes of these trends.
 #'    Default is "no trend".
 #' @param slope A numeric value specifying the slope of the trend, to be used
 #'    with trend = "linear" or trend = "cos1linear".
 #' @param amp A numeric value specifying the amplitude of the seasonal trend.
-#'    Must be between 0 and 1.
+#'    Must be between -1 and 1.
 #' @param custom_func A character string specifying a customized function from
 #'    which to create a trend variable.
 #' @param ... optional arguments to a custom trend function
@@ -33,7 +32,12 @@
 #' @export
 calc_t <- function(n, trend = "no trend", slope=1, amp = .6, custom_func = NULL, ...){
   day <- c(1:n)
-  if (trend == "cos1"){
+  if (!is.null(custom_func)) {
+    arguments <- list(...)
+    arguments$n <- n
+    seasont <- do.call(custom_func, arguments)
+  }
+  else if (trend == "cos1"){
     seasont <- 1 + amp * cos(2 * pi * (day / 365))
     } else if (trend == "cos2"){
     seasont <- 1 + amp * cos(2 * pi * (day / 365)) +
@@ -48,10 +52,6 @@ calc_t <- function(n, trend = "no trend", slope=1, amp = .6, custom_func = NULL,
       seasont <- (1 + slope*(day / n)) * (1 + amp * cos(2 * pi * (day / 365)))
     } else if (trend == "no trend"){
       seasont <- rep(1, n)
-    } else if (trend == "custom" & !is.null(custom_func)) {
-      arguments <- list(...)
-      arguments$n <- n
-      seasont <- do.call(custom_func, arguments)
     } else {
       stop(paste0("`trend` value is not a valid choice. Please check the",
                   " function documentation to select a valid option."))
@@ -73,7 +73,6 @@ calc_t <- function(n, trend = "no trend", slope=1, amp = .6, custom_func = NULL,
 #'      \item{"linear"}
 #'      \item{"monthly"}
 #'      \item{"no trend"}
-#'      \item{"custom"}
 #'    }
 #'
 #' @param p A numeric value giving the mean probability of exposure
[…]

### 2017-05-16 R/assess_performance.R
Change to use custom model, change some spacing in code

@@ -14,7 +14,8 @@
 #'             exposure_type="continuous", exposure_trend = "cos1",
 #'             exposure_amp = .6, average_outcome = 22,
 #'             outcome_trend = "no trend", outcome_amp = .6, rr = 1.01)
-#' fits <- fit_mods(outcome = sims, model = "spline")
+#' fits <- fit_mods(data = sims, custom_model = spline_mod,
+#'                  custom_model_args = list(df_year = 1))
 #' mean_beta(df=fits)
 #'
 #' @export
@@ -47,7 +48,8 @@ mean_beta <- function(df){
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.01)
-#' fits <- fit_mods(outcome = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, custom_model = spline_mod,
+#'                  custom_model_args  = list(df_year = 1))
 #' beta_var(fits)
 #'
 #' @export
@@ -76,7 +78,8 @@ beta_var <- function(df){
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.01)
-#' fits <- fit_mods(data = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, custom_model = spline_mod,
+#'                  custom_model_args = list(df_year = 1))
 #' beta_bias(fits, true_rr = 1.02)
 #'
 #' @export
@@ -105,7 +108,8 @@ beta_bias <- function(df, true_rr){
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.01)
-#' fits <- fit_mods(data = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, custom_model = spline_mod,
+#'                  custom_model_args = list(df_year = 1))
 #' coverage_beta(df=fits, true_rr = 1.02)
 #'
 #' @export
@@ -132,7 +136,8 @@ coverage_beta <- function(df, true_rr){
 #'                     average_outcome = 20,
 #'                     outcome_trend = "no trend",
 #'                     rr = 1.01)
-#' fits <- fit_mods(data = sims, model = "spline", df_year = 1)
+#' fits <- fit_mods(data = sims, custom_model = spline_mod,
+#'                  custom_model_args = list(df_year = 1))
 #' power_beta(fits)
 #'
 #' @export
[…]

### 2017-05-16 R/customizable_functions.R

@@ -165,20 +165,20 @@ custom_baseline <- function(n, df = dlnm::chicagoNMMAPS, outcome_type = "cvd",
 #' @param amp A numeric value specifying the amplitude of the seasonal trend.
 #'    Must be between 0 and 1.
 #' @param ... optional arguments to a custom baseline function
-#' @param average_outcome A numeric value specifying the average outcome value
+#' @param average_baseline A numeric value specifying the average outcome value
 #' @param cust_base_func A character string specifying a user-made custom
 #'    function for baseline trend
 #'
 #' @return A numeric vector of baseline outcome values
 #'
 #' @examples
-#' create_baseline(n = 5, average_outcome = 22, trend = "linear")
-#' create_baseline(n = 5, average_outcome = NULL, trend = NULL,
-#'                 custom_func = "custom_baseline", outcome_type = "death")
+#' create_baseline(n = 5, average_baseline = 22, trend = "linear")
+#' # create_baseline(n = 5, custom_func = "custom_baseline", outcome_type = "death")
 #'
 #' @export
 #'
-create_baseline <- function(n, average_baseline, trend, slope, amp,
+create_baseline <- function(n, average_baseline = NULL, trend = "no trend",
+                            slope = 1, amp = 0.6,
                             cust_base_func = NULL, ...){
   if(is.null(cust_base_func)){
     lambda <- average_baseline
@@ -213,13 +213,12 @@ create_baseline <- function(n, average_baseline, trend, slope, amp,
 #'
 #' @examples
 #' base <- create_baseline(n = 10, average_baseline = 22, trend = "linear",
-#'                         slope=.4)
+#'                         slope = .4)
 #' exp <- sim_exposure(n = 5, central = 100, sd = 10, amp = .6,
 #'                     exposure_type = "continuous")
-#' create_lambda(baseline = base, exposure = exp, rr = 1.01)
+#' create_lambda(baseline = base, exposure = exp$x, rr = 1.01)
 #'
 #' @export
-
 create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
   if(is.null(cust_lambda_func)){
     if("data.frame" %in% class(baseline)){ baseline <- baseline$baseline }
@@ -245,6 +244,7 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
 #'    simulating values
 #' @inheritParams create_baseline
 #' @inheritParams create_lambda
[…]

### 2017-05-16 R/fit_models.R
Take out an example that wasn't working in `casecross`

@@ -38,9 +38,6 @@ spline_mod <- function(df, df_year = 7){
 #' @return Summary of the estimated log relative risk for a 1-unit increase in
 #'    exposure
 #'
-#' @examples
-#' casecross_mod(df = out)
-#'
 #' @export
 casecross_mod <- function(df){
   df$stratum <- factor(format(df$date, "%Y.%m"))
@@ -79,9 +76,6 @@ casecross_mod <- function(df){
 #' @return Summary of the estimated log relative risk for a 1-unit increase in
 #'    exposure
 #'
-#' @examples
-#' crossyear_mod(df = out)
-#'
 #' @export
 crossyear_mod <- function(df){
   df$stratum <- factor(format(df$date, "%j"))
@@ -89,13 +83,11 @@ crossyear_mod <- function(df){
 
   if (sum(df$x == 0 | df$x == 1) == length(df$x)){
     event.check <- as.matrix(table(df$stratum, df$x))
-    informative.strata <- rownames(event.check)[apply(event.check,
-                                                      1, prod) > 0]
-    df <- subset(df, stratum %in% informative.strata)
+    informative.strata <- rownames(event.check)[apply(event.check, 1, prod) > 0]
+    df <- dplyr::filter(df, ~ stratum %in% informative.strata)
 
     if(length(informative.strata) > 1){
-      mod <- stats::glm(y ~ x + stratum,
-                 data = df,
+      mod <- stats::glm(y ~ x + stratum, data = df,
                  family = stats::quasipoisson(link = "log"))
     } else {
       mod <- stats::glm(y ~ x,
@@ -103,8 +95,7 @@ crossyear_mod <- function(df){
                  family = stats::quasipoisson(link = "log"))
     }
   } else {
-    mod <- stats::glm(y ~ x + stratum,
-               data = df,
+    mod <- stats::glm(y ~ x + stratum, data = df,
                family = stats::quasipoisson(link = "log"))
   }
 

### 2017-05-16 R/plots.R
Change to use second argument of `eesim` output for coverage plot examples

@@ -18,10 +18,8 @@
 #' @examples
 #' ex_sim <- eesim(n_reps = 100, n = 1000, central = 100, sd = 10,
 #'           exposure_type = "continuous", average_outcome = 20, rr = 1.02,
-#'           model = "spline", df_year = 1)
-#' coverage_plot(ex_sim[[1]], true_param = 1.02)
-#'
-#' @importFrom dplyr %>%
+#'           custom_model = spline_mod, custom_model_args = list(df_year = 1))
+#' coverage_plot(ex_sim[[2]], true_param = 1.02)
 #'
 #' @export
 #'

### 2017-05-16 writing/EpiDraft.Rmd

@@
 
-```{r, echo=FALSE}
-#I used the variable "Legionellosis, Current week" so hopefully that's right
-leg1 <- read.csv("../Legionnaires data/2014Leg.csv")[,c(2,3,24)]
-leg2 <- read.csv("../Legionnaires data/2015Leg.csv")[,c(2,3,24)]
-leg3 <- read.csv("../Legionnaires data/2016Leg.csv")[,c(2,3,24)]
-leg <- data.frame(year = c(leg1$MMWR.Year, leg2$MMWR.Year, leg3$MMWR.Year), week = c(leg1$MMWR.Week, leg2$MMWR.Week, leg3$MMWR.Week), cases = c(leg1$Legionellosis..Current.week, leg2$Legionellosis..Current.week, leg3$Legionellosis..Current.week))
-plot(leg$cases, xlab = "Week", ylab = "Number of Cases", main = "Legionnaire's Disease 2014-2016")
-```
-
 ```{r, echo=F, warning=F, message=F}
@@
 #Row 48 is the only row with different states. 569 and 574 have NA in every column.
+
 ```
@@
 parain <- readRDS("data/pa.rds")
-nyrain <- readRDS("data/pa.rds")
+nyrain <- readRDS("data/ny.rds")
 paquant <- quantile(parain$prcp, probs = c(.75, .95, .99), na.rm=T)
@@
 #Monthly rainfall above 75th, 95th, and 99th percentiles for Philly and NYC
-pamonthly <- group_by(parain, month, year) %>% summarise(quant75 =mean(quant75), quant95=mean(quant95), quant99=mean(quant99))
-nymonthly <- group_by(nyrain, month, year) %>% summarise(mean(quant75), mean(quant95), mean(quant99))
+pamonthly <- group_by(parain, year, month) %>% summarise(quant75 =mean(quant75), quant95=mean(quant95), quant99=mean(quant99))
+nymonthly <- group_by(nyrain, year, month) %>% summarise(mean(quant75), quant95 = mean(quant95), mean(quant99))
 #What pattern is this?
 ggplot(pamonthly, aes(x=1:136, y=quant95))+geom_line()+geom_vline(xintercept=seq(from=1, to=136, by=12), col="red") + ggtitle("Monthly rainfall for Philadelphia 2006-2017")+ ylab("Proportion of days over 95th percentile")
+
+#Check if proportion constant averaging over months:
+plot(pamonthly %>% group_by(month) %>% summarise(mean1 = mean(quant95, na.rm=T)))
+plot(nymonthly %>% group_by(month) %>% summarise(mean1 = mean(quant95, na.rm=T)))
+#These two plots display a similar pattern, so we will use the monthly probabilities for PA in our simulation.
+rainprobs <-pamonthly %>% group_by(month) %>% summarise(mean1 = mean(quant95, na.rm=T))
 ```
 
-```{r}
-#Power analysis using eesim
+### Simulated Data with Monthly Exposure Trend
 
-```
-
-We can pull the trend from this data:
-
-```{r}
-```
-
-### Simulated Data with Similar Trend
-
-Many epidemiologic data sets are of a sensitive nature and not publicly available.  We can use eesim to simulate data with a similar seasonal trend to fit and assess models.
+Many epidemiologic data sets are of a sensitive nature and not publicly available.
+We can use eesim to simulate data with a similar seasonal trend to fit and assess models.
+Because rainfall in Pennsylvania and New York exhibited similar montly patterns, we will use the average proportion of days each month from 2006-2017 which had rainfall in the 95th percentile.
+We will specify these proportions in the "monthly" exposure trend.
 
@@
 #(ma=14.633, nyc=4.69, pa=5.314)
-exposure <- continuous_exposure(n=1001, mu = 65, sd = 5, trend = "cos1", amp=.4)
-#Make this into weekly data instead of daily.
-x <- seq.int(from=1, to=1001, by=7)
-weeklyexp <- exposure[x,]
-plot(weeklyexp, main = "Plot of Exposure values")
-outcome <- sim_outcome(weeklyexp, average_outcome = 4.69, trend = "no trend")
-plot(outcome$outcome)
-
+simdat <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", average_outcome = 5.314453)
 ```

### 2017-05-16 writing/custom_examples.Rmd

@@
   day <- c(1:n)
-  seasont <- p + y * sin(2 * pi * (day / 365))
+  p + y * sin(2 * pi * (day / 365))
 }
@@
 plot(bin_t(1000, p=.3, trend="cos1"))
-plot(bin_t(1000, p=.3, custom_func=binsin, y=.1))
+plot(bin_t(n= 1000, p=.3, custom_func=binsin, y=.1))
 ```
@@
 ```
-
-##Custom function for model
-
-```{r}
-trivial_mod <- function(df, constant){
-  newx <- constant*df$x
-  mod <- lm(outcome~newx, data=df)
-  out_1 <- summary(mod)$coef[2, ]
-  out_2 <- stats::confint.default(mod)[2, ]
-  out <- c(out_1, out_2)
-  return(out)
-}
-```

### 2017-05-16 R/customizable_functions.R
Update documentation

@@ -494,7 +494,7 @@ fit_mods <- function(data, custom_model = NULL, custom_model_args = list()){
 #'
 #' @return Output with the correct column names for use as a custom model in eesim.
 #'
-#' @example
+#' @examples
 #' dat <- data.frame(x=rnorm(1000, 0, 1), outcome = rnorm(1000, 5, 1))
 #' lin_mod <- lm(outcome~x, data=dat)
 #' format_out(lin_mod)

### 2017-05-16 R/sim_exposure.R
Close #30

@@ -101,6 +101,7 @@ bin_t <- function(n, p, trend = "no trend", slope = 1, amp = .01,
     arguments$n <- n
     arguments$p <- p
     seasont <- do.call(custom_func, arguments)
+    return(seasont)
   }
   else if (trend == "monthly"){
   }

### 2017-05-18 writing/EpiDraft.Rmd
Fixes during meeting 5/16

@@
 We can use eesim to simulate data with a similar seasonal trend to fit and assess models.
-Because rainfall in Pennsylvania and New York exhibited similar montly patterns, we will use the average proportion of days each month from 2006-2017 which had rainfall in the 95th percentile.
+Because rainfall in Philadelphia and New York exhibited similar montly patterns, we will use the average proportion of days each month from 2006-2017 which had rainfall in the 95th percentile.
 We will specify these proportions in the "monthly" exposure trend.
+5.3 is the weekly mean Legionnaire outcome value in Pennsylvania, and we divide by 7 to get the daily mean.
 
@@
 #(ma=14.633, nyc=4.69, pa=5.314)
-simdat <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", average_outcome = 5.314453)
+simdat <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 
+0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", 
+average_outcome = 5.3/7, custom_model = spline_mod)
 ```

### 2017-05-18 R/assess_performance.R

@@ -237,7 +237,6 @@ power_calc <- function(varying, values, n_reps, custom_model, central, exposure_
                        cust_base_func = NULL, cust_lambda_func = NULL,
                        cust_base_args = NULL, cust_lambda_args = NULL,
                        custom_model_args = NULL, plot = FALSE){
-
   msg <- paste("This function may take a minute or two to run, especially with lots of",
                "replications (`n_reps`) or options for `values`.")
   msg <- paste(strwrap(msg), collapse = "\n")

### 2017-05-18 R/sim_exposure.R

@@ -21,7 +21,7 @@
 #' @param amp A numeric value specifying the amplitude of the seasonal trend.
 #'    Must be between -1 and 1.
 #' @param custom_func A character string specifying a customized function from
-#'    which to create a trend variable.
+#'    which to create a trend variable. Must accept an argument "n".
 #' @param ... optional arguments to a custom trend function
 #'
 #' @return A numeric vector used to generate data with seasonal trends.
@@ -82,6 +82,8 @@ calc_t <- function(n, trend = "no trend", slope=1, amp = .6, custom_func = NULL,
 #'    Must be between -.5 and .5.
 #' @param start.date A date of the format "yyyy-mm-dd" from which to begin
 #'    simulating values
+#' @param custom_func A character string specifying a customized function from
+#'    which to create a trend variable. Must accept arguments "n" and "p".
 #' @inheritParams calc_t
 #'
 #' @return A numeric vector used to generate binary exposure data with seasonal

### 2017-05-18 R/customizable_functions.R
Add cust_outdraw to sim_exposure

@@ -48,6 +48,13 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
 #'    exposure values.
 #' @param cust_exp_args A list of arguments used in the user-specified custom
 #'    function.
+#' @param cust_expdraw A character string specifying a user-created function
+#'    which determines the distribution of random noise off of the trend line.
+#'    This function must have inputs "n" and "prob" for a binary exposure
+#'    function and inputs "n" and "mean" for a continuous exposure function.
+#'    The custom function must output a vector of simulated exposure values.
+#' @param cust_expdraw_args A list of arguments other than "n" and "prob" or
+#' "mean" required by the cust_expdraw function.
 #' @inheritParams std_exposure
 #' @inheritParams continuous_exposure
 #' @inheritParams binary_exposure
@@ -71,7 +78,8 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
 sim_exposure <- function(n, central = NULL, sd = NULL, trend = "no trend",
                          slope = 1, amp = .6, exposure_type = NULL,
                          start.date = "2001-01-01", cust_exp_func = NULL,
-                         cust_exp_args = NULL){
+                         cust_exp_args = NULL, cust_expdraw = NULL,
+                         cust_expdraw_args = list()){
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = n)
   if(is.null(cust_exp_args)){
@@ -240,6 +248,12 @@ create_lambda <- function(baseline, exposure, rr, cust_lambda_func = NULL, ...){
 #'    user-specified custom lambda function
 #' @param cust_base_args A list of arguments and their values used in the
 #'    user-specified custom baseline function
+#' @param cust_outdraw A character string specifying a user-created function to
+#' randomize the outcome values off of the baseline for outcome values. This
+#' function must take inputs "n" and "lambda" and output a vector of outcome
+#' values.
+#' @param cust_outdraw_args A list of arguments besides "n" and "lamba" passed
+#' to the user-created custom outcome draw function.
 #' @param start.date A date of the format "yyyy-mm-dd" from which to begin
 #'    simulating values
 #' @inheritParams create_baseline
[…]

### 2017-05-18 R/customizable_functions.R
Add custom draw arguments to other functions

@@ -397,6 +397,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 #' @param cust_lambda_args A list of arguments and their values used in the
 #'    user-specified custom lambda function
 #' @inheritParams std_exposure
+#' @inheritParams sim_exposure
 #' @inheritParams sim_outcome
 #'
 #' @return A list object, in which each list element is one of the synthetic datasets
@@ -422,22 +423,28 @@ create_sims <- function(n_reps, n, central, sd=NULL, exposure_type, exposure_tre
                         outcome_trend, outcome_slope=1,
                         outcome_amp, rr, start.date = "2000-01-01",
                         cust_exp_func = NULL, cust_exp_args = NULL,
+                        cust_expdraw = NULL, cust_expdraw_args = NULL,
                         cust_base_func = NULL, cust_lambda_func = NULL,
-                        cust_base_args = NULL, cust_lambda_args = NULL){
+                        cust_base_args = NULL, cust_lambda_args = NULL,
+                        cust_outdraw=NULL, cust_outdraw_args=NULL){
 
   exposure <- lapply(rep(n, times = n_reps), sim_exposure, central = central,
                      sd = sd, exposure_type = exposure_type,
                      slope = exposure_slope, amp = exposure_amp,
                      trend = exposure_trend, start.date = start.date,
                      cust_exp_func = cust_exp_func,
-                     cust_exp_args = cust_exp_args)
+                     cust_exp_args = cust_exp_args,
+                     cust_expdraw = cust_expdraw,
+                     cust_expdraw_args = cust_expdraw_args)
   outcome <- lapply(exposure, sim_outcome, average_outcome = average_outcome,
                     trend = outcome_trend, slope = outcome_slope,
                     amp = outcome_amp, rr = rr,
                     start.date = start.date, cust_base_func = cust_base_func,
                     cust_lambda_func = cust_lambda_func,
                     cust_base_args = cust_base_args,
-                    cust_lambda_args = cust_lambda_args)
+                    cust_lambda_args = cust_lambda_args,
+                    cust_outdraw=cust_outdraw,
+                    cust_outdraw_args=cust_outdraw_args)
   return(outcome)
 }
 
@@ -596,8 +603,10 @@ eesim <- function(n_reps, n, rr, exposure_type, custom_model,
                   outcome_trend = "no trend", outcome_slope = NULL,
                   outcome_amp = NULL, start.date = "2000-01-01",
[…]

### 2017-05-31 R/customizable_functions.R
Just add some spacing in some code

@@ -277,7 +277,7 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
 
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = nrow(exposure))
-  average_baseline <- average_outcome/exp(log(rr)*mean(exposure$x))
+  average_baseline <- average_outcome / exp(log(rr) * mean(exposure$x))
   if(is.null(cust_base_func) & is.null(cust_lambda_func)){
     if(is.null(average_outcome)){
       stop(paste0("If custom functions are not used to generate outcomes,

### 2017-05-31 R/customizable_functions.R
Make sure average outcome is the mean value we want

@@ -274,7 +274,6 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
                         cust_base_func = NULL, cust_lambda_func = NULL,
                         cust_base_args = list(), cust_lambda_args = list(),
                         cust_outdraw = NULL, cust_outdraw_args = list()){
-
   start.date <- as.Date(start.date)
   date <- seq(from = start.date, by = 1, length.out = nrow(exposure))
   average_baseline <- average_outcome / exp(log(rr) * mean(exposure$x))
@@ -311,10 +310,12 @@ sim_outcome <- function(exposure, average_outcome = NULL, trend = "no trend",
     cust_lambda_args$exposure <- exposure$x
     lambda <- do.call(cust_lambda_func, cust_lambda_args)
   }
+
+  lambda <- lambda * average_outcome / mean(lambda) # Make sure we have desired mean
+
   if (is.null(cust_outdraw)){
   outcome <- stats::rpois(n = nrow(exposure), lambda = lambda)
-  }
-  else {
+  } else {
     cust_outdraw_args$lambda <- lambda
     cust_outdraw_args$n <- nrow(exposure)
     outcome <- do.call(cust_outdraw, cust_outdraw_args)

### 2017-06-01 R/assess_performance.R
Use `spline_mod` instead of case-crossover for `power_calc` example

@@ -224,7 +224,7 @@ check_sims <- function(df, true_rr){
 #'            central = 100, sd = 10, rr = 1.001, exposure_type = "continuous",
 #'            exposure_trend = "cos1", exposure_amp = .6, average_outcome = 22,
 #'            outcome_trend = "no trend", outcome_amp = .6,
-#'            custom_model = casecross_mod, plot = TRUE)
+#'            custom_model = spline_mod, plot = TRUE)
 #'}
 #'
 #' @export

### 2017-06-02 R/assess_performance.R
Make edits to helpfiles

@@ -1,13 +1,14 @@
 #' Average Estimated Coefficient
 #'
-#' This function gives the mean value of the \eqn{\hat{\beta}}s and the mean
-#' estimated relative risk over the n simulations.
+#' This function gives the mean value of the estimated log relative risks (\eqn{\hat{\beta}}s)
+#' and the mean of the estimated relative risk values over the \code{n} simulations.
 #'
 #' @param df A data frame of replicated simulations which must include a column
-#'    titled "Estimate"
+#'    titled "Estimate" with the effect estimate from the fitted model.
 #'
-#' @return A data frame with the mean estimated coefficient and mean estimated
-#'    relative risk
+#' @return A data frame with the mean estimated log relative risk and mean estimated
+#'    relative risk. The mean estimated risk is based on first calculating the
+#'    mean log relative risk and then exponentiating this mean value.
 #'
 #' @examples
 #' sims <- create_sims(n_reps=10, n=50, central = 100, sd = 10,
@@ -24,15 +25,16 @@ mean_beta <- function(df){
     stop("Input must be a data frame")
   }
   beta_hat <- mean(df$Estimate)
-  rr_hat <- mean(exp(df$Estimate))
+  rr_hat <- mean(beta_hat)
   out <- data.frame(beta_hat, rr_hat)
   return(out)
 }
 
 #' Standard Deviation of Estimated Coefficients
 #'
-#' This function gives the variance of the point estimates of beta hat over
-#' the n simulations and the mean of the variances of each beta hat.
+#' Measures the variance of the point estimates of the estimated log relative risk
+#' (\eqn{\hat{beta}}{b}) over the \code{n_rep} simulations and the mean of
+#' the variances of each \eqn{\hat{\beta}}{b}.
 #'
 #' @param df A data frame of replicated simulations which must include columns
 #'    titled "Estimate" and "Std.Error".
@@ -64,11 +66,18 @@ beta_var <- function(df){
 #'
 #' This function returns the relative bias of the mean of the estimated coefficients.
 #'
-#' @inheritParams mean_beta
-#' @param true_rr The true relative risk used to simulate your data
 #'
-#' @return The percent bias of the mean of the estimated coefficients over n
-#'    simulations
+#' @details This function estimates the percent bias in the estimated log relative risk
+#'    (\eqn{b}) as:
+#' \deqn{100 * \frac{\beta - \hat{\beta}}{\beta}}{100 ((\beta - b) / \beta)}
[…]

### 2017-06-02 R/customizable_functions.R

@@ -1,5 +1,7 @@
 #' Pull exposure series from data set
 #'
+#' Example of a custom exposure function that can be passed to \code{\link{eesim}} or
+#' \code{\link{power_calc}}.
 #' By default, this function pulls exposure data from the Chicago NMMAPS data
 #' set in the dlnm package.  The user may specify a different data set from
 #' which to pull exposure values.
@@ -15,11 +17,12 @@
 #'      \item"rhum"
 #'      \item"pm10"
 #'      \item"o3"}
+#' (Note: These are the column names for exposure measurements in the observed data.)
 #' @param start.date A date of the format "yyyy-mm-dd" from which to begin
 #'    pulling exposure values. Dates in the Chicago NMMAPS data set are from
 #'    1987-01-01 to 2000-12-31.
 #'
-#' @return A numeric vector of exposure values
+#' @return A numeric vector of length \code{n} giving exposure values.
 #'
 #' @examples
 #' custom_exposure(n = 5, metric = "temp", start.date = "2000-01-01")
@@ -44,17 +47,17 @@ custom_exposure <- function(n, df = dlnm::chicagoNMMAPS, metric = "temp",
 #' without seasonal trends. It also allows the user to use a custom function for
 #' the shape of the exposure trend.
 #'
-#' @param cust_exp_func The name of a function to use to generate custom
+#' @param cust_exp_func An R object specifying the function to use to generate custom
 #'    exposure values.
 #' @param cust_exp_args A list of arguments used in the user-specified custom
 #'    function.
-#' @param cust_expdraw A character string specifying a user-created function
+#' @param cust_expdraw An R object specifying a user-created function
 #'    which determines the distribution of random noise off of the trend line.
-#'    This function must have inputs "n" and "prob" for a binary exposure
-#'    function and inputs "n" and "mean" for a continuous exposure function.
+#'    This function must have inputs \code{n} and \code{prob} for a binary exposure
+#'    function and inputs \code{n} and \code{mean} for a continuous exposure function.
 #'    The custom function must output a vector of simulated exposure values.
-#' @param cust_expdraw_args A list of arguments other than "n" and "prob" or
-#' "mean" required by the cust_expdraw function.
+#' @param cust_expdraw_args A list of arguments other than \code{n} required by the
+#'    \code{cust_expdraw} function.
 #' @inheritParams std_exposure
 #' @inheritParams continuous_exposure
 #' @inheritParams binary_exposure
[…]

### 2017-06-02 R/plots.R

@@ -1,16 +1,18 @@
-#' Coverage Plot
+#' Plot coverage of empirical confidence intervals
 #'
-#' This plot displays the relative risk point estimates and their confidence
-#' intervals for each repitition of the simulation in relation to the true
-#' relative risk.  It is a visualization of the coverage of the specified
+#' Plots the relative risk point estimates and their confidence
+#' intervals for model fit results for each simulation, compared to the true
+#' relative risk. This gives a visualization of the coverage of the specified
 #' method for the relative risk. The confidence intervals which do not contain
-#' the true relative risk appear in red.
+#' the true relative risk appear in red. The input to this function should be
+#' either the output of \code{\link{fit_mods}} or the second element of the
+#' output of \code{\link{eesim}}.
 #'
 #' @param summarystats A list or data frame of summary statistics from many
-#' repitions of a simulation.  Must include columns titled "Estimate",
-#' "lower_ci", and "upper_ci". This could be the first object from the output of
-#' eesim, specified by using the format "eesim_output[[1]]".
-#' @param true_param The true value of the parameter being estimated
+#' repitions of a simulation.  Must include columns titled \code{Estimate},
+#' \code{lower_ci}, and \code{upper_ci}. This could be the second object from the output of
+#' \code{\link{eesim}}, specified by using the format \code{eesim_output[[2]]}.
+#' @param true_param The true value of the relative risk used to simulate the data.
 #'
 #' @return A plot displaying the coverage for the true value of the parameter by
 #' the confidence intervals resulting from each repitition of the simulation.
@@ -43,19 +45,24 @@ coverage_plot <- function(summarystats, true_param){
     ggplot2::scale_x_discrete(name = "", breaks = NULL)
   return(out)
 }
+
+#' Create calendar plot
 #'
-#' Calendar Plot
+#' Creates a calendar plot of a time series of continuous or discrete data. The time series
+#' data frame input to this function must have only two columns, one for the date and one with
+#' the values to plot.
 #'
-#' This plot displays daily values of a binary or continuous variable on a
-#' calendar via color.
-#'
-#' @param df Data frame with one column for date with entries in the format
-#' "yyyy-mm-dd" and one column for the daily values of the variable.
+#' @param df Data frame with one column named \code{date} for date with entries in the format
+#' "yyyy-mm-dd" and one column for the daily values of the variable to plot.
 #' @param type Character string specifying whether the exposure is continuous or
 #' discrete
[…]

### 2017-06-05 writing/EpiDraft.Rmd
Add plot aggregating Leg data monthly

@@
 
+leg_cases$mo <- month(leg_cases$week_end_date)
+leg_cases$yr <- year(leg_cases$week_end_date)
+leg_cases$moyr <- ymd(strtrim(leg_cases$week_end_date, width=7), truncated=3)
+
+#Plot of monthly cases for Pennsylvania
+leg_cases[,c(1,4,8,9, 10)] %>% group_by(moyr) %>% summarise(sum=sum(penn, na.rm=T)) %>% ggplot(aes(x=moyr, y=sum))+ geom_bar(stat="identity") + labs(x="Date", y="Cases", title = "Monthly Legionnaire's Cases for Pennsylvania")
+
 ```
@@
 #(ma=14.633, nyc=4.69, pa=5.314)
-simdat <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 
-0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", 
-average_outcome = 5.3/7, custom_model = spline_mod)
+simdat <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", outcome_trend="cos1", average_outcome = 5.3/7, custom_model = spline_mod)
+
+#Create new month/year variable to be able to aggregate by months
+simdat$simulated_datasets[[1]]$moyr <- ymd(strtrim(simdat$simulated_datasets[[1]]$date, width=7), truncated=3)
+
+#Plot of monthly cases 
+simdat$simulated_datasets[[1]] %>% group_by(moyr) %>% summarise(sum=sum(outcome, na.rm=T)) %>% ggplot(aes(x=moyr, y=sum))+ geom_bar(stat="identity") + labs(x="Date", y="Cases", title = "Simulated Monthly Legionnaire's Cases")
 ```

### 2017-06-05 README.Rmd
Add Travis CI checking

@@
 output: github_document
+bibliography: vignettes/eesim_bibfile.bib
 ---
@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
+
+[![Travis-CI Build Status](https://travis-ci.org/sakoehler7/eesim.svg?branch=master)](https://travis-ci.org/sakoehler7/eesim)
 
@@
 library(tidyr)
-library(viridisLite)
 library(gridExtra)
@@
 
-1. Generation of exposure data; 
-2. Generation of outcome data;
-3. Fitting models to simulated data; and 
-4. Evaluating model performance on simulated data.
+1. Generation of exposure data (simulated or from real data); 
+2. Generation of simulated outcome data;
+3. Fitting models to generated data; and 
+4. Evaluating model performance on generated data.
 
@@
 
-The package creates time series that are relevant for environmental epidemiology studies of ambient exposures (e.g., studies of acute mortality risks associated with daily air pollution concentration, daily temperature, or occurance of a community-wide extreme event like a heat wave).
+The package creates simulated time series data that are relevant for environmental epidemiology studies of ambient exposures (e.g., studies of acute mortality risks associated with daily air pollution concentration, daily temperature, or occurance of a community-wide extreme event like a heat wave).
+Simulated environmental datasets like those created by the package can be used in to assess the performance of statistical models meant to estimate the association between exposure level and outcome risk, to estimate power for a planned study, and to develop a better understanding of the data generating processes behind observed environmental datasets.
+Such time series are often characterized by both seasonal and long-term trends in both the exposure of interest and the outcome.
+For example, the following plot shows time series of daily ozone concentration (in parts per billion [ppb]) and cardiovascular deaths in Chicago, IL (1996--2000), with smoothed lines overlaid on the raw data to show patterns over time.
+
+```{r message = FALSE, echo = FALSE, fig.width = 5, fig.align = "center"}
+library(dlnm)
+data("chicagoNMMAPS")
+chicagoNMMAPS %>% 
+  tbl_df() %>% 
+  filter(year >= 1996) %>% 
+  select(date, cvd, o3) %>% 
+  rename(`Cardiovascular deaths` = cvd,
+         `Ozone concentration (ppb)` = o3) %>% 
+  gather(variable, value, -date) %>% 
+  ggplot(aes(x = date, y = value)) + 
+  geom_point(alpha = 0.5, size = 1) + 
+  geom_smooth(se = FALSE, span = 0.1, method = "loess", color = "red") +
+  facet_wrap(~ variable, ncol = 1, scales = "free_y") + 
+  theme_classic() + 
+  labs(x = "Date", y = "Variable value")
+```
 
@@
 The main function of this package is the `eesim` function.
-You can use the `eesim` function to conduct all four steps of the simulation process at once (simulate exposure data, simulate outcome data, fit models to simulated data, and evaluate model performance).
+You can use the `eesim` function to conduct all four steps of the simulation process at once (generate exposure data, generate outcome data, fit models to simulated data, and evaluate model performance).
 
-The `eesim` function requires [required inputs].
-The function returns a list with two elements.
-The first element gives simulation-specific results: the estimated effect, standard error, t- and p-values, and upper and lower 95% confidence bounds when a model was applied to each set of simulated data.
-The second element gives some measures of model assessment, assessed over all simulations, including the mean beta and relative risk estimates across simulations.
+The `eesim` function requires inputs on:
 
-For example, you can use the following call to (1) generate 10 observations of a continuous exposure with mean 100 and standard deviation 10, with a seasonal trend in expected value; (2) generate 10 associated outcome values, where the outcome has an average value of 20 and a relative risk of 1.10 per one-unit increase in the exposure; (3) fit a generalized linear model that controls for long-term and seasonal trends with a natural cubic spline with 1 degree of freedom per year; and (4) evaluate the performace of that model on the simulated data:
+- `n`: The desired number of observations per simulated dataset (for a daily time series, this is the desired number of days in the simulated dataset)
+- `n_reps`: The desired number of simulated datasets
+- `exposure_type`: Whether the exposure is binary (e.g., occurence of an extreme event like a heat wave or wildfire) or continuous (e.g., concentration of a pollutant)
+- `rr`: The relative rate of the outcome associated with the exposure.
+For a binary exposure, this is the relative rate associated with the exposure compared to a similar day without the exposure.
+For a continuous exposure, this is the relative rate associated with a one-unit increase in the exposure.
+- `model`: The model to be used to estimate the association between exposure and outcome in the simulated datasets, either to estimate power of a planned analysis or to otherwise evaluate the performance (e.g., coverage, bias) of a model on the simulated datasets.
 
-[To do-- increase `n` and `n_reps`, but wait until we've edited the vignette, because that will increase time to render.
-We could change `df` then, too (and maybe change to case-crossover?
-That may be easier for a first example)]
+A number of optional inputs can also be specified, including arguments to adjust the shape of seasonal or long-term trends in the exposure or outcome data or custom arguments to use at different steps of the data generation.
+
+The function returns a list with three elements.
+The first element is a list with all the simulated datasets.
+The second element gives simulation-specific results for each simulated dataset: the estimated effect, standard error, t- and p-values, and upper and lower 95% confidence bounds when a model was applied to each of the simulated datasets.
+The third element gives some measures of model assessment, assessed over all simulations, including the mean beta and relative risk estimates across simulations.
+
+```{r message = FALSE, echo = FALSE, results='hide'}
+chicagoNMMAPS %>% 
+  tbl_df() %>% 
+  filter(year >= 1996) %>% 
+  select(date, cvd, o3) %>% 
+  gather(variable, value, -date) %>% 
+  group_by(variable) %>% 
+  summarize(mean = mean(value, na.rm = TRUE),
+            sd = sd(value, na.rm = TRUE))
+
+library(splines)
+a <- chicagoNMMAPS %>% 
+  tbl_df() %>% 
+  filter(year >= 1996) %>% select(date, o3)
+mod <- lm(o3 ~ ns(scale(time, scale = FALSE, center = TRUE), 7 * 14), data = chicagoNMMAPS)
+sd(residuals(mod))
+```
+
+For example, in the observed data from Chicago, IL, shown in the plots above, daily ozone concentrations have a mean of about 20 ppb and standard deviation of about 7 ppb after removing seasonal and long-term trends.
+The average number of cardiovascular deaths per day is around 50.
+Here is the code, and a plot of the resulting data, for generating a dataset with similar characteristics for use in a power analysis or to evaluate model performance (later in the vignette, we will show how to use customization to further improve the simulation of data for this example, including avoiding negative values of ozone concentration in simulated data):
 
 ```{r, warning = FALSE, message = FALSE}
-ex_sim <- eesim(n_reps = 100, n = 10, central = 100, sd = 10,
-                exposure_type = "continuous", 
-                average_outcome = 20, rr = 1.10,
-                model = "spline", df_year = 1)
-head(ex_sim[[1]])
-ex_sim[[2]]
+sim_chicago <- create_sims(n_reps = 1, n = 365 * 5, central = 20, sd = 7,
+                           exposure_type = "continuous", exposure_trend = "cos1",
+                           exposure_amp = -.6, average_outcome = 50,
+                           outcome_trend = "cos1", outcome_amp = 0.2, 
+                           rr = 1.0005, start.date = "1996-01-01")
+head(sim_chicago[[1]])
 ```
 
-The first element of the returned object can be used to explore the behavior of individual simulations.
-For example, to look at the relative risk point estimate and 95% confidence interval from each of the 100 simulations, as well as which 95% confidence intervals include the true relative risk using the `coverage_plot` function that comes with the package:
+```{r echo = FALSE, message = FALSE, echo = FALSE, fig.width = 5, fig.align = "center"}
+sim_chicago[[1]] %>% 
+  rename(Exposure = x,
+         Outcome = outcome) %>% 
+  gather(variable, value, -date) %>% 
+  mutate(variable = factor(variable, levels = c("Outcome", "Exposure"))) %>% 
+  ggplot(aes(x = date, y = value)) + 
+  geom_point(alpha = 0.5, size = 1) + 
+  facet_wrap(~ variable, ncol = 1, scales = "free_y") + 
+  theme_classic() + 
+  geom_smooth(se = FALSE, method = "loess", color = "red", span = 0.1)
+```
+
+This simulated data can also be visualized using the `calendar_plot` function that comes with the package:
+
+```{r fig.align = "center", fig.width = 8, fig.height = 7.5}
+a <- calendar_plot(sim_chicago[[1]] %>% select(date, outcome), type = "continuous", 
+                   legend_name = "Outcome") + 
+  ggtitle("Outcome")
+b <- calendar_plot(sim_chicago[[1]] %>% select(date, x), type = "continuous") + 
+  ggtitle("Exposure")
+grid.arrange(a, b, ncol = 1)
+```
+
+You can use the `eesim` function to generate multiple similar simulated datasets and investigate the performance of a specified model in estimating the association between ozone concentration and the risk of cardiovascular death in 20 simulated datasets.
+You must write a function with the code to fit the model you desire to fit to the simulated data (more details for writing this function are provided later in the vignette), and then you can use the `eesim` function to generate lots of simulated datasets, fit that model, and assess its performance using a call like:
+
+```{r echo = FALSE}
+set.seed(16)
+```
+
+```{r, warning = FALSE, message = FALSE}
+ex_sim <- eesim(n_reps = 100, n = 365 * 5, central = 20, sd = 7,
+                exposure_type = "continuous", exposure_trend = "cos1",
+                exposure_amp = -.6, average_outcome = 50,
+                outcome_trend = "cos1", outcome_amp = 0.2, 
+                rr = 1.2, start.date = "1996-01-01",
+                custom_model = spline_mod, custom_model_args = list(df_year = 7))
+```
+
+The `eesim` function returns a list with three elements:
+
+```{r}
+names(ex_sim)
+```
+
+The first element of the returned object is a list with all of the simulated datasets. For example, you can create a calendar plot of exposure in the first simulated dataset using the call:
+
+```{r fig.height = 3, fig.width = 7, fig.align = "center"}
+calendar_plot(ex_sim[["simulated_datasets"]][[1]] %>% select(date, x), type = "continuous")
+```
+
+The second element of the returned object gives the results of fitting the model to each of the simulated datasets. It can be used to explore the behavior of individual simulations:
+
+```{r}
+head(ex_sim[["indiv_performance"]])
+```
+
+After running the simulation, you can look at the relative risk point estimate and 95% confidence interval from each of the 100 simulations, as well as which 95% confidence intervals include the true relative rate, using the `coverage_plot` function that comes with the package:
 
 ```{r fig.width = 4, fig.height = 5, fig.align = "center"}
-coverage_plot(ex_sim[[1]], true_param = 1.10)
+coverage_plot(ex_sim[["indiv_performance"]], true_param = 1.2)
 ```
 
-The second element of the returned object gives the following overall summaries of model performance across all simulations:
+The third element of the list returned by a call to `eesim` gives the following overall summaries of model performance across all simulations:
 
@@
 eesim_overall_output <- data_frame(element = paste0("`",
-                                                    colnames(ex_sim[[2]]),
+                                                    colnames(ex_sim[[3]]),
                                                     "`"),
-                                   desc = c("The mean estimated effect (log relative risk) over all simulations",
-                                            "The mean relative risk over all simulations",
-                                            "Variance of the point estimates of the estimated effect over all simulations",
-                                            "The mean across simulations of the variances of each estimated effect",
-                                            "Relative bias of the mean of the estimated coefficients",
-                                            "Percent of simulations for which the 95% confidence interval estimate includes the true value",
-                                            NA))
+                                   desc = c("**Mean estimate**: The mean of the estimated log relative rate over all simulations.",
+                                            "**Mean estimated relative rate**: The mean of the estimated relative rate over all simulations.",
+                                            "**Variance across estimates**: Variance of the point estimates (estimated log relative risk) over all simulations.",
+                                            "**Mean variance of estimate**: The mean of the variances of the estimated effect (estimated log relative risk) across all simulations.",
+                                            "**Relative bias**: Difference between the estimated log relative risk and true log relative risk as a proportion of the true log relative risk.",
+                                            "**95% confidence inverval coverage**: Percent of simulations for which the 95% confidence interval estimate of log relative risk includes the true value of log relative risk.",
+                                            "**Power**: Percent of simulations for which the null hypothesis that the log relative risk equals zero is rejected based on a p-value of 0.05."))
 knitr::kable(eesim_overall_output, col.names = c("Variable", "Description"))
@@
 
-##Piece-by-piece breakdown of package utility
+For example, here are the overall results for the simulation fit above:
+
+```{r}
+ex_sim[["overall_performance"]]
+```
+
+In later sections of this vignette, we will show how to customize steps in the generation of the simulated data to further improve this example simulation.
+
+As another basic example, here is a plot of the dates of extreme heat days (defined as a day with temperature at or above the 98\textsuperscript{th} percentile temperature in Chicago between 1987 and 2000) in the observed Chicago dataset (points are jittered along the y-axis to limit overlapping):
+
+```{r echo = FALSE, message = FALSE, echo = FALSE, fig.width = 5, fig.align = "center", fig.height = 2}
+chicagoNMMAPS %>% 
+  mutate(temp = temp >= quantile(temp, probs = 0.98)) %>% 
+  tbl_df() %>% 
+  filter(year >= 1996) %>% 
+  select(date, temp) %>% 
+  rename(`Extreme heat day` = temp) %>% 
+  gather(variable, value, -date) %>% 
+  ggplot(aes(x = date, y = value)) + 
+  geom_jitter(alpha = 0.5, size = 0.7, fill = NA, width = 0, height = 0.1) + 
+  facet_wrap(~ variable, ncol = 1, scales = "free_y") + 
+  theme_classic() + 
+  labs(x = "Date", y = "Variable value")
+```
+
+```{r echo = FALSE, results="hide"}
+chicagoNMMAPS %>% 
+  mutate(temp = temp >= quantile(temp, probs = 0.98)) %>% 
+  tbl_df() %>% 
+  filter(year >= 1996) %>% 
+  group_by(month) %>% 
+  summarize(p_heat = mean(temp))
+```
+
+In this observed data, there is (unsurprisingly) a strong seasonal trend in this binary exposure of extreme heat days.
+The percent of days that are extreme heat days is 0% for all months expect June (about 5% of days in observed data were extreme heat days), July (about 12% of days), and August (about 2% of days).
+Similar exposure time series can be simulated with the call:
+
+```{r}
+sim_chicago2 <- create_sims(n_reps = 1, n = 365 * 5, sd = 1,
+                            central = c(0, 0, 0, 0, 0, 0.05, 0.12, 0.02, 0, 0, 0, 0),
+                            exposure_type = "binary", exposure_trend = "monthly",
+                            exposure_amp = -.6, average_outcome = 50,
+                            outcome_trend = "cos1", outcome_amp = 0.2, 
+                            rr = 1.05, start.date = "1996-01-01")
+```
+
+Here is an example of the simulated exposure data:
+
+```{r echo = FALSE, message = FALSE, echo = FALSE, fig.width = 5, fig.align = "center", fig.height = 2}
+sim_chicago2[[1]] %>% 
+  select(date, x) %>% 
+  rename(`Extreme heat day` = x) %>% 
+  gather(variable, value, -date) %>% 
+  ggplot(aes(x = date, y = value)) + 
+  geom_jitter(alpha = 0.5, size = 0.7, fill = NA, width = 0, height = 0.1) + 
+  facet_wrap(~ variable, ncol = 1, scales = "free_y") + 
+  theme_classic() + 
+  labs(x = "Date", y = "Variable value")
+```
+
+Again, both the observed and simulated exposure data can also be plotted using the `calendar_plot` function:
+
+```{r fig.align = "center", fig.width = 8, fig.height = 7}
+a <- chicagoNMMAPS %>% 
+  mutate(temp = temp >= quantile(temp, probs = 0.98)) %>% 
+  tbl_df() %>% 
+  filter(year >= 1996) %>% 
+  select(date, temp) %>% 
+  calendar_plot(type = "discrete", labels = c("Extreme heat day", "Other day")) + 
+  ggtitle("Observed exposure data")
+b <- sim_chicago2[[1]] %>% 
+  select(date, x) %>% 
+  calendar_plot(type = "discrete", labels = c("Extreme heat day", "Other day")) + 
+  ggtitle("Simulated exposure data")
+grid.arrange(a, b, ncol = 1)
+```
+
+The comparison of the observed and simulated data in this case suggests some clustering in the observed data that is not evident in the simulated data, suggesting that the probability of exposure may be higher on a day near other extreme heat days.
+
+The `eesim` function can be used to assess the performance of a GLM in estimating relative risk of cardiovascular mortality for extreme heat days compared to other days using:
+
+```{r echo = FALSE}
+set.seed(2416)
+```
+
+```{r}
+ex_sim2 <- eesim(n_reps = 100, n = 365 * 5, 
+                 central = c(0, 0, 0, 0, 0, 0.05, 0.12, 0.02, 0, 0, 0, 0),
+                 exposure_type = "binary", exposure_trend = "monthly",
+                 exposure_amp = -.6, average_outcome = 50,
+                 outcome_trend = "cos1", outcome_amp = 0.2, 
+                 rr = 1.05, start.date = "1996-01-01",
+                 custom_model = spline_mod, custom_model_args = list(df_year = 7))
+```
+
+As before, a plot of CI coverage can be created with `coverage_plot`:
+
+```{r fig.width = 4, fig.height = 5, fig.align = "center"}
+coverage_plot(ex_sim2[["indiv_performance"]], true_param = 1.05)
+```
+
+Here are the overall estimates in this case for model performance:
+
+```{r}
+ex_sim2[["overall_performance"]]
+```
+
+In this case, the expected power is low.
+
+The `power_calc` function in the package allows you to extend on this simulation functionality to create power curves for an analysis given an anticipated underlying process of data generation.
+This function will create simulations for several different values of number of days in the study (`n`), average daily outcome counts (`average_outcome`), or expected association between exposure and outcome (`rr`).
+
+For example, the following call generates a power curve that explores how expected power changes with an increasing number of days for the heat wave analysis example just presented (as a warning, this call takes a few minutes to run, since it's simulating many datasets):
+
+```{r fig.width = 5, fig.height = 3.5, fig.align = "center"}
+ex_power_calc <- power_calc(varying = "n", values = floor(365.25 * seq(1, 20, by = 5)),
+                            n_reps = 100, rr = 1.05,
+                            central = c(0, 0, 0, 0, 0, 0.05, 0.12, 0.02, 0, 0, 0, 0),
+                            exposure_type = "binary", exposure_trend = "monthly", 
+                            exposure_amp = -.6, average_outcome = 50,
+                            outcome_trend = "cos1", outcome_amp = 0.2, 
+                            custom_model = spline_mod, custom_model_args = list(df_year = 7),
+                            plot = FALSE)
+ex_power_calc %>% 
+  ggplot(aes(x = values, y = power)) + 
+  geom_line() + 
+  ylim(0, 1) + 
+  labs(x = "Number of days in the study", y = "Power") + 
+  theme_bw()
+```
+
+## Piece-by-piece breakdown of package utility
 
@@
 The first task of the package is generating exposure data.
-This can be dones with the `sim_exposure` function.
-In this function, the user can specify whether they would like the exposure data to be binary or continuous (`exposure_type`).
-For continuous exposure data, the user can specify the mean (`central`) and standard deviation (`sd`) of the exposure data:
+This can be done with the `sim_exposure` function.
+In this function, the user can specify whether he or she would like to generate exposure data that is binary or continuous (`exposure_type`).
+For continuous exposure data, the user must specify the mean (`central`) and standard deviation (`sd`) of the exposure data.
+For example, the following call simulates a dataframe of exposure data for an exposure that is normally distributed, with a mean value of 50, a standard deviation of 5, and no long-term or seasonal trends:
 
 ```{r}
-x_cont <- sim_exposure(n = 1000, central = 50, sd = 5,
-                      exposure_type = "continuous") 
+x_cont <- sim_exposure(n = 1000, central = 50, sd = 5, exposure_type = "continuous") 
 x_cont %>% slice(1:5)
@@
 ```{r fig.width = 4, fig.height = 2, fig.align = "center"}
-ggplot(x_cont, aes(x = date, y = x)) + geom_point(alpha = 0.2)
+ggplot(x_cont, aes(x = date, y = x)) + geom_point(alpha = 0.2) + 
+  theme_classic()
 ```
 
-You can plot a calendar plot of this simulated exposure time series using the `calendar_plot` function that comes with the package:
+You can plot a calendar plot of this simulated exposure time series using the `calendar_plot` function that comes with the package.
+Within this function, the type of data ("continuous" or "discrete") must be specified:
 
@@
 
-For binary exposure data, the `central` argument gives the probability of exposure on a study day:
+You can similarly use the `sim_exposure` function to simulate a binary exposure (e.g., occurence of an extreme event).
+For binary exposure data, the `central` argument of `sim_exposure` must also be expressed, but in this case it gives the probability of exposure on a study day:
 
 ```{r}
-x_bin <- sim_exposure(n = 1000, central = 0.1,
-                      exposure_type = "binary")
+x_bin <- sim_exposure(n = 1000, central = 0.05, exposure_type = "binary")
 x_bin %>% slice(1:5)
@@
 
-```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
+Again, the `calendar_plot` function can be used to visualize the generated time series.
+In the case of binary exposure data, the labels to be used in the legend for each outcome level must also be specified using the `labels` argument:
+
+```{r fig.align = "center", fig.width = 8, fig.height = 2.25}
 calendar_plot(x_bin, type = "discrete", labels = c("Not exposed", "Exposed"))
@@
 
-For environmental epidemiology applications, exposure data often has a seasonal trend and / or long-term trend.
+So far, these `sim_exposure` calls have been used to simulate basic exposure data, without long-term or seasonal trends.
+However, for environmental epidemiology applications, exposure data often has a seasonal trend and / or long-term trend, and these temporal trends can serve as confounders in assessing the association between time-varying environmental exposures and health outcomes.
+The `sim_exposure` function therefore includes options to generate exposure data with long-term and seasonal trends relevant to environmental time series studies, through the `trend` argument.
+
 The default for `sim_exposure` is to simulate the exposure data without a time trend (`trend = "no trend"`).
-However, we have also built in several time trends from which a user can to choose to simulate exposure data with a time trend, either seasonal or long-term or both.
+However, we have also built in several time trends from which a user can to choose to simulate exposure data with a time trend, either seasonal or long-term or both, based on trend patterns used in a simulation study of case-crossover studies as a method of controlling for seasonal and long-term trends in environmental epidemiology studies [@Bateson1999].
 These trend patterns differ slightly depending on whether the user is simulating binary or continuous data.
-[add reference for Schwartz paper where we got some of the trend equations] Below are plots of the built-in trends for continuous exposure data from which the user may choose.
-
+Below are plots of the built-in trends for continuous exposure data from which the user may choose.
 ```{r fig.width = 7, fig.height = 4, echo = FALSE, message = FALSE}
@@
 
-The trend options are similar for binary exposure, but exclude  "curvilinear" and "cos1linear" and include a "monthly" trend, for which the user enters a vector of 12 probabilities of exposure, one for each month.
+You can use the `amp` argument to adjust the seasonal trend in any of the patterns with a seasonal trend.
+For example, here are plots of trends using `tren = "cos1linear"` for different  values of `amp`:
 
-Here is an example of generating continuous exposure data with a "cos1linear"" 
-trend:
-
-```{r, fig.width = 5, fig.height = 3, fig.align = "center"}
-testexp <- sim_exposure(n=1000, central = 50, sd = 5, trend = "cos1linear",
-                        amp = .6, exposure_type = "continuous")
-head(testexp)
-qplot(testexp$date, testexp$x)+geom_point()+coord_cartesian(ylim = c(0,110)) + 
-  labs(title = "Exposure with a cos1linear trend", x = "Date", y="Exposure")+theme_bw()
+```{r fig.width = 7, fig.height = 2, echo = FALSE, message = FALSE}
+data.frame(day = 1:1000) %>%
+  mutate(`"amp = 0.5"` = calc_t(n = 1000, trend = "cos1linear", amp = 0.5),
+         `"amp = 0.1"` = calc_t(n = 1000, trend = "cos1linear", amp = 0.1),
+         `"amp = 0.9"` = calc_t(n = 1000, trend = "cos1linear", amp = 0.9),
+         `"amp = -0.5"` = calc_t(n = 1000, trend = "cos1linear", amp = -0.5)) %>%
+  gather(trend_method, trend_value, -day) %>%
+  ggplot(aes(x = day, y = trend_value)) + 
+  geom_line() + facet_wrap(~ trend_method, ncol = 4) + 
+  theme_bw()
 ```
 
-```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-calendar_plot(testexp, type = "continuous")
+The long-term trend in expected values can be changed in a similar way with the `exposure_trend` argument in `eesim`.
+
+Here is an example of generating continuous exposure data with a "cos1linear" trend for an exposure with a mean value of 50 and a standard deviation of 10:
+
+```{r, fig.width = 5.5, fig.height = 5, fig.align = "center"}
+testexp <- sim_exposure(n = 365 * 3, central = 50, sd = 10, trend = "cos1linear",
+                        exposure_type = "continuous")
+a <- ggplot(testexp, aes(x = date, y = x)) +  
+  geom_point(alpha = 0.5, size = 0.8) + 
+  coord_cartesian(ylim = c(0,110)) + 
+  labs(title = "Exposure with a 'cos1linear' trend", x = "Date", y="Exposure") + 
+  theme_classic()
+b <- calendar_plot(testexp, type = "continuous") + 
+  ggtitle("Calendar plot of simulated exposure data") + 
+  theme(legend.position = "bottom")
+grid.arrange(a, b, ncol = 1)
 ```
 
-The default amplitude for the trend is .6, but we can adjust the "amp" parameter
-for a smaller or larger trend amplitude:
+Here is an example of changing the seasonal trend by changing the value for `amp` (the default value is 0.6) to simulate exposure data for an exposure with a smaller seasonal trend and with higher exposures typical in the summer than the winter:
 
-```{r, fig.width = 5, fig.height = 3, fig.align = "center"}
-smallamp <- sim_exposure(n=1000, central = 50, sd = 5, trend = "cos1linear",
-                        amp = .2, exposure_type = "continuous")
-qplot(smallamp$date, smallamp$x) +
-  geom_point() + coord_cartesian(ylim = c(0,110)) +
-  labs(title = "Cos1linear exposure with smaller amplitude", x="Date", y="Exposure") + theme_bw()
+```{r, fig.width = 5.5, fig.height = 5,  fig.align = "center"}
+small_amp <- sim_exposure(n = 365 * 3, central = 50, sd = 10, trend = "cos1linear",
+                        amp = -0.3, exposure_type = "continuous")
+a <- ggplot(small_amp, aes(x = date, y = x)) +  
+  geom_point(alpha = 0.5, size = 0.8) + 
+  coord_cartesian(ylim = c(0,110)) + 
+  labs(title = "Exposure with a 'cos1linear' trend", x = "Date", y="Exposure") + 
+  theme_classic()
+b <- calendar_plot(small_amp, type = "continuous") + 
+  ggtitle("Calendar plot of simulated exposure data") + 
+  theme(legend.position = "bottom")
+grid.arrange(a, b, ncol = 1)
 ```
 
-```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-calendar_plot(smallamp, type = "continuous")
+The trend options are similar for binary exposure, but exclude  "curvilinear" and "cos1linear".
+Further, binary exposures can also be simulated using a "monthly" trend (`trend = "monthly"`), in which the probability of exposure can vary by month.
+When using this "monthly" trend option, the `cental` argument to `sim_exposure` should include a vector with 12 separate probabilities (the first is for January, the second for February, etc.) rather than a single probability.
+Here is an example of generating binary exposure data with a monthly trend, starting from June 1, 2002, with higher probability of the exposure in summer months than in winter months:
+
+```{r, message = FALSE, fig.width = 5.5, fig.height = 5,  fig.align = "center"}
+testbin <- sim_exposure(n=1000, central = c(.05, .05, .1, .2, .4, .4, .5, .7, .5, .2, .1, .05),
+                        trend = "monthly", exposure_type = "binary", 
+                        start.date = "2002-06-01")
+a <- testbin %>% 
+  mutate(x = factor(x, levels = c(0, 1), labels = c("Not exposed", "Exposed"))) %>% 
+  ggplot(aes(x = date, y = x)) + 
+  geom_jitter(alpha = 0.5, size = 0.7, fill = NA, width = 0, height = 0.1) + 
+  theme_classic() + 
+  labs(x = "Date", y = "Exposure")
+b <- calendar_plot(testbin, type = "discrete", labels = c("Not exposed", "Exposed")) + 
+  ggtitle("Calendar plot of simulated exposure data") + 
+  theme(legend.position = "bottom")
+grid.arrange(a, b, ncol = 1)
 ```
 
-Here is an example of generating binary exposure data with a "monthly" trend 
-starting from June 1, 2002.  The probabilities for the central parameter are for each month of the year, starting in January.
-
-```{r, message = FALSE, fig.width = 5, fig.height = 3, fig.align = "center"}
-testbin <- sim_exposure(n=1000, central = c(.1,.1,.2,.3,.4,.4,.5,.6,.5,.3,.2,.1),
-                        trend = "monthly", exposure_type = "binary", 
-                        start.date = "2002-06-01")
-head(testbin)
-qplot(testbin$date, testbin$x)+geom_point()+labs(title = "Binary exposure data with monthly trend", x="Date", y="Exposure")+theme_bw()
-```
-
-```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-calendar_plot(testbin, type = "discrete", labels = c("Not exposed", "Exposed"))
-```
-
-The following figure shows the steps taken in simulating exposure ...
+The `sim_exposure` function works by first calculating the expected exposure on any date in the simulated time series (figure below, left).
+This expected value is a mean for a continuous exposure and a probability for a binary exposure.
+The `sim_exposure` function then draws random values from the appropriate distribution (normal distribution for a continuous exposure, binomial distribution for a binary exposure) based on this day-specific expected exposure value and, in the case of continuous exposure, the standard deviation of the exposure (figure below, right).
+For continuous exposure data, the standard deviation specified in the call to `eesim` should measure the standard deviation of each point from its expected value (i.e., from the expected line shown on the left below), not the overall standard deviation of exposure values across all days in the simulated data.
 
 ```{r echo = FALSE, fig.width = 8, fig.height = 3}
-# binary_trend 
-# binary_continuous
 continuous_trend <- data_frame(t = calc_t(n = 1000, trend = "cos1")) %>% 
@@
          date = seq(from = as.Date("2000-01-01"),
-                    by = 1, length.out = 1000)) %>%
-  ggplot(aes(x = date, y = x)) + geom_point() + theme_minimal() + 
+                    by = 1, length.out = 1000)) 
+a <- continuous_trend %>%
+  ggplot(aes(x = date, y = x)) + geom_line(color = "red") + theme_classic() + 
   ylim(c(0, 200))
+
 continuous_simulated <- continuous_exposure(n = 1000, mu = 100,
-                                            sd = 10, trend = "cos1") %>%
-  ggplot(aes(x = date, y = x)) + geom_point() + theme_minimal() + 
+                                            sd = 10, trend = "cos1") 
+b <- continuous_simulated %>%
+  ggplot(aes(x = date, y = x)) + geom_point() + 
+  geom_line(data = continuous_trend, color = "red") + theme_classic() + 
   ylim(c(0, 200))
-grid.arrange(continuous_trend, continuous_simulated, ncol = 2)
+
+grid.arrange(a, b, ncol = 2)
 ```
+
+Later in this vignette, we show how you can further customize this step of generating exposure data through the use of a user-created function, allowing extensive further flexibility in simulating exposure data.
 
@@
 
-Next, the eesim package uses the exposure data to generate health outcome data with or without seasonal trends.
-eesim creates "baseline" values for outcome based on the user-specified trend and average outcome.
-Next it uses the following function to relate exposure and relative risk to outcomes:
-$$\lambda = e^{\log{(baseline)}+\log{(relative risk)}*exposure}$$
-The values of lambda are then used to randomize the outcome values around the baseline with a Poisson($\lambda\$) distribution.
+Next, the `sim_outcome` function simulates outcome data.
+The health data can have an underlying seasonal and / or long term trend in its baseline value, and then that baseline is adjusted for the risk associated with exposure, based on the generated exposure data for that day.
+The baseline outcome count for a given day ($B_t$) are based on a user-specified trend and user-specified average outcome per day over the simulated time period.
+Further, the expected outcome count on a given day is adjusted for exposure-related risk through a user-specified relative rate per unit increase in exposure ($RR$) and the simulated exposure for that day ($X_t$).
+The `eesim` function then uses the following equation to calculate the expected outcome count ($\lambda$) on a given day in the simulated time series, based on the expected baseline rate and exposure-related risk for that day:
 
-Here is an example of generating health outcome data with an upward linear trend using exposure data with a "cos1" trend:
+$$
+\log(\lambda_t) = \log{(B_t)}+\log{(RR)}*X_t
+$$
 
-```{r fig.width = 5, fig.height = 3, fig.align = "center"}
-testexp2 <- sim_exposure(n=1000, central = 100, sd = 10, trend = "cos1", exposure_type = "continuous")
-testout <- sim_outcome(exposure = testexp2, average_outcome = 22, trend = "linear", rr = 1.01)
-qplot(testout$date, testout$outcome) + geom_point() +
-  labs(title = "Health outcomes with a linear trend", x="Date", y="Outcome") + theme_bw()
+For a binary outcome, the baseline count on a given day ($B_t$) is the expected outcome count for the day if there is not an event (e.g., in a heat wave study, a non-heat wave day).
+For a continuous exposure, the baseline count on a given day ($B_t$) is the expected outcome count for the day if exposure is at its mean value.
+
+Once the expected count ($\lambda_t$) on each day of the simulated time series is calculated using this equation, the simulated count on each day is drawn as a random variable from a Poisson distribution with mean $\lambda_t$.
+Later we describe how customization can be used to simulate output counts in other ways.
+
+Here is an example of generating health outcome data with an upward linear trend using exposure data with a "cos1" trend.
+In this case, there is a steady increase in the baseline outcome count over time, as well as a seasonal trend linked to the risk associated with the seasonally-varying exposure:
+
+```{r}
+testexp2 <- sim_exposure(n = 1000, central = 100, sd = 10, trend = "cos1",
+                         exposure_type = "continuous")
+testout <- sim_outcome(exposure = testexp2, average_outcome = 22,
+                       trend = "linear", rr = 1.01)
 ```
 
-```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-calendar_plot(testout %>% select(date, outcome), type = "continuous")
+Here are plots of the resulting output:
+
+```{r echo = FALSE, fig.width = 5.5, fig.height = 5, fig.align = "center"}
+a <- ggplot(testout, aes(x = date, y = outcome)) + 
+  geom_point(alpha = 0.5, size = 0.8) +
+  labs(title = "Health outcomes with a linear trend", x = "Date", y = "Outcome") + 
+  theme_classic()
+b <- calendar_plot(testout %>% select(date, outcome), type = "continuous", 
+                   legend_name = "Outcome") + 
+  theme(legend.position = "bottom") + 
+  ggtitle("Calendar plot of simulated outcome data")
+grid.arrange(a, b, ncol = 1)
 ```
 
-###Fitting models
+As with the exposure simulation step, this step can also be extensively customized by using a user-created function. This customization will be demonstrated in a later section of the vignette.
 
-eesim next uses many generated data sets to fit statistical models to relate exposure to outcome and estimate relative risk.
-The built-in model choices are a spline model and a case-crossover model.
-The fit_mods function outputs a data frame with estimates of the log relative risk, p-values, and upper and lower 95% confidence bounds for each simulated data set.
+### Fitting models
 
-Here is an example of fitting the spline model with 7 degrees of freedom:
+Next, the `eesim` package uses this process to generate many simulated data sets and then to fit statistical models to these generated datasets.
+This step allows tests of model performance.
+You must create an R function that fits the model you'd like to fit to the simulated dataset.
+This function needs to follow certain input / output rules to work correctly in the `eesim` framework.
+First, it must input the simulated dataframe with the argument `df`.
+When writing the function, you should assume that this simulated dataframe has at least the columns `date` (in a Date format), `x` (numeric class, this gives a daily value for exposure, with 0 for unexposed and 1 for exposed in the case of binary exposure), and `outcome` (non-negative integer, this gives the simulated outcome count each day).
+Other arguments can also be passed to this function if desired.
+The function should fit a desired model to the simulated dataframe and then should return a numeric vector of length 6 with values, in this order, for the log relative risk point estimate from the model (`Estimate`), the standard error for this point estimate (`Std.
+Error`), the t-statistic for a hypothesis test with the null hypothesis that this estimate is zero (`t value`), a p-value for that test (`Pr(>|t|)`), and the lower and upper 95% confidence intervals for the point estimate (`2.5 %` and `97.5 %`).
+
+The `spline_mod` function that comes with the package is an example of such a function.
+In this case, the function fits a GLM to the simulated data, with a natural cubic spline used to control for long-term and seasonal trends in mortality.
+The function, in addition to inputing the dataframe of simulated data (`df`), also allows an argument to use to set the smoothness of the time spline (`df_year`).
+Since the function is included in the package, you can see its code by running the bare function name at the console:
 
 ```{r}
+spline_mod
+```
+
+Here are examples of applying this function to a simulated dataframe:
+
+```{r}
+# Create simulated data
 sims <- create_sims(n_reps = 10, n = 100, central = 100, sd = 10,
@@
              outcome_trend = "no trend", outcome_amp = .6, rr = 1.01)
-fits <- fit_mods(outcome = sims, model = "spline", df_year = 7)
+head(sims[[1]])
+```
+
+```{r}
+# Apply `spline_mod` to the data
+spline_mod(df = sims[[1]])
+spline_mod(df = sims[[1]], df_year = 6)
+```
+
+The `format_function` function can be used within the modeling function to get the output in the correct format if running a GLM or similar model.
+
+Once you've created the function, you can input it in a call to `eesim` using the `custom_model` argument.
+You can pass any additional arguments (`df_year` in our example) through to the function using the `custom_model_args` argument.
+This argument takes a list with the argument name and value for each argument you wish to pass to the modeling function.
+For example, the following call passes the `spline_mod` function shown above as the function to use for modeling the simulated data as well as a value for its `df_year` argument:
+
+```{r eval = FALSE}
+ex_sim2 <- eesim(n_reps = 100, n = 365 * 5, 
+                 central = c(0, 0, 0, 0, 0, 0.05, 0.12, 0.02, 0, 0, 0, 0),
+                 exposure_type = "binary", exposure_trend = "monthly",
+                 exposure_amp = -.6, average_outcome = 50,
+                 outcome_trend = "cos1", outcome_amp = 0.2, 
+                 rr = 1.05, start.date = "1996-01-01",
+                 custom_model = spline_mod, custom_model_args = list(df_year = 7))
+```
+
+The `eesim` function does this by applying the modeling function across all simulated datasets using a function called `fit_mods`.
+If you'd like, you can run that function independently.
+The `fit_mods` function outputs a data frame with estimates of the log relative risk, p-values, and upper and lower 95% confidence bounds for each simulated data set.
+
+Here is an example of fitting the spline model coded in the `spline_mod` function, with 7 degrees of freedom per year used to model long-term and seasonal trends (`df_year = 7` passed in a list to the model with the `custom_model_args` argument):
+
+```{r}
+fits <- fit_mods(data = sims, custom_model = spline_mod, 
+                 custom_model_args = list(df_year = 7))
 fits
@@
 
-###Evaluating the models
+As a note, the output of the `fit_mods` function is the output given as the second element of the list returned by a call to `eesim`.
 
-Lastly, eesim evaluates model performance with several different measures.
-The check_sims function takes the true relative risk as input and returns mean beta and relative risk estimates across all simulated data sets, variance of the estimates of beta, the mean of the variances of each beta hat, the relative bias of the mean of the beta hats, the percent coverage of the true beta, and the power of the test at the 5% significance level.
+### Evaluating the models
 
-Here is an example of the use of the check_sims function:
+The final step of the `eesim` function is to evaluate model performance across all simulations with several different measures.
+Within the `eesim` function, the `check_sims` function takes as inputs the true relative risk as well as the results from fitting the modeling function to all the simulations using the `fim_mods` function.
+It returns values for the mean effect estimate (log relative risk) and relative risk estimates across all simulated data sets, variance of the estimates of beta, the mean of the variances of each estimated log relative risk, the relative bias of the mean of the log relative risks, the percent coverage confidence intervals of the true log relative risk, and the power of the test at the 5% significance level (see the table near the beginning of the vignette).
+
+Here is an example of the use of the `check_sims` function:
 
@@
 
-##Using custom functions
+In a run of `eesim`, this output is given in the third element of the returned list.
 
-An important feature of eesim is that the user may input custom functions for any part of the simulation process.
-For example, the user may wish to generate exposure data with a custom trend, then automate the process of generating outcomes, fitting models, and evaluating performance using the built-in features of eesim.
-Functions the user has the option to customize within the eesim framework are exposure trend, outcome baseline, and outcome lambda.
-A custom model may also be used as long as returns the necessary output if the model assessmen tools in eesim are to be used.
+Internally, the functions used for this model assessment are:
 
-To use custom functions within eesim, the user must input the name of the custom function as well as a list of all arguments for the custom function and their values.
-Many inputs for the eesim functions may no longer be necessary when a custom function is used, in which case they can simply be left out.
+- `beta_bias`
+- `beta_var`
+- `coverage_beta`
+- `mean_beta`
+- `power_beta`
 
-###Custom exposure trend
+A few more details about how some of these assessments are measured are given below.
 
-Here is an example of using a custom function to specify an exposure trend that is not built in to eesim:
+### Variance across estimated log relative risk and mean variance of estimates
 
-```{r, fig.width = 5, fig.height = 3, fig.align = "center"}
-sintrend <- function(n, mu, y){
-  day <- c(1:n)
-  base <- mu + y * sin(2 * pi * (day / 365))
-  rnorm(n, mean = base, sd = 1)
-}
-customexp <- sim_exposure(n=1000, exposure_type = "continuous", cust_exp_func = sintrend, cust_exp_args = list(n=1000, mu = 75, y=3))
-head(customexp)
-qplot(customexp$date, customexp$x)+geom_point()+labs(title = "Exposure values with a custom trend", x="Date", y="Exposure")+theme_bw()
+Two values are measured by the `beta_var` function. First, the variance across all estimates of log relative risk is measured across all the simulations, using the equation:
+
+$$
+\text{variance of estimates} = E\left[\left(\hat{\beta_i} - \frac{1}{n}\sum_{i = 1}^n{\hat{\beta_i}}\right)\right]
+$$
+
+where $\hat{\beta_i}$ is the estimated log relative risk for a single simulation out of $n$ total simulations and $E$ represents the expected value.
+
+Second, the function measures the mean value of the variance estimated for $\hat{\beta}$ for each simulation:
+
+$$
+\text{mean of estimate variances} = \frac{1}{n}\sum_{i = 1}^{n}{var(\hat{\beta_i})}
+$$
+
+where  $var(\hat{\beta_i})$ is the estimated variance of the estimated log relative risk for a single simulation out of $n$ total simulations.
+
+#### Relative bias
+
+Here is the equation used by the `beta_bias` function to estimate relative bias in estimates from the simulated data:
+
+$$
+\text{relative bias} = 100*\frac{\beta - \frac{1}{n}\sum_{i = 1}^{n}{\hat{\beta}}}{\beta}
+$$
+where $\beta$ is the true log relative risk used to simulate the data and $\hat{\beta}$ is the estimated log relative risk from simulation $i$ (out of a total of $n$ simulations).
+
+### Generating power curves
+
+The other main functionality of the `eesim` package is to run through simulations under varying data generating scenarios to estimate expected power of an analysis under different scenarios.
+For example, you can explore how expected power varies for different expected effect sizes (relative risk of the outcome associated with a change in exposure) or for different average daily number of outcomes.
+This is run using the `power_calc` function in the package.
+
+The `power_calc` function allows you to put in varying values for one of the following three specifications in the simulations:
+
+- Relative risk (`rr`): How you strongly expect the exposure and outcome to be associated
+- Number of days in the study (`n`): How long you expect the study to last (or how many days of historical data you expect to be able to collect)
+- Average daily count of outcomes (`average_outcome`): For the outcome of interest, how common it is on average on days in the study (this will usually be strongly associated with the size of the population being studied)
+
+For whichever of these you choose to vary, you can specify different values to test.
+The `power_calc` function then loops through those values and runs `eesim` for each of them.
+From this, it can estimate the power for each value of the varying parameter.
+
+Here is an example of running a power calculation with varying number of days in the study (`n`).
+The `values` argument is used to specify different values of `n` we would like to test (here, it's testing power for studies with daily data for between 1 and 21 years):
+
+```{r fig.align = "center", fig.width = 5, fig.height = 3}
+pow <- power_calc(varying = "n", values = floor(365.25 * seq(1, 21, by = 5)), n_reps = 20,
+                  central = 100, sd = 10, rr = 1.001, exposure_type = "continuous",
+                  exposure_trend = "cos1", exposure_amp = .6, average_outcome = 22,
+                  outcome_trend = "no trend", outcome_amp = .6,
+                  custom_model = spline_mod, plot = TRUE)
 ```
 
-```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
-calendar_plot(customexp, type = "continuous")
+This call returns a dataframe with the estimated power for each of the values of `n` tested:
+
+```{r}
+pow
 ```
 
-This custom exposure function can be called with eesim:
+Because the argument `plot` is set to `TRUE`, it also generates a power curve plot as a side effect, as shown above.
 
-```{r, warning=F, error = TRUE}
-ex_sim2 <- eesim(n_reps = 3, n = 10, central = 100, sd = 10,
-                exposure_type = "continuous", exposure_trend = "custom", 
-                exposure_amp = .6, cust_exp_func = sintrend,
-                cust_exp_args = list(n=10, mu=75,y=3),
-                average_outcome = 22, rr = 1.01, 
-                model = "spline", df_year = 1)
-ex_sim2
+Here is another example, but this time we assume that the study will have 4,000 days of daily data, but we explore estimated power as the average daily outcome count varies:
+
+```{r fig.align = "center", fig.width = 5, fig.height = 3}
+pow2 <- power_calc(varying = "average_outcome", values = c(1, 5, 10, 20, 30, 40),
+                   n_reps = 20,
+                   central = 100, sd = 10, rr = 1.001, exposure_type = "continuous",
+                   exposure_trend = "cos1", exposure_amp = .6, n = 4000,
+                   outcome_trend = "no trend", outcome_amp = .6,
+                   custom_model = spline_mod, plot = TRUE)
 ```
 
-###Custom outcomes
+```{r}
+pow2
+```
 
-There are two ways to customize the simulated outcome data: creating a custom baseline for outcome values or customizing the relationship between outcome and exposure.
+Because these power curves and calculations are based on simulated data, there will be some randomness to results.
+Curves will be smoother as more simulations are used for each run (`n_reps`), although this will also increase the time needed to run the simulation.
 
-The outcome baseline is comprised of the values the user expects the outcomes to take on without exposure factored in.
-The user may write a function to specify the trend of the baseline, then use it as an input in sim\_outcome or eesim.
-Here is an example of creating a custom baseline function and using it in the eesim function:
+## Using custom functions
 
-```{r, warning = F}
+An important feature of `eesim` is that the user can create and use custom functions for any part of the simulation process.
+For example, the user may wish to generate exposure data with a custom trend, then automate the process of generating outcomes, fitting models, and evaluating performance using the built-in features of eesim.
+Functions the user has the option to customize within the `eesim` framework are:
+
+- The underlying expected exposure value on each day.
+Through this, the user can use customized long-term and seasonal trend patterns or can build a simulation starting from a running mean of observed exposure data.
+- How exposure values are randomized from the underlying trend.
+This allows a user to, for example, use a distribution other than normal (for continuous exposure data) or binomial (for binary data) as the underlying distribution of the exposure data.
+- The underlying pattern in the expected outcome baseline, before the influence of the exposure is added.
+Through this, the user can use a custom pattern of long-term and seasonal trends in expected health outcome rates in the simulated data.
+This functionality can also be used to include any expected influence on the baseline outcome rate from daily-varying values other than the exposure of interest.
+- How exposure influences the expected outcome rate.
+The user can create a function that inputs the expected baseline outcome count and simulated exposure levels for each day and outputs the expected outcome rate on each day, including any added or reduced risks caused by the exposure.
+This functionality can be used, for example, to simulate outcomes with a non-linear relationship with the exposure or with lagged exposure effects.
+- How outcome counts are randomized from the underlying expected outcome rate.
+This allows users to, for example, use a negative binomial or overdispersed Poisson distribution as the underlying distribution of the outcome counts.
+
+To use custom functions within `eesim`, the user must input the name of the custom function as well as a list of all arguments for the custom function and their values (examples shown below).
+This allows the user to pass the function and required arguments directly within a call to the main `eesim` function.
+When a custom function is used, many inputs that are otherwise required for the `eesim` function may no longer be necessary, in which case they can simply be left out of the `eesim` call.
+As a note, if extensive customize is required for several steps of the simulation process, it may make more sense to code the full simulation by hand rather than using the `eesim` framework.
+
+### Customizing the exposure trend
+
+To take advantage of any of the customization options, you need to write a function that follows certain input and output (i.e., interface) rules.
+First, you can use a custom function for the underlying trend in expected exposure.
+This function must take the inputs:
+
+- `n` (the number of days to simulate)
+- Either `mean` for a continuous exposure (the average value of the outcome) or `prob` for a binary exposure (the average probability of exposure)
+
+The function can take any other additional inputs, as well, but any such extra arguments (as well as `mean`) will need to be input to the `eesim` function in a list for the `cust_expdraw_args` argument (example below).
+The value for `n` will pass through directly from the `n` value specified for the call to `eesim`.
+The function must output a numeric vector that gives the simulated exposure values for each day in the simulated data.
+
+For example, the following function creates a custom exposure trend with a long-term and seasonal trend, similar to trends available through the default package options.
+However, this function specifies a minimum value that the exposure trend cannot fall below-- if the `base` exposure value is every set below this `minimum` within the algorithm, the value is reset to the `minimum` before the final values are output.
+This function can be useful in cases where the exposure cannot fall below a certain value (for example, a pollution concentration could not be lower than 0).
+This custom exposure function can also be used to customize how values are simulated from the expected exposure on each day (based on the expected distribution of the exposure).
+In the case of the example ozone concentration data from Chicago shown earlier in this vignette, we may want to simulate exposure based on the assumption that the square root of exposure is normally distributed, which will prevent negative values and may also help to simulate occasional very high values.
+
+```{r}
+above_min_trend <- function(n, mean, sd_of_sqrt, minimum = 0){
+  day <- c(1:n)
+
+  ## Calculate a baseline exposure for each day
+  base <- mean + -10 * cos(2 * pi * (day / 365))
+  base[base < minimum] <- minimum            # Reset any values below 0 to 0
+
+  ## Simulate exposure values from the baseline
+  sqrt_base <- sqrt(base)                   # Transform to square root
+  sqrt_sim <- rnorm(n, mean = sqrt_base, sd = sd_of_sqrt)
+  sqrt_sim ^ 2                              # Transform back
+}
+```
+
+Here is an example of running this custom exposure simulation function over 5 years, with a smooth line added to the plot to help show the seasonal trend included:
+
+```{r fig.width = 5, fig.height = 2, fig.align = "center"}
+above_min_trend(n = 365.25 * 5, mean = 20, minimum = 0, sd_of_sqrt = 0.9) %>% 
+  tbl_df() %>% 
+  mutate(day = 1:n()) %>% 
+  ggplot(aes(x = day, y = value)) + 
+  geom_point(alpha = 0.5, size = 0.8) + 
+  theme_classic() + 
+  geom_smooth(se = FALSE, span = 0.1, method = "loess", color = "red")
+```
+
+You can then pass this custom function into the `eesim` function using the `cust_exp_func` argument.
+The value for `n` input to the custom function will be the value you input to `eesim` for `n`.
+For any other arguments you want to pass to the function (in the function we just created, you'll want to pass values for `mean`, `minimum`, and `sd_of_sqrt`), you can include specifications for these as a list for the `cust_exp_args` argument of `eesim`.
+For example, the following call would run a simulation using this custom function for exposure:
+
+```{r message = FALSE}
+ex_sim2 <- eesim(n_reps = 1, n = round(365.25 * 5), 
+                 exposure_type = "continuous",
+                 cust_exp_func = above_min_trend,
+                 cust_exp_args = list(mean = 20, minimum = 0, sd_of_sqrt = 0.9),
+                 average_outcome = 50, rr = 1.01, 
+                 custom_model = spline_mod, custom_model_args = list(df_year = 7))
+```
+
+### Customizing the outcome simulation
+
+There are three ways to customize the simulated outcome data: creating a custom baseline for outcome values, customizing the relationship between outcome and exposure, and, as with the exposure values, customizing the randomization of the outcome values.
+
+The outcome baseline ($B_t$) is comprised of the values the user expects the outcomes to have on each day of the simulated dataset without risk associated with the exposure factored in.
+The user may write a function to specify the trend of the baseline, then use it as an input in `sim_outcome` or `eesim`.
+Here is an example of creating a custom baseline function and using it in the `eesim` function:
+
+```{r fig.width = 5, fig.height = 2, fig.align = "center"}
 custombase <- function(n, slope, intercept){
   day <- c(1:n)
-  baseline <- day*slope + intercept
+  baseline <- day * slope + intercept
   return(baseline)
@@
 
-ex_sim3 <- eesim(n_reps = 3, n = 10, central = 100, sd = 10,
+ex_sim3 <- eesim(n_reps = 3, n = 1000, central = 100, sd = 10,
                 exposure_type = "continuous", exposure_trend = "cos1",
                 exposure_amp = .6, average_outcome = 22, rr = 1.01, 
-                cust_base_func = custombase, cust_base_args = list(n=10, slope = .5,                        intercept = 12), model = "spline", df_year = 1)
-ex_sim3
+                cust_base_func = custombase,
+                cust_base_args = list(n=1000, slope = 5, intercept = 12),
+                custom_model = spline_mod, custom_model_args = list(df_year = 2))
+ggplot(ex_sim3$simulated_datasets[[1]], aes(x=date, y=outcome))+ geom_point() + geom_point(alpha = 0.5, size = 0.8) + 
+  theme_classic() + 
+  geom_smooth(se = FALSE, span = 0.1, method = "loess", color = "red")
+
 ```
 
-Here is an example of creating a custom lambda, meaning a custom function relating relative risk and exposure to outcomes, and using it in eesim with the custom baseline function created above:
+The second way of customizing the outcome simulation is to use a custom function to incorporate the added risk from the exposure when calculating the expected daily outcome count for a day, $\lambda_t$, from the inputs of exposure ($X_t$) and outcome baseline ($B_t$) for the day.
+Here is an example of creating a custom lambda, meaning a custom function relating relative risk and exposure to outcomes, and using it in eesim with the custom baseline function created above.
+The custom lambda function must input arguments `exposure`, `rr`, and `baseline` and output a vector of lambda values.
 
@@
 }
+
+ex_sim4 <- eesim(n_reps = 3, n = 1000, central = 100, sd = 10,
+                exposure_type = "continuous", exposure_trend = "cos1",
+                exposure_amp = .6, average_outcome = 22, rr = 1.01, 
+                cust_base_func = custombase,
+                cust_base_args = list(n=1000, slope = .5, intercept = 12),
+                cust_lambda_func = customlambda, cust_lambda_args = list(constant=10),
+                custom_model = spline_mod, custom_model_args = list(df_year = 2))
 ```
+
+The third way to customize the outcome simulation is to customize the randomization of the outcome values from the trend created by relating the baseline outcomes and the exposure (what we have called lambda).
+When the `cust_outdraw` argument is not specified in the `eesim` function, the function draws outcome values from a Poisson distribution with mean lambda.
+A custom function for outcome draws must input values called `n` and `lambda`, and any other arguments must be included in the `cust_outdraw_args` argument.
+Here is an example of using the custom functions to specify a negative binomial distribution for outcome randomization:
+
+```{r}
+custnbinom <- function(n, lambda, prob){
+  out <- rnbinom(n=n, size=lambda, prob=prob)
+  return(out)
+}
+
+ex_sim5 <- eesim(n_reps = 3, n = 1000, central = 100, sd = 10,
+                exposure_type = "continuous", exposure_trend = "cos1",
+                exposure_amp = .6, average_outcome = 22, rr = 1.01, 
+                cust_base_func = custombase,
+                cust_base_args = list(n=1000, slope = .5, intercept = 12),
+                cust_lambda_func = customlambda, cust_lambda_args = list(constant=10),
+                cust_outdraw = custnbinom, cust_outdraw_args = list(prob=.3), 
+                custom_model = spline_mod, custom_model_args = list(df_year = 2))
+```
+
+## References

### 2017-06-08 writing/EpiDraft.Rmd
Simulated monthly leg data, comparison plots

--- 2017-05-18 writing/EpiDraft.Rmd
@@
 
+leg_cases$mo <- month(leg_cases$week_end_date)
+leg_cases$yr <- year(leg_cases$week_end_date)
+leg_cases$moyr <- ymd(strtrim(leg_cases$week_end_date, width=7), truncated=3)
+
+#Plot of monthly cases for Pennsylvania
+realplot <- leg_cases[,c(1,4,8,9, 10)] %>% group_by(moyr) %>% summarise(sum=sum(penn, na.rm=T)) %>% ggplot(aes(x=moyr, y=sum))+ geom_bar(stat="identity") + labs(x="Date", y="Cases", title = "Monthly Legionnaire's Cases for Pennsylvania")
+
 ```
@@
 ```{r, echo=F}
+
 #Means of real data:
@@
 #(ma=14.633, nyc=4.69, pa=5.314)
-simdat <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 
-0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", 
-average_outcome = 5.3/7, custom_model = spline_mod)
+simdat1 <- eesim(n=1095, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", exposure_amp = .05, outcome_trend = "cos1linear", outcome_slope=2, outcome_amp=.8,  average_outcome = 5.3/7, custom_model = spline_mod)
+simdat2 <- eesim(n=2555, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", exposure_amp = .05, outcome_trend = "cos1linear", outcome_slope=-.8, outcome_amp=.8,  average_outcome = 5.3/7, custom_model = spline_mod)
+
+#Create new month/year variable to be able to aggregate by months
+simdat1$simulated_datasets[[1]]$date <- as.Date(1:1095, origin="2006-01-01")
+simdat1$simulated_datasets[[1]]$moyr <- ymd(strtrim(simdat1$simulated_datasets[[1]]$date, width=7), truncated=3)
+simdat2$simulated_datasets[[1]]$date <- as.Date(1:2555, origin = "2009-01-01")
+simdat2$simulated_datasets[[1]]$moyr <- ymd(strtrim(simdat2$simulated_datasets[[1]]$date, width=7), truncated=3)
+
+#Combine 2 parts
+simdat <- rbind(simdat1$simulated_datasets[[1]], simdat2$simulated_datasets[[1]])
+
+#Plot of monthly cases 
+simplot <- simdat %>% group_by(moyr) %>% summarise(sum=sum(outcome, na.rm=T)) %>% ggplot(aes(x=moyr, y=sum))+ geom_bar(stat="identity") + labs(x="Date", y="Cases", title = "Simulated Monthly Legionnaire's Cases") + ylim(0, 120)
+
+#plots side by side
+library(gridExtra)
+grid.arrange(realplot, simplot, ncol=1)
 ```

### 2017-06-08 R/sim_exposure.R

@@ -96,7 +96,7 @@ calc_t <- function(n, trend = "no trend", slope=1, amp = .6, custom_func = NULL,
 #' bin_t(n = 5, p = .3, trend = "cos1", amp = .3)
 #'
 #' @export
-bin_t <- function(n, p, trend = "no trend", slope = 1, amp = .01,
+bin_t <- function(n, p, trend = "no trend", slope = 1, amp = .05,
                   start.date = "2000-01-01", custom_func = NULL,...){
   day <- c(1:n)
   start.date <- as.Date(start.date)

### 2017-06-08 writing/EpiDraft.Rmd
Simulate leg data with custom baseline

--- 2017-05-18 writing/EpiDraft.Rmd
@@
 We will specify these proportions in the "monthly" exposure trend.
-5.3 is the weekly mean Legionnaire outcome value in Pennsylvania, and we divide by 7 to get the daily mean.
+For the outcome trend we wrote a custom baseline function which is a variation of the `cos1linear` trend option, allowing the outcomes to first have a positive slope, then a negative slope, with an outlier where the slope changes.
+These specifications allow us to simulate data that looks roughly like the real data.
 
 ```{r, echo=F}
+
 #Means of real data:
@@
 #(ma=14.633, nyc=4.69, pa=5.314)
-simdat <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 
-0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", 
-average_outcome = 5.3/7, custom_model = spline_mod)
+
+#Custom trend for outcome:
+
+custbase <- function(n1, n2, slope1, slope2, amp1, amp2){
+  day1 <- c(1:n1)
+  day2 <- c(1:n2)
+  out1 <- (slope1*(day1 / n1)) * (1 + amp1 * cos(2 * pi * (day1/ 365)))
+  out2 <- (1 + slope2*(day2 / n2)) * (1 + amp2 * cos(2 * pi * (day2/ 365)))
+  out1[(n1-10):n1] <- 2*out1[(n1-10):n1]
+  out2[1:10] <- 2*out2[1:10]
+  out <- c(out1, out2)
+  return(out)
+}
+
+simdat3 <- eesim(n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", exposure_amp = .05, cust_base_func = custbase, cust_base_args = list(n1=1095, n2=2555, slope1=1.5, slope2=-.8, amp1=.5, amp2=.7),  average_outcome = 5.3/7, custom_model = spline_mod)
+simdat3$simulated_datasets[[1]]$moyr <- ymd(strtrim(simdat3$simulated_datasets[[1]]$date, width=7), truncated=3)
+#plot
+simplot2 <- simdat3$simulated_datasets[[1]] %>% group_by(moyr) %>% summarise(sum=sum(outcome, na.rm=T)) %>% ggplot(aes(x=moyr, y=sum))+ geom_bar(stat="identity") + labs(x="Date", y="Cases", title = "Simulated Monthly Legionnaire's Cases") + ylim(0, 120)
+
+grid.arrange(realplot, simplot2, ncol=1)
+
 ```
+
+#Power calculations
+
+We can use our simulated data to perform power calculations. For the Legionnaire's disease example, we may want to study how many cases we must see monthly to have adequate power.
+
+```{r}
+power_calc(varying="average_outcome", values = c())
+```

### 2017-06-09 writing/EpiDraft.Rmd
Add power calculation to leg example

@@
 ```{r}
-power_calc(varying="average_outcome", values = c())
+power_calc(varying="average_outcome", values = c(.5, .75, 1, 2, 5), n=3650, n_reps=1, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", exposure_amp = .05, cust_base_func = custbase, cust_base_args = list(n1=1095, n2=2555, slope1=1.5, slope2=-.8, amp1=.5, amp2=.7),  custom_model = spline_mod, plot=TRUE)
 ```

### 2017-06-11 writing/EpiDraft.Rmd
Work on poster.

--- 2017-06-08 writing/EpiDraft.Rmd
@@
 
-In `eesim`, the function `power_calc` can be used to investigate the effects of various sample sizes or relative risks on power.
-This function requires arguments including the parameter to be varied (sample size or relative risk), the desired values of that parameter, and lists of the same arguments as used in the `eesim` function.
+In `eesim`, the function `power_calc` can be used to investigate the effects of various sample sizes, relative risks, or average outcome values on power.
+This function requires arguments including the parameter to be varied (sample size, relative risk, or average outcome), the desired values of that parameter, and lists of the same arguments as used in the `eesim` function.
 The `power_calc` function outputs a table of the values of the varied parameter along with their corresponding power.
@@
 ```{r}
-power_calc(varying="average_outcome", values = c())
+power_calc(varying="average_outcome", values = c(.5, .75, 1, 2, 5), n=3650, n_reps=10, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", exposure_amp = .05, cust_base_func = custbase, cust_base_args = list(n1=1095, n2=2555, slope1=1.5, slope2=-.8, amp1=.5, amp2=.7),  custom_model = spline_mod, plot=TRUE)
 ```

### 2017-06-12 writing/EpiDraft.Rmd
Change n_reps in power_calc example

@@
 ```{r}
-power_calc(varying="average_outcome", values = c(.5, .75, 1, 2, 5), n=3650, n_reps=10, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", exposure_amp = .05, cust_base_func = custbase, cust_base_args = list(n1=1095, n2=2555, slope1=1.5, slope2=-.8, amp1=.5, amp2=.7),  custom_model = spline_mod, plot=TRUE)
+power_calc(varying="average_outcome", values = c(.5, .75, 1, 2, 5, 10, 20), n=3650, n_reps=10, rr=1.1, exposure_type = "binary", central = c(0.048, 0.035, 0.046, 0.036, 0.035, 0.055, 0.053, 0.067, 0.070, 0.053, 0.033, 0.067), exposure_trend = "monthly", exposure_amp = .05, cust_base_func = custbase, cust_base_args = list(n1=1095, n2=2555, slope1=1.5, slope2=-.8, amp1=.5, amp2=.7),  custom_model = spline_mod, plot=TRUE)
 ```
